### 2021-02-16 README.Rmd
use readme rmd

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# skater

<!-- badges: start -->
<!-- badges: end -->

## Installation

Install from GitHub with:

``` r
remotes::install_github("signaturescience/skater", auth_token = "?????")
```

## Example

Lorem ipsum

### 2021-02-17 R/fam.R
add fam2ped and ped2kinpair from signaturescience/skatesim#2

@@ -1,0 +1,47 @@
+#' Fam to pedigree
+#'
+#' @param famfile PLINK-formatted fam file
+#' @param plotped Logical, TRUE if you want to create a PDF with pedigree drawings at the same location as the fam file with extension .pedigree.pdf
+#'
+#' @return A tibble with new listcol `ped` containing pedigrees from `kinship2::pedigree`.
+#' @md
+#' @export
+fam2ped <- function(famfile, plotped=TRUE) {
+  # Read in the fam file
+  fam <- readr::read_delim(famfile,
+                           delim=" ",
+                           col_names=c("fid", "id", "dadid", "momid", "sex", "affected"),
+                           col_types="ccccii")
+  # Replace 0 father and mother IDs with NA
+  fam <- fam %>% dplyr::mutate(dplyr::across(c(dadid, momid), dplyr::na_if, 0))
+  # Create a listcol with the pedigree object for each family
+  peds <-
+    fam %>%
+    tidyr::nest(data=c(-fid)) %>%
+    dplyr::mutate(ped=purrr::map(data, ~with(., kinship2::pedigree(id, dadid, momid, sex, affected))))
+  # Create a plot if plotped=TRUE
+  if (plotped) {
+    pdfoutfile <- paste0(famfile, ".pedigree.pdf")
+    message(paste("Saving PDF plots of pedigrees to file:", pdfoutfile))
+    grDevices::pdf(file = pdfoutfile, width=10, height=8, onefile=TRUE)
+    purrr::walk(peds$ped, ~plot.pedigree(., mar=c(1,4,1,4)))
+    grDevices::dev.off()
+  }
+  return(peds)
+}
+
+#' Pedigree to pairwise kinship
+#'
+#' @param ped A "pedigree" class object from \link[skater]{fam2ped}.
+#' @return A tibble containing all pairwise kinship coefficients from the input pedigree.
+#' @md
+#' @export
+ped2kinpair <- function(ped) {
+  if (class(ped)!="pedigree") stop("Input must be of class 'pedigree'.")
+  kinship2::kinship(ped) %>%
+    corrr::as_cordf(diagonal=0.5) %>%
+    corrr::shave() %>%
+    corrr::stretch() %>%
+    stats::na.omit() %>%
+    dplyr::transmute(id1=pmin(x, y), id2=pmax(x, y), k=r)
+}

### 2021-02-17 R/skater-package.R

@@ -8,4 +8,13 @@
 NULL
 
 # Quiet CRAN check notes for unbound global variables
-if(getRversion() >= "2.15.1")  utils::globalVariables(c("."))
+if(getRversion() >= "2.15.1")  utils::globalVariables(c(".",
+                                                        "fid",
+                                                        "dadid",
+                                                        "momid",
+                                                        "dadid",
+                                                        "dadid",
+                                                        "x",
+                                                        "y",
+                                                        "r",
+                                                        "data"))

### 2021-02-17 R/fam.R
explicitly use kinship2::

@@ -24,7 +24,7 @@ fam2ped <- function(famfile, plotped=TRUE) {
     pdfoutfile <- paste0(famfile, ".pedigree.pdf")
     message(paste("Saving PDF plots of pedigrees to file:", pdfoutfile))
     grDevices::pdf(file = pdfoutfile, width=10, height=8, onefile=TRUE)
-    purrr::walk(peds$ped, ~plot.pedigree(., mar=c(1,4,1,4)))
+    purrr::walk(peds$ped, ~kinship2::plot.pedigree(., mar=c(1,4,1,4)))
     grDevices::dev.off()
   }
   return(peds)

### 2021-02-18 R/fam.R
roxygen2md

@@ -4,7 +4,6 @@
 #' @param plotped Logical, TRUE if you want to create a PDF with pedigree drawings at the same location as the fam file with extension .pedigree.pdf
 #'
 #' @return A tibble with new listcol `ped` containing pedigrees from `kinship2::pedigree`.
-#' @md
 #' @export
 fam2ped <- function(famfile, plotped=TRUE) {
   # Read in the fam file
@@ -34,9 +33,8 @@ fam2ped <- function(famfile, plotped=TRUE) {
 
 #' Pedigree to pairwise kinship
 #'
-#' @param ped A "pedigree" class object from \link[skater]{fam2ped}.
+#' @param ped A "pedigree" class object from [fam2ped][skater::fam2ped].
 #' @return A tibble containing all pairwise kinship coefficients from the input pedigree.
-#' @md
 #' @export
 ped2kinpair <- function(ped) {
   if (class(ped)!="pedigree") stop("Input must be of class 'pedigree'.")

### 2021-02-18 R/fam.R
improve docs

@@ -1,5 +1,9 @@
 #' Fam to pedigree
 #'
+#' Converts a [PLINK-formatted fam file](https://www.cog-genomics.org/plink/1.9/formats#fam) to a pedigree object using [kinship2::pedigree].
+#'
+#' `r lifecycle::badge("experimental")`
+#'
 #' @param famfile PLINK-formatted fam file
 #' @param plotped Logical, TRUE if you want to create a PDF with pedigree drawings at the same location as the fam file with extension .pedigree.pdf
 #'
@@ -32,6 +36,10 @@ fam2ped <- function(famfile, plotped=TRUE) {
 }
 
 #' Pedigree to pairwise kinship
+#'
+#' Converts a pedigree class object from [fam2ped][skater::fam2ped] to a pairwise list of relationships and their expected/theoretical kinship coefficient.
+#'
+#' `r lifecycle::badge("experimental")`
 #'
 #' @param ped A "pedigree" class object from [fam2ped][skater::fam2ped].
 #' @return A tibble containing all pairwise kinship coefficients from the input pedigree.

### 2021-02-19 R/skater-package.R
suppress note from #2 they way it's done in tidyverse https://github.com/tidyverse/tidyverse/blob/62710238560b04ea2a9fe3f3cc70ceede5aa4273/R/tidyverse.R#L4-L23

@@ -7,6 +7,9 @@
 ## usethis namespace: end
 NULL
 
+# Suppress R CMD check note
+#' @importFrom lifecycle badge
+
 # Quiet CRAN check notes for unbound global variables
 if(getRversion() >= "2.15.1")  utils::globalVariables(c(".",
                                                         "fid",

### 2021-02-19 R/skater-package.R
add degree tibble (dibble) function #16

@@ -11,7 +11,7 @@ NULL
 #' @importFrom lifecycle badge
 
 # Quiet CRAN check notes for unbound global variables
-if(getRversion() >= "2.15.1")  utils::globalVariables(c(".",
+if(getRversion() >= "2.15.1")  utils::globalVariables(c("data",
                                                         "fid",
                                                         "dadid",
                                                         "momid",
@@ -20,4 +20,8 @@ if(getRversion() >= "2.15.1")  utils::globalVariables(c(".",
                                                         "x",
                                                         "y",
                                                         "r",
-                                                        "data"))
+                                                        "k",
+                                                        "u",
+                                                        "l",
+                                                        "degree",
+                                                        "."))

### 2021-02-19 R/utils.R

@@ -1,0 +1,30 @@
+#' Degree tibble
+#'
+#' Creates a tibble with degree, expected kinship coefficient, and inference boundaries.
+#'
+#' Rows will be created up to the `max_degree`, with an additional row for any
+#' relationship more distant than `max_degree`. The `degree` value for the final
+#' row will be `NA`. This represents inference criteria for "unrelated"
+#' individuals. See examples.
+#'
+#' @param max_degree The most distant degree you want to measure.
+#'
+#' @return A tibble containing the degree, expected kinship coefficient (`k`),
+#'   lower (`l`) and upper (`u`) inference bounds for that degree.
+#' @export
+#'
+#' @examples
+#' dibble(3)
+#' dibble(10)
+dibble <- function(max_degree=3L) {
+  stopifnot(is.numeric(max_degree))
+  max_degree <- as.integer(round(max_degree))
+  tibble::tibble(degree=0:(max_degree+1)) %>%
+    dplyr::mutate(k=.5^(degree+1)) %>%
+    dplyr::mutate(l=sqrt(dplyr::lead(k, default=dplyr::last(k)/2)  * k)) %>%
+    dplyr::mutate(u=sqrt(dplyr::lag(k,  default=dplyr::first(k)*2) * k)) %>%
+    dplyr::mutate(u=ifelse(dplyr::row_number()==1L, 1, u)) %>%
+    dplyr::mutate(l=ifelse(dplyr::row_number()==dplyr::n(), -1, l)) %>%
+    dplyr::mutate(degree=ifelse(dplyr::row_number()==dplyr::n(), NA_integer_, degree)) %>%
+    dplyr::mutate(k=ifelse(dplyr::row_number()==dplyr::n(), 0, k))
+}

### 2021-02-20 R/utils.R
initial work on infer degree function #16 not working

@@ -11,11 +11,12 @@
 #'
 #' @return A tibble containing the degree, expected kinship coefficient (`k`),
 #'   lower (`l`) and upper (`u`) inference bounds for that degree.
-#' @export
 #'
 #' @examples
 #' dibble(3)
 #' dibble(10)
+#'
+#' @export
 dibble <- function(max_degree=3L) {
   stopifnot(is.numeric(max_degree))
   max_degree <- as.integer(round(max_degree))
@@ -27,4 +28,28 @@ dibble <- function(max_degree=3L) {
     dplyr::mutate(l=ifelse(dplyr::row_number()==dplyr::n(), -1, l)) %>%
     dplyr::mutate(degree=ifelse(dplyr::row_number()==dplyr::n(), NA_integer_, degree)) %>%
     dplyr::mutate(k=ifelse(dplyr::row_number()==dplyr::n(), 0, k))
+}
+
+
+#' Infer degree for a kinship coefficient
+#'
+#' @param k Kinship coefficient (numeric, typically between 0 and .5, although KING can produce values <0).
+#' @param degree_resolution The maximum degree relationship you want to infer.
+#'   Defaults to third-degree relatives. Anything lower than `degree_resolution`
+#'   will produce `NA` (i.e., unrelated). See [dibble][skater::dibble].
+#'
+#' @return Inferred degree, up to the `degree_resolution` (anything more distant is `NA`, i.e., unrelated).
+#'
+#' @examples
+#' infer_degree(0.0625)
+#' infer_degree(0.125)
+#' infer_degree(0.25)
+#' infer_degree(0.5)
+#'
+#' @export
+infer_degree <- function(k, degree_resolution=3L) {
+  stopifnot(is.numeric(k))
+  stopifnot(k %>% dplyr::between(-1, 1))
+  d <- dibble(degree_resolution)
+  as.integer(d$degree[which(purrr::map2_lgl(d$l, d$u, ~dplyr::between(k, .x, .y)))])
 }

### 2021-02-20 R/utils.R
pull dibble contruction out of infer_degree

@@ -9,8 +9,8 @@
 #'
 #' @param max_degree The most distant degree you want to measure.
 #'
-#' @return A tibble containing the degree, expected kinship coefficient (`k`),
-#'   lower (`l`) and upper (`u`) inference bounds for that degree.
+#' @return A tibble (class "dibble") containing the degree, expected kinship
+#'   coefficient (`k`), lower (`l`) and upper (`u`) inference bounds.
 #'
 #' @examples
 #' dibble(3)
@@ -27,29 +27,33 @@ dibble <- function(max_degree=3L) {
     dplyr::mutate(u=ifelse(dplyr::row_number()==1L, 1, u)) %>%
     dplyr::mutate(l=ifelse(dplyr::row_number()==dplyr::n(), -1, l)) %>%
     dplyr::mutate(degree=ifelse(dplyr::row_number()==dplyr::n(), NA_integer_, degree)) %>%
-    dplyr::mutate(k=ifelse(dplyr::row_number()==dplyr::n(), 0, k))
+    dplyr::mutate(k=ifelse(dplyr::row_number()==dplyr::n(), 0, k)) %>%
+    structure(class=c("dibble", class(.)))
 }
 
 
 #' Infer degree for a kinship coefficient
 #'
 #' @param k Kinship coefficient (numeric, typically between 0 and .5, although KING can produce values <0).
-#' @param degree_resolution The maximum degree relationship you want to infer.
-#'   Defaults to third-degree relatives. Anything lower than `degree_resolution`
-#'   will produce `NA` (i.e., unrelated). See [dibble][skater::dibble].
+#' @param dibble Degree tibble from [dibble][skater::dibble].
 #'
-#' @return Inferred degree, up to the `degree_resolution` (anything more distant is `NA`, i.e., unrelated).
+#' @return Inferred degree, up to the maximum degree in `dibble` (anything more distant is `NA`, i.e., unrelated).
 #'
 #' @examples
-#' infer_degree(0.0625)
-#' infer_degree(0.125)
-#' infer_degree(0.25)
-#' infer_degree(0.5)
-#'
+#' d3 <- dibble(3)
+#' infer_degree(0.5, d3)
+#' infer_degree(0.25, d3)
+#' infer_degree(0.125, d3)
+#' infer_degree(0.0625, d3)
+#' infer_degree(0, d3)
[…]

### 2021-02-20 R/utils.R
additional examples

@@ -47,7 +47,11 @@ dibble <- function(max_degree=3L) {
 #' infer_degree(0.0625, d3)
 #' infer_degree(0, d3)
 #' infer_degree(-0.05, d3)
+#' # Run on multiple kinship coefficients
 #' k <- seq(.02, .5, .03)
+#' # Not vectorized, this will not work:
+#' # infer_degree(k, d3)
+#' # This does:
 #' purrr::map_int(k, infer_degree, d3)
 #' tibble::tibble(k=k) %>% dplyr::mutate(degree=purrr::map_int(k, infer_degree, d3))
 #' @export
@@ -57,3 +61,4 @@ infer_degree <- function(k, dibble) {
   stopifnot("dibble" %in% class(dibble))
   as.integer(dibble$degree[which(purrr::map2_lgl(dibble$l, dibble$u, ~dplyr::between(k, .x, .y)))])
 }
+

### 2021-02-20 R/utils.R
vectorize

@@ -47,18 +47,13 @@ dibble <- function(max_degree=3L) {
 #' infer_degree(0.0625, d3)
 #' infer_degree(0, d3)
 #' infer_degree(-0.05, d3)
-#' # Run on multiple kinship coefficients
 #' k <- seq(.02, .5, .03)
-#' # Not vectorized, this will not work:
-#' # infer_degree(k, d3)
-#' # This does:
-#' purrr::map_int(k, infer_degree, d3)
-#' tibble::tibble(k=k) %>% dplyr::mutate(degree=purrr::map_int(k, infer_degree, d3))
+#' infer_degree(k, d3)
+#' tibble::tibble(k=k) %>% dplyr::mutate(degree=infer_degree(k, d3))
 #' @export
 infer_degree <- function(k, dibble) {
-  stopifnot(is.numeric(k))
-  stopifnot(k %>% dplyr::between(-1, 1))
+  stopifnot(all(is.numeric(k)))
+  stopifnot(all(k %>% dplyr::between(-1, 1)))
   stopifnot("dibble" %in% class(dibble))
-  as.integer(dibble$degree[which(purrr::map2_lgl(dibble$l, dibble$u, ~dplyr::between(k, .x, .y)))])
+  vapply(k, function(k) as.integer(dibble$degree[which(purrr::map2_lgl(dibble$l, dibble$u, ~dplyr::between(k, .x, .y)))]), FUN.VALUE=1L)
 }
-

### 2021-02-20 R/utils.R
vectorize with vapply (rename)

@@ -9,8 +9,8 @@
 #'
 #' @param max_degree The most distant degree you want to measure.
 #'
-#' @return A tibble (class "dibble") containing the degree, expected kinship
-#'   coefficient (`k`), lower (`l`) and upper (`u`) inference bounds.
+#' @return A tibble containing the degree, expected kinship coefficient (`k`),
+#'   lower (`l`) and upper (`u`) inference bounds.
 #'
 #' @examples
 #' dibble(3)
@@ -27,33 +27,39 @@ dibble <- function(max_degree=3L) {
     dplyr::mutate(u=ifelse(dplyr::row_number()==1L, 1, u)) %>%
     dplyr::mutate(l=ifelse(dplyr::row_number()==dplyr::n(), -1, l)) %>%
     dplyr::mutate(degree=ifelse(dplyr::row_number()==dplyr::n(), NA_integer_, degree)) %>%
-    dplyr::mutate(k=ifelse(dplyr::row_number()==dplyr::n(), 0, k)) %>%
-    structure(class=c("dibble", class(.)))
+    dplyr::mutate(k=ifelse(dplyr::row_number()==dplyr::n(), 0, k))
 }
 
 
-#' Infer degree for a kinship coefficient
+#' Kinship coefficient to degree
+#'
+#' Infers relationship degree given a kinship coefficient.
 #'
 #' @param k Kinship coefficient (numeric, typically between 0 and .5, although KING can produce values <0).
-#' @param dibble Degree tibble from [dibble][skater::dibble].
+#' @param max_degree Max degree resolution (default 3). Used to seed
+#'   [dibble][skater::dibble]. Anything below the inference range of
+#'   `max_degree` will report `NA`. See [dibble][skater::dibble].
 #'
 #' @return Inferred degree, up to the maximum degree in `dibble` (anything more distant is `NA`, i.e., unrelated).
 #'
 #' @examples
-#' d3 <- dibble(3)
-#' infer_degree(0.5, d3)
-#' infer_degree(0.25, d3)
-#' infer_degree(0.125, d3)
-#' infer_degree(0.0625, d3)
-#' infer_degree(0, d3)
-#' infer_degree(-0.05, d3)
+#' kin2degree(0.5)
+#' kin2degree(0.25)
+#' kin2degree(0.125)
+#' kin2degree(0.0625)
[…]

### 2021-02-20 R/fam.R
whitespace

@@ -8,6 +8,7 @@
 #' @param plotped Logical, TRUE if you want to create a PDF with pedigree drawings at the same location as the fam file with extension .pedigree.pdf
 #'
 #' @return A tibble with new listcol `ped` containing pedigrees from `kinship2::pedigree`.
+#'
 #' @export
 fam2ped <- function(famfile, plotped=TRUE) {
   # Read in the fam file
@@ -42,7 +43,9 @@ fam2ped <- function(famfile, plotped=TRUE) {
 #' `r lifecycle::badge("experimental")`
 #'
 #' @param ped A "pedigree" class object from [fam2ped][skater::fam2ped].
+#'
 #' @return A tibble containing all pairwise kinship coefficients from the input pedigree.
+#'
 #' @export
 ped2kinpair <- function(ped) {
   if (class(ped)!="pedigree") stop("Input must be of class 'pedigree'.")

### 2021-02-20 tests/testthat.R
use_testthat

@@ -1,0 +1,4 @@
+library(testthat)
+library(skater)
+
+test_check("skater")

### 2021-02-21 tests/testthat/test-utils.R
test kinship inference fixes #18

@@ -1,0 +1,10 @@
+library(skater)
+library(testthat)
+
+test_that("Degree inference from kinship coefficient", {
+  expect_equal(kin2degree(.25), 1L)
+  expect_equal(kin2degree(.125), 2L)
+  expect_equal(kin2degree(.0625), 3L)
+  expect_identical(kin2degree(0), NA_integer_)
+  expect_identical(kin2degree(c(.5, .25, .125, .0625, .03125)), c(0L, 1L, 2L, 3L, NA))
+})

### 2021-02-21 R/utils.R
errors/warnings

@@ -7,7 +7,7 @@
 #' row will be `NA`. This represents inference criteria for "unrelated"
 #' individuals. See examples.
 #'
-#' @param max_degree The most distant degree you want to measure.
+#' @param max_degree The most distant degree you want to measure (usually between 3-9, default 3).
 #'
 #' @return A tibble containing the degree, expected kinship coefficient (`k`),
 #'   lower (`l`) and upper (`u`) inference bounds.
@@ -18,7 +18,11 @@
 #'
 #' @export
 dibble <- function(max_degree=3L) {
-  stopifnot(is.numeric(max_degree))
+  if (!is.numeric(max_degree)) stop("max_degree must be numeric")
+  if (max_degree<1) stop("max_degree must be >=1")
+  if (max_degree<3) warning("max_degree should be >=3")
+  if (max_degree>=12) stop("max_degree must be <12")
+  if (max_degree>=10) warning("max_degree should be <10")
   max_degree <- as.integer(round(max_degree))
   tibble::tibble(degree=0:(max_degree+1)) %>%
     dplyr::mutate(k=.5^(degree+1)) %>%

### 2021-02-21 tests/testthat/test-utils.R
test dibble #18

@@ -1,6 +1,14 @@
 library(skater)
 library(testthat)
 
+test_that("Degree tibble", {
+  expect_equal(3:9 %>% purrr::map(dibble) %>% purrr::map_int(nrow), (3:9)+2)
+  expect_error(dibble(12))
+  expect_warning(dibble(10))
+  expect_error(dibble(0))
+  expect_warning(dibble(2))
+})
+
 test_that("Degree inference from kinship coefficient", {
   expect_equal(kin2degree(.25), 1L)
   expect_equal(kin2degree(.125), 2L)

### 2021-02-21 R/fam.R
add read_fam #20

@@ -1,3 +1,30 @@
+#' Read PLINK-formatted .fam file
+#'
+#' Reads in a [PLINK-formatted .fam file](https://www.cog-genomics.org/plink/1.9/formats#fam). Input `file` must have six columns:
+#' 1. Family ID
+#' 2. Individual ID
+#' 3. Father ID
+#' 4. Mother ID
+#' 5. Sex
+#' 6. Affected Status
+#'
+#' @param file Input file path
+#'
+#' @return A tibble containing the 6 columns from the fam file.
+#'
+#' @examples
+#' famfile <- system.file("extdata", "3gens.fam", package="skater", mustWork=TRUE)
+#' read_fam(famfile)
+#'
+#' @export
+read_fam <- function(file) {
+  readr::read_delim(famfile,
+                    delim=" ",
+                    col_names=c("fid", "id", "dadid", "momid", "sex", "affected"),
+                    col_types="ccccii")
+}
+
+
 #' Fam to pedigree
 #'
 #' Converts a [PLINK-formatted fam file](https://www.cog-genomics.org/plink/1.9/formats#fam) to a pedigree object using [kinship2::pedigree].

### 2021-02-21 R/fam.R
split read fam / plot ped fixes #20

@@ -14,11 +14,12 @@
 #'
 #' @examples
 #' famfile <- system.file("extdata", "3gens.fam", package="skater", mustWork=TRUE)
-#' read_fam(famfile)
+#' fam <- read_fam(famfile)
+#' fam
 #'
 #' @export
 read_fam <- function(file) {
-  readr::read_delim(famfile,
+  readr::read_delim(file,
                     delim=" ",
                     col_names=c("fid", "id", "dadid", "momid", "sex", "affected"),
                     col_types="ccccii")
@@ -31,18 +32,17 @@ read_fam <- function(file) {
 #'
 #' `r lifecycle::badge("experimental")`
 #'
-#' @param famfile PLINK-formatted fam file
-#' @param plotped Logical, TRUE if you want to create a PDF with pedigree drawings at the same location as the fam file with extension .pedigree.pdf
+#' @param fam A tibble with six columns of PLINK .fam data as read in by [read_fam].
 #'
 #' @return A tibble with new listcol `ped` containing pedigrees from `kinship2::pedigree`.
 #'
+#' @examples
+#' famfile <- system.file("extdata", "3gens.fam", package="skater", mustWork=TRUE)
+#' fam <- read_fam(famfile)
+#' fam2ped(fam)
+#'
 #' @export
-fam2ped <- function(famfile, plotped=TRUE) {
-  # Read in the fam file
-  fam <- readr::read_delim(famfile,
-                           delim=" ",
-                           col_names=c("fid", "id", "dadid", "momid", "sex", "affected"),
-                           col_types="ccccii")
+fam2ped <- function(fam) {
   # Replace 0 father and mother IDs with NA
   fam <- fam %>% dplyr::mutate(dplyr::across(c(dadid, momid), dplyr::na_if, 0))
   # Make affected status 1
@@ -52,15 +52,25 @@ fam2ped <- function(famfile, plotped=TRUE) {
     fam %>%
     tidyr::nest(data=c(-fid)) %>%
     dplyr::mutate(ped=purrr::map(data, ~with(., kinship2::pedigree(id, dadid, momid, sex, affected))))
-  # Create a plot if plotped=TRUE
-  if (plotped) {
-    pdfoutfile <- paste0(famfile, ".pedigree.pdf")
-    message(paste("Saving PDF plots of pedigrees to file:", pdfoutfile))
[…]

### 2021-02-21 R/fam.R
whitespace

@@ -73,6 +73,7 @@ plot_pedigree <- function(ped, file=NULL, width=10, height=8) {
   grDevices::dev.off()
 }
 
+
 #' Pedigree to pairwise kinship
 #'
 #' Converts a pedigree class object from [fam2ped][skater::fam2ped] to a pairwise list of relationships and their expected/theoretical kinship coefficient.

### 2021-02-21 R/fam.R
minor roxygen formatting

@@ -1,5 +1,6 @@
 #' Read PLINK-formatted .fam file
 #'
+#' @description
 #' Reads in a [PLINK-formatted .fam file](https://www.cog-genomics.org/plink/1.9/formats#fam). Input `file` must have six columns:
 #' 1. Family ID
 #' 2. Individual ID
@@ -28,9 +29,11 @@ read_fam <- function(file) {
 
 #' Fam to pedigree
 #'
+#' @description
+#' `r lifecycle::badge("experimental")`
+#'
 #' Converts a [PLINK-formatted fam file](https://www.cog-genomics.org/plink/1.9/formats#fam) to a pedigree object using [kinship2::pedigree].
 #'
-#' `r lifecycle::badge("experimental")`
 #'
 #' @param fam A tibble with six columns of PLINK .fam data as read in by [read_fam].
 #'
@@ -76,9 +79,11 @@ plot_pedigree <- function(ped, file=NULL, width=10, height=8) {
 
 #' Pedigree to pairwise kinship
 #'
+#' @description
+#' `r lifecycle::badge("experimental")`
+#'
 #' Converts a pedigree class object from [fam2ped][skater::fam2ped] to a pairwise list of relationships and their expected/theoretical kinship coefficient.
 #'
-#' `r lifecycle::badge("experimental")`
 #'
 #' @param ped A "pedigree" class object from [fam2ped][skater::fam2ped].
 #'

### 2021-02-21 R/utils.R

@@ -1,5 +1,6 @@
 #' Degree tibble
 #'
+#' @description
 #' Creates a tibble with degree, expected kinship coefficient, and inference boundaries.
 #'
 #' Rows will be created up to the `max_degree`, with an additional row for any
@@ -37,6 +38,7 @@ dibble <- function(max_degree=3L) {
 
 #' Kinship coefficient to degree
 #'
+#' @description
 #' Infers relationship degree given a kinship coefficient.
 #'
 #' @param k Kinship coefficient (numeric, typically between 0 and .5, although KING can produce values <0).

### 2021-02-21 README.Rmd
update readme fixes #8

@@
 <!-- badges: start -->
+[![R-CMD-check](https://github.com/signaturescience/skater/workflows/R-CMD-check/badge.svg)](https://github.com/signaturescience/skater/actions)
 <!-- badges: end -->
+
+**S**NP-based **K**inship **A**nalysis **T**esting and **E**valuation: miscellaneous **R** utilties.
 
@@
 
-Install from GitHub with:
+Install from GitHub:
 
 ``` r
-remotes::install_github("signaturescience/skater", auth_token = "?????")
+remotes::install_github("signaturescience/skater", auth_token = github_pat(quiet))
 ```
 
-## Example
+A GitHub [personal access token](https://github.com/settings/tokens) must be in the environment variable `GITHUB_PAT`.
 
-Lorem ipsum
+## Examples
+
+```{r}
+library(skater)
+```
+
+### Pedigrees and PLINK .fam files
+
+Read a PLINK-formatted .fam file:
+
+```{r}
+famfile <- system.file("extdata", "3gens.fam", package="skater", mustWork=TRUE)
+fam <- read_fam(famfile)
+fam
+```
+
+Convert each family into a pedigree object (see the [kinship2 vignette](https://cran.r-project.org/web/packages/kinship2/vignettes/pedigree.html)):
+
+```{r}
+peds <- fam2ped(fam)
+peds
+```
+
+In the example above, the resulting tibble is nested by family ID.
+The `data` column contains the individual family information, while the `ped` column contains the pedigree object for that family.
+You can unnest any particular family:
+
+```{r}
+peds %>% 
+  dplyr::filter(fid=="testped1") %>% 
+  tidyr::unnest(cols=data)
+```
+
+You can also look at a single pedigree:
+
+```{r}
+peds$ped[[1]]
+```
+
+Or plot that pedigree:
+
+```{r plotped}
+plot(peds$ped[[1]], mar=c(1,4,1,4))
+```
+
+The `plot_pedigree` function in the skater package will walk over a list of pedigree objects, writing a multi-page PDF to file, with each page containing a pedigree from each of the families:
+
+```{r, eval=FALSE}
+plot_pedigree(peds$ped, file="3gens.ped.pdf")
+```
+
+The `ped2kinpair()` function takes a pedigree object and produces a pairwise list of relationships with their expected kinship coefficient.
+
+Run on a single family:
+
+```{r}
+ped2kinpair(peds$ped[[1]])
+```
+
+Map over all families:
+
+```{r}
+kinpairs <- 
+  peds %>% 
+  dplyr::mutate(pairs=purrr::map(ped, ped2kinpair)) %>% 
+  dplyr::select(fid, pairs) %>% 
+  tidyr::unnest(cols=pairs)
+kinpairs
+```
+
+Note that this maps `ped2kinpair` over all `ped` objects in the input tibble, and that relationships are not shown for between-family relationships (which should all be zero).
+
+### Degree inference
+
+The `dibble()` function creates a **d**egree **i**nference tibble, with degrees up to the specified `max_degree` (default=3), expected kinship coefficient, and lower (`l`) and upper (`u`) inference ranges (see [Manichaikul 2010](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3025716/)).
+Degree 0 corresponds to self / identity / MZ twins, with an expected kinship coefficient of 0.5, with inference range >=0.354.
+Anything beyond the maximum degree resolution is considered unrelated (degree `NA`), with expected kinship coefficient of 0.
+
+```{r}
+dibble()
+```
+
+The degree inference `max_degree` default is 3. Change this argument to allow more granular degree inference ranges:
+
+```{r}
+dibble(max_degree = 5)
+```
+
+Note that the distance between relationship degrees becomes smaller as the relationship degree becomes more distant.
+The `dibble()` function will throw a warning with `max_degree` >=10, and will stop with an error at >=12.
+
+The `kin2degree()` function infers the relationship degree given a kinship coefficient and a `max_degree` up to which anything more distant is treated as unrelated. Example first degree relative:
+
+```{r}
+kin2degree(.25, max_degree=3)
+```
+
+Example 4th degree relative, but using the default max_degree resolution of 3:
+
+```{r}
+kin2degree(.0312, max_degree=3)
+```
+
+Example 4th degree relative, but increasing the degree resolution:
+
+```{r}
+kin2degree(.0312, max_degree=5)
+```
+
+The `kin2degree()` function is vectorized over values of `k`, so can be used inside of a `mutate` on a tibble of kinship coefficients:
+
+```{r}
+# Get two pairs from each type of relationship we have in kinpairs:
+kinpairs_subset <- 
+  kinpairs %>% 
+  dplyr::group_by(k) %>% 
+  dplyr::slice(1:2)
+kinpairs_subset
+
+# Infer degree out to third degree relatives:
+kinpairs_subset %>% 
+  dplyr::mutate(degree=kin2degree(k, max_degree=3))
+```

### 2021-02-22 data-raw/generate_sysdata.R
1000G unrelateds fixes #11

@@ -1,0 +1,26 @@
+## code to prepare `DATASET` dataset goes here
+
+library(tidyverse)
+
+# Read ped file from 1kg FTP
+ped1kg <- read_tsv("http://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/integrated_call_samples_v3.20200731.ALL.ped", guess_max = 4000)
+# Clean names
+ped1kg <- ped1kg %>% janitor::clean_names()
+# make first 6 names same as result from read_fam
+names(ped1kg)[1:6] <- c("fid", "id", "dadid", "momid", "sex", "affected")
+# limit to 2504 with sequencing data
+ped1kg <- ped1kg %>% filter(phase_3_genotypes %>% near(1))
+# Remove unneeded columns
+ped1kg <- ped1kg %>% select(-phase_3_genotypes, -related_genotypes, -omni_genotypes, -affy_genotypes)
+
+# Get all samples unrelated to anyone
+ped1kg_unrel <-
+  ped1kg %>%
+  filter(dadid=="0") %>%
+  filter(momid=="0") %>%
+  filter(siblings=="0") %>%
+  filter(second_order=="0") %>%
+  filter(third_order=="0") %>%
+  filter(children=="0")
+
+usethis::use_data(ped1kg, ped1kg_unrel, internal = TRUE, overwrite = TRUE)

### 2021-02-22 data-raw/generate_sysdata.R
documentation #11

@@ -1,6 +1,6 @@
-## code to prepare `DATASET` dataset goes here
-
 library(tidyverse)
+
+## 1000 Genomes pedigree
 
 # Read ped file from 1kg FTP
 ped1kg <- read_tsv("http://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/integrated_call_samples_v3.20200731.ALL.ped", guess_max = 4000)

### 2021-02-22 data-raw/generate_sysdata.R
Write out individual files containing IDs for unrelated individuals #11

@@ -23,4 +23,12 @@ ped1kg_unrel <-
   filter(third_order=="0") %>%
   filter(children=="0")
 
+# Write out individual files containing IDs for unrelated individuals
+unrel_path <- here::here("data-raw/1000g_unrel")
+dir.create(unrel_path, showWarnings = FALSE)
+for(pop in unique(ped1kg_unrel$population)) {
+  write_lines(ped1kg_unrel %>% filter(population==pop) %>% pull(id),
+              paste0(file.path(unrel_path, pop), ".txt"))
+}
+
 usethis::use_data(ped1kg, ped1kg_unrel, internal = TRUE, overwrite = TRUE)

### 2021-02-22 README.Rmd
updated documentation with package data info

@@
 ```
+
+### Package data
+
+Unexported objects `ped1kg` and `ped1kg_unrel` contain pedigree data obtained from 1000 Genomes FTP (http://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/integrated_call_samples_v3.20200731.ALL.ped) for all 2504 individuals with sequencing data, and all individuals unrelated to anyone else in the data (momid, dadid, siblings, second_order, and third_order all =0).
+See [data-raw/generate_sysdata.R](data-raw/generate_sysdata.R) for details.
+
+```{r}
+skater:::ped1kg
+skater:::ped1kg_unrel
+```

### 2021-03-08 data-raw/generate_sysdata.R
less restrictive finding unrelateds #22

@@ -1,8 +1,6 @@
 library(tidyverse)
 
-## 1000 Genomes pedigree
-
-# Read ped file from 1kg FTP
+# Read 1000Gped file from 1kg FTP
 ped1kg <- read_tsv("http://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/integrated_call_samples_v3.20200731.ALL.ped", guess_max = 4000)
 # Clean names
 ped1kg <- ped1kg %>% janitor::clean_names()
@@ -13,22 +11,26 @@ ped1kg <- ped1kg %>% filter(phase_3_genotypes %>% near(1))
 # Remove unneeded columns
 ped1kg <- ped1kg %>% select(-phase_3_genotypes, -related_genotypes, -omni_genotypes, -affy_genotypes)
 
-# Get all samples unrelated to anyone
-ped1kg_unrel <-
-  ped1kg %>%
-  filter(dadid=="0") %>%
-  filter(momid=="0") %>%
-  filter(siblings=="0") %>%
-  filter(second_order=="0") %>%
-  filter(third_order=="0") %>%
-  filter(children=="0")
+# Who are the relatives in the ped file (e.g., mom, dad, sibs, 2d, 3d, and children)?
+relatives <-
+  with(ped1kg, c(dadid, momid, siblings, second_order, third_order, children)) %>%
+  strsplit(",") %>%
+  unlist() %>%
+  grep("0", ., invert=TRUE, value=TRUE) %>%
+  unique() %>%
+  print()
 
-# Write out individual files containing IDs for unrelated individuals
-unrel_path <- here::here("data-raw/1000g_unrel")
-dir.create(unrel_path, showWarnings = FALSE)
-for(pop in unique(ped1kg_unrel$population)) {
-  write_lines(ped1kg_unrel %>% filter(population==pop) %>% pull(id),
-              paste0(file.path(unrel_path, pop), ".txt"))
-}
+# See how many individual IDs are any of the relatives to any other individual ID:
+ped1kg %>% filter(id %in% relatives)
+# # A tibble: 2 x 13
+#   fid     id      dadid momid   sex affected population relationship siblings second_order third_order children other_comments
+#   <chr>   <chr>   <chr> <chr> <dbl>    <dbl> <chr>      <chr>        <chr>    <chr>        <chr>       <chr>    <chr>
[…]

### 2021-03-08 README.Rmd
updated docs #22

--- 2021-02-21 README.Rmd
@@
 ```
+
+## Package data
+
+### Package data objects
+
+Unexported objects `ped1kg` and `ped1kg_unrel` contain [pedigree data obtained from 1000 Genomes FTP](http://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/integrated_call_samples_v3.20200731.ALL.ped) for all 2504 individuals with sequencing data, and all individuals unrelated to anyone else in the data (momid, dadid, siblings, second_order, and third_order all =0).
+This code gets all the unique IDs for anyone listed as a mother, father, sibling, child, second order, or third order relative, _who also has sequencing data_, and removes any _individual ID_ who matches one of these relatives.
+See [data-raw/generate_sysdata.R](data-raw/generate_sysdata.R), specifically the code that begins with the `relatives <-` assignment.
+
+```{r}
+skater:::ped1kg
+skater:::ped1kg_unrel
+```
+
+There are **`r nrow(skater:::ped1kg)-nrow(skater:::ped1kg_unrel)`** samples who have known relatives in the 1000 Genomes ped file (included in `ped1kg` but not in `ped1kg_unrel`):
+
+```{r}
+skater:::ped1kg %>% 
+  dplyr::anti_join(skater:::ped1kg_unrel) %>% 
+  dplyr::select(fid, id, population, second_order:other_comments) %>% 
+  knitr::kable()
+```
+
+### Text files with unrelated founder sample IDs
+
+After creating the `ped1kg_unrel` object in [data-raw/generate_sysdata.R](data-raw/generate_sysdata.R) as described above, the code in [data-raw/write-sampleids-1000g-unrelated.R](data-raw/write-sampleids-1000g-unrelated.R) then writes one sample ID per line for each of those unrelated samples to `inst/extdata`.
+These can be used to pass to `bcftools view --samples-file <POP.txt>`.
+
+The directory containing these files can be found on any system where skater is installed:
+
+```{r}
+system.file("extdata", "sampleids-1000g-unrelated", package="skater", mustWork=TRUE)
+```
+
+The table below shows how many unrelated founders are available in any given population:
+
+```{r}
+system.file("extdata", "sampleids-1000g-unrelated", package="skater", mustWork=TRUE) %>% 
+  list.files(full.names=TRUE) %>% 
+  rlang::set_names(basename) %>% 
+  purrr::map_int(~length(readr::read_lines(.))) %>% 
+  tibble::enframe() %>% 
+  knitr::kable()
+```

### 2021-03-08 README.Rmd
small update to readme

@@
 
-The directory containing these files can be found on any system where skater is installed:
+The directory containing these files can be found on any system where skater is installed.
+On MacOS, this is usually something like `/Library/Frameworks/R.framework/Versions/4.0/Resources/library/skater/extdata/sampleids-1000g-unrelated`.
 
-```{r}
+```{r, eval=FALSE}
 system.file("extdata", "sampleids-1000g-unrelated", package="skater", mustWork=TRUE)

### 2021-03-08 R/confusion_matrix.R
initial cut at modified confusion_matrix fixes #25

@@ -1,0 +1,519 @@
+utils::globalVariables(c('Class', 'Positive', 'N Positive', 'N Negative', 'N'))
+
+#' Calculate Accuracy
+#'
+#' @description Calculates accuracy and related metrics.
+#'
+#' @author Michael Clark (see [m-clark/confusion_matrix](https://github.com/m-clark/confusionMatrix)).
+#'
+#' @param tabble A frequency table created with \code{\link{table}}
+#'
+#' @details Calculates accuracy, lower and upper bounds, the guessing rate and
+#'   p-value of the accuracy vs. the guessing rate. This function is called by
+#'   \code{confusion_matrix}, but if this is all you want, you can simply supply
+#'   the table to this function.
+#'
+#' @return A tibble with the corresponding statistics
+#'
+#' @seealso \code{\link{binom.test}}
+#'
+#' @examples
+#' p = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
+#' o = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
+#' calc_accuracy(table(p, o))
+#'
+#' @importFrom stats binom.test
+#'
+#' @export
+calc_accuracy <- function(tabble) {
+
+  acc <- sum(diag(tabble))/sum(tabble)
+
+  acc_ci <-
+    try(
+      stats::binom.test(sum(diag(tabble)), sum(tabble))$conf.int,
+      silent = TRUE
+    )
+
+  if(inherits(acc_ci, "try-error"))
+    acc_ci <- rep(NA, 2)
+
+  acc_p <- try(
+    stats::binom.test(
+      sum(diag(tabble)),
+      sum(tabble),
+      p = max(colSums(tabble)/sum(tabble)),
+      alternative = "greater"
+    ),
+    silent = TRUE)
+
+  if (inherits(acc_p, "try-error"))
+    acc_p <- c("null.value.probability of success" = NA, p.value = NA)
+  else
+    acc_p <- unlist(acc_p[c("null.value", "p.value")])
+
+  tibble(
+    Accuracy = acc,
+    `Accuracy LL` = acc_ci[1],
+    `Accuracy UL` = acc_ci[2],
+    `Accuracy Guessing` = acc_p[1],
+    `Accuracy P-value` = acc_p[2]
+  )
+}
+
+
+#' Calculate various statistics from a confusion matrix
+#'
+#' @description Given a frequency table of predictions versus target values,
[…]

### 2021-03-08 R/confusion_matrix.R
fix examples

@@ -327,7 +327,8 @@ calc_stats <- function(tabble, prevalence = NULL, positive, ...) {
 #' prediction = c(rep(1, 50), rep(2, 40), rep(3, 60))
 #' target     = c(rep(1, 50), rep(2, 50), rep(3, 50))
 #' confusion_matrix(prediction, target)
-#' confusion_matrix(prediction, target, longer=TRUE) %>% purrr::pluck("Table")
+#' confusion_matrix(prediction, target) %>% purrr::pluck("Table")
+#' confusion_matrix(prediction, target, longer=TRUE)
 #' confusion_matrix(prediction, target, longer=TRUE) %>%
 #'   purrr::pluck("Other") %>%
 #'   tidyr::spread(Class, Value)

### 2021-03-09 README.Rmd
updated docs #25

@@
 
+### Confusion matrix
+
+The skater package adapts functionality from the [m-clark/confusion_matrix](https://github.com/m-clark/confusionMatrix) package to create a confusion matrix and statistics with minimal dependencies.
+
+First, take the `kinpairs` data from above and randomly flip ~20% of the true relationship degrees.
+
+```{r}
+# Function to randomly flip levels of a factor (at 20%, by default)
+randomflip <- function(x, p=.2) ifelse(runif(length(x))<p, sample(unique(x)), x)
+
+# Infer degree (truth/target) using kin2degree, then randomly flip 20% of them
+set.seed(42)
+kinpairs_inferred <- kinpairs %>% 
+  dplyr::mutate(degree_truth=kin2degree(k, max_degree=3)) %>% 
+  dplyr::mutate(degree_truth=tidyr::replace_na(degree_truth, "unrelated")) %>% 
+  dplyr::mutate(degree_inferred=randomflip(degree_truth))
+kinpairs_inferred
+```
+
+The `confusion_matrix` function on its own outputs a list with three objects:
+
+1. Calculates accuracy, lower and upper bounds, the guessing rate and p-value of the accuracy vs. the guessing rate. 
+2. Calculates by class the following statistics:
+  - Sensitivity = A/(A+C)
+  - Specificity = D/(B+D)
+  - Prevalence = (A+C)/(A+B+C+D)
+  - Positive Predictive Value = (sensitivity * prevalence)/((sensitivity*prevalence) + ((1-specificity)*(1-prevalence)))
+  - Negative Predictive Value = (specificity * (1-prevalence))/(((1-sensitivity)*prevalence) + ((specificity)*(1-prevalence)))
+  - Detection Rate = A/(A+B+C+D)
+  - Detection Prevalence = (A+B)/(A+B+C+D)
+  - Balanced Accuracy = (sensitivity+specificity)/2
+  - Precision = A/(A+B)
+  - Recall = A/(A+C)
+  - F1 = harmonic mean of precision and recall
+  - False Discovery Rate = 1 - Positive Predictive Value
+  - False Omission Rate = 1 - Negative Predictive Value
+  - False Positive Rate = 1 - Specificity
+  - False Negative Rate = 1 - Sensitivity
+3. The contingency table object itself.
+
+```{r}
+confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
+                 target = kinpairs_inferred$degree_truth)
+```
+
+Pluck out just the contingency table:
+
+```{r}
+confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
+                 target = kinpairs_inferred$degree_truth) %>% 
+  purrr::pluck("Table")
+```
+
+Output in tidy (`longer=TRUE`) format, then spread stats by class:
+
+```{r}
+confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
+                 target = kinpairs_inferred$degree_truth, 
+                 longer = TRUE) %>% 
+  purrr::pluck("Other") %>% 
+  tidyr::spread(Class, Value) %>% 
+  dplyr::relocate(Average, .after=dplyr::last_col()) %>% 
+  dplyr::mutate_if(rlang::is_double, signif, 2) %>% 
+  knitr::kable()
+```
+
 ## Package data

### 2021-03-09 R/confusion_matrix.R
remove rd links to caret and confusionMatrix pkgs to silence note on CI

@@ -303,8 +303,9 @@ calc_stats <- function(tabble, prevalence = NULL, positive, ...) {
 #'   functions below to find out more. Originally inspired by the
 #'   \code{confusionMatrix} function from the \code{caret} package.
 #'
-#' @seealso \code{\link[caret]{confusionMatrix}}  \code{\link{calc_accuracy}}
-#'   \code{\link{calc_stats}} \code{\link[confusionMatrix]{confusion_matrix}}
+#' @seealso
+#'   \code{\link{calc_accuracy}} \code{\link{calc_stats}}
+#'   \code{\link{confusion_matrix}}
 #'
 #' @return A list of tibble(s) with the associated statistics and possibly the
 #'   frequency table as list column of the first element.

### 2021-03-09 README.Rmd

@@
     - Prevalence = (A+C)/(A+B+C+D)
-  - Positive Predictive Value = (sensitivity * prevalence)/((sensitivity*prevalence) + ((1-specificity)*(1-prevalence)))
-  - Negative Predictive Value = (specificity * (1-prevalence))/(((1-sensitivity)*prevalence) + ((specificity)*(1-prevalence)))
+    - PPV = (sensitivity * prevalence)/((sensitivity * prevalence) + ((1-specificity) * (1-prevalence)))
+    - NPV = (specificity * (1-prevalence))/(((1-sensitivity) * prevalence) + ((specificity) * (1-prevalence)))
     - Detection Rate = A/(A+B+C+D)
@@
     - F1 = harmonic mean of precision and recall
-  - False Discovery Rate = 1 - Positive Predictive Value
-  - False Omission Rate = 1 - Negative Predictive Value
+    - False Discovery Rate = 1 - PPV
+    - False Omission Rate = 1 - NPV
     - False Positive Rate = 1 - Specificity

### 2021-03-15 README.Rmd
fix badge

@@
 <!-- badges: start -->
-[![R-CMD-check](https://github.com/signaturescience/skater/workflows/R-CMD-check/badge.svg)](https://github.com/signaturescience/skater/actions)
+[![R-CMD-check-stable](https://github.com/signaturescience/skater/workflows/R-CMD-check-stable/badge.svg)](https://github.com/signaturescience/skater/actions)
 <!-- badges: end -->

### 2021-03-15 README.Rmd
dev badge

--- 2021-03-09 README.Rmd
@@
 <!-- badges: start -->
-[![R-CMD-check](https://github.com/signaturescience/skater/workflows/R-CMD-check/badge.svg)](https://github.com/signaturescience/skater/actions)
+[![R-CMD-check-stable](https://github.com/signaturescience/skater/workflows/R-CMD-check-stable/badge.svg)](https://github.com/signaturescience/skater/actions)
+
+[![R-CMD-check-dev](https://github.com/signaturescience/skater/workflows/R-CMD-check-dev/badge.svg)](https://github.com/signaturescience/skater/actions)
 <!-- badges: end -->

### 2021-03-16 README.Rmd

@@
 ``` r
-remotes::install_github("signaturescience/skater", auth_token = github_pat(quiet))
+remotes::install_github("signaturescience/skater")
 ```
 
-A GitHub [personal access token](https://github.com/settings/tokens) must be in the environment variable `GITHUB_PAT`.
+A GitHub [personal access token](https://github.com/settings/tokens) must be in the environment variable `GITHUB_PAT` or supplied as a string to `auth_token`. See `?remotes::install_github`.
 

### 2021-03-16 R/utils.R
add kin2cm function fixes #28

@@ -69,3 +69,36 @@ kin2degree <- function(k, max_degree=3L) {
   d <- dibble(max_degree)
   vapply(k, function(k) as.integer(d$degree[which(purrr::map2_lgl(d$l, d$u, ~dplyr::between(k, .x, .y)))]), FUN.VALUE=1L)
 }
+
+
+#' Kinship coefficient to cM
+#'
+#' @description
+#' `r lifecycle::badge("experimental")`
+#'
+#' "Converts" a kinship coefficient to put on the same scale as shared cM using the formula
+#' \eqn{cm <- pmin(3560, 4*pmax(0, k)*3560)}.
+#'
+#'
+#' @param k Kinship coefficient (numeric, typically between 0 and .5, although KING can produce values <0).
+#'
+#' @return Estimated cM, ranging from 0-3560.
+#'
+#' @examples
+#' kin2cm(.25)
+#' kin2cm(.125)
+#' kin2cm(.0625)
+#' dibble(9) %>% dplyr::mutate(cm=kin2cm(k))
+#'
+#' @references <https://dnapainter.com/tools/sharedcmv4>.
+#' @references <https://www.ancestry.com/dna/resource/whitePaper/AncestryDNA-Matching-White-Paper.pdf>.
+#' @references <https://verogen.com/wp-content/uploads/2021/03/snp-typing-uas-kinship-estimation-gedmatch-pro-tech-note-vd2020058-a.pdf>.
+#'
+#'
+#' @export
+kin2cm <- function(k) {
+  stopifnot(all(is.numeric(k)))
+  stopifnot(all(k %>% dplyr::between(-1, 1)))
+  cm <- pmin(3560, 4*pmax(0, k)*3560)
+  return(cm)
+}

### 2021-06-28 paper/paper.Rmd
paper bioinformatics template

---
title_full: "`skater`: An R package for SNP-based Kinship Analysis and Benchmarking"
title_short: "skater package for kinship analysis"

# one of `Original Paper`, `Discovery Note`, `Application Note`, `Review`, `Letter to the Editor`, `Conference Paper`    
subject_section: "Application Note"

access_date:
  - day: DD
    month: MM
    year: YY

author:
  - name: Stephen D. Turner
    affiliation: "1"
  - name: V.P. Nagraj
    affiliation: "1"
  - name: Matthew Scholz
    affiliation: "1"
  - name: Shakeel Jessa
    affiliation: "1"
  - name: Carlos Acevedo
    affiliation: "1"
  - name: Jianye Ge
    affiliation: "2"
  - name: August Woener
    affiliation: "2"
  - name: Bruce Budowle
    affiliation: "2"

address:
  - code: 1
    address: Signature Science, LLC., Charlottesville, Virginia, 22911, USA.
  - code: 2
    address: University of North Texas Health Science Center, Fort Worth, Texas, 76107, USA.

# Full email address to be given, preferably an institution email address.  
corresponding_author_text: "To whom correspondence should be addressed. E-mail: sturner\\@signaturescience.com"
corresponding_author_email: "sturner\\@signaturescience.com"

motivation: |
  This section should specifically state the scientific question within the context of the field of study.

results: |
  This section should summarize the scientific advance or novel results of the study, and its impact on computational biology.

availability: |
  This section should state software availability if the paper focuses mainly on software development or on the implementation of an algorithm. Examples are: 'Freely available on the web at XXX.' Website implemented in Perl, MySQL and Apache, with all major browsers supported'; or 'Source code and binaries freely available for download at URL, implemented in C++ and supported on linux and MS Windows'. The complete address (URL) should be given. If the manuscript describes new software tools or the implementation of novel algorithms the software must be freely available to non-commercial users. Authors must also ensure that the software is available for a full TWO YEARS following publication. The editors of Bioinformatics encourage authors to make their source code available and, if possible, to provide access through an open source license.

supplementary_info: |
  Supplementary data are available at Bioinformatics Online.

output:
  rticles::bioinformatics_article:
    citation_package: natbib

bibliography: bibliography.bib
---

# Introduction

Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

# The skater package

Describe the approach. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Describe the approach. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Describe the approach. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Describe the approach. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

# Functions and usage

Detailed methods. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Detailed methods. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Detailed methods. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

## Pedigree processing

Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

## IBD segment analysis

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

## Additional features and built-in data

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

# Conclusion

Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.

# Acknowledgements {.unnumbered}

These should be included at the end of the text and not in footnotes. Please ensure you acknowledge all sources of funding, see funding section below.

Details of all funding sources for the work in question should be given in a separate section entitled 'Funding'. This should appear before the 'Acknowledgements' section.

# Funding {.unnumbered}

The following rules should be followed:

-   The sentence should begin: 'This work was supported by ...' -
-   The full official funding agency name should be given, i.e. 'National Institutes of Health', not 'NIH' (full RIN-approved list of UK funding agencies)
-   Grant numbers should be given in brackets as follows: '[grant number xxxx]'
-   Multiple grant numbers should be separated by a comma as follows: '[grant numbers xxxx, yyyy]'
-   Agencies should be separated by a semi-colon (plus 'and' before the last funding agency)
-   Where individuals need to be specified for certain sources of funding the following text should be added after the relevant agency or grant number 'to [author initials]'.

An example is given here: 'This work was supported by the National Institutes of Health [AA123456 to C.S., BB765432 to M.H.]; and the Alcohol & Education Research Council [hfygr667789].'

Oxford Journals will deposit all NIH-funded articles in PubMed Central. See Depositing articles in repositories -- information for authors for details. Authors must ensure that manuscripts are clearly indicated as NIH-funded using the guidelines above.

### 2021-06-28 paper/paper.tex
paper bioinformatics template

\documentclass{bioinfo}
\copyrightyear{2016} \pubyear{2016}

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Pandoc citation processing

% hyperref makes the margins screwy.
% https://groups.google.com/forum/#!topic/latexusersgroup/4W_SwGk6zx4
% http://ansuz.sooke.bc.ca/software/latex-tricks.php
% \usepackage[colorlinks=true, allcolors=blue]{hyperref}

\access{Advance Access Publication Date:   }
\appnotes{Manuscript Category}

\begin{document}
\firstpage{1}

\subtitle{Application Note}

\title[skater package for kinship analysis]{\texttt{skater}: An R
package for SNP-based Kinship Analysis and Benchmarking}

\author[FirstAuthorLastName \textit{et~al}.]{
Stephen D. Turner\,\textsuperscript{1},
V.P. Nagraj\,\textsuperscript{1},
Matthew Scholz\,\textsuperscript{1},
Shakeel Jessa\,\textsuperscript{1},
Carlos Acevedo\,\textsuperscript{1},
Jianye Ge\,\textsuperscript{2},
August Woener\,\textsuperscript{2},
Bruce Budowle\,\textsuperscript{2},
}

\address{
\textsuperscript{1}Signature Science, LLC., Charlottesville, Virginia,
22911, USA.\\
\textsuperscript{2}University of North Texas Health Science Center, Fort
Worth, Texas, 76107, USA.\\
}

\corresp{To whom correspondence should be addressed. E-mail:
sturner@signaturescience.com}

\history{Received on XXX; revised on XXX; accepted on XXX}

\editor{Associate Editor: XXX}

\abstract{
\textbf{Motivation:} This section should specifically state the
scientific question within the context of the field of study.\\
\textbf{Results:} This section should summarize the scientific advance
or novel results of the study, and its impact on computational
biology.\\
\textbf{Availability:} This section should state software availability
if the paper focuses mainly on software development or on the
implementation of an algorithm. Examples are: `Freely available on the
web at XXX.' Website implemented in Perl, MySQL and Apache, with all
major browsers supported'; or `Source code and binaries freely available
for download at URL, implemented in C++ and supported on linux and MS
Windows'. The complete address (URL) should be given. If the manuscript
describes new software tools or the implementation of novel algorithms
the software must be freely available to non-commercial users. Authors
must also ensure that the software is available for a full TWO YEARS
following publication. The editors of Bioinformatics encourage authors
to make their source code available and, if possible, to provide access
through an open source license.\\
\textbf{Contact:}sturner@signaturescience.com\\
\textbf{Supplementary information:} Supplementary data are available at
Bioinformatics Online.}

\maketitle

\section{Introduction}

Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

\section{The skater package}

Describe the approach. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Describe the approach. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Describe the approach. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Describe the approach. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

\section{Functions and usage}

Detailed methods. Lorem ipsum ad nauseum. Introduce your topic. Lorem
ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Detailed methods. Lorem ipsum ad nauseum. Introduce your topic. Lorem
ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Detailed methods. Lorem ipsum ad nauseum. Introduce your topic. Lorem
ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

\subsection{Pedigree processing}

Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

\subsection{IBD segment analysis}

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

\subsection{Additional features and built-in data}

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
Lorem ipsum ad nauseum.

\section{Conclusion}

Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum
ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your
topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad
nauseum.

Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum
ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your
topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad
nauseum.

Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum
ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your
topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad
nauseum.

Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum
ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your
topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad
nauseum.

\section*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

These should be included at the end of the text and not in footnotes.
Please ensure you acknowledge all sources of funding, see funding
section below.

Details of all funding sources for the work in question should be given
in a separate section entitled `Funding'. This should appear before the
`Acknowledgements' section.

\section*{Funding}
\addcontentsline{toc}{section}{Funding}

The following rules should be followed:

\begin{itemize}
\tightlist
\item
  The sentence should begin: `This work was supported by \ldots{}' -
\item
  The full official funding agency name should be given, i.e.~`National
  Institutes of Health', not `NIH' (full RIN-approved list of UK funding
  agencies)
\item
  Grant numbers should be given in brackets as follows: `{[}grant number
  xxxx{]}'
\item
  Multiple grant numbers should be separated by a comma as follows:
  `{[}grant numbers xxxx, yyyy{]}'
\item
  Agencies should be separated by a semi-colon (plus `and' before the
  last funding agency)
\item
  Where individuals need to be specified for certain sources of funding
  the following text should be added after the relevant agency or grant
  number `to {[}author initials{]}'.
\end{itemize}

An example is given here: `This work was supported by the National
Institutes of Health {[}AA123456 to C.S., BB765432 to M.H.{]}; and the
Alcohol \& Education Research Council {[}hfygr667789{]}.'

Oxford Journals will deposit all NIH-funded articles in PubMed Central.
See Depositing articles in repositories -- information for authors for
details. Authors must ensure that manuscripts are clearly indicated as
NIH-funded using the guidelines above.

% Bibliography
\bibliographystyle{natbib}
\bibliography{bibliography.bib}

\end{document}

### 2021-06-30 README.Rmd
update readme
updated readme to reference vignette

@@
 
-**S**NP-based **K**inship **A**nalysis **T**esting and **E**valuation: miscellaneous **R** utilties.
+**S**NP-based **K**inship **A**nalysis **T**esting and **E**valuation: miscellaneous **R** data analysis utilties.
 
@@
 ``` r
-remotes::install_github("signaturescience/skater")
+remotes::install_github("signaturescience/skater", build_vignettes=TRUE)
 ```
@@
 
-## Examples
+## Usage
 
-```{r}
-library(skater)
+The "Basic Usage" vignette steps through the primary functionality of the `skater` package:
+
+``` r
+vignette("basic_usage", package = "skater")
 ```
-
-### Pedigrees and PLINK .fam files
-
-Read a PLINK-formatted .fam file:
-
-```{r}
-famfile <- system.file("extdata", "3gens.fam", package="skater", mustWork=TRUE)
-fam <- read_fam(famfile)
-fam
-```
-
-Convert each family into a pedigree object (see the [kinship2 vignette](https://cran.r-project.org/web/packages/kinship2/vignettes/pedigree.html)):
-
-```{r}
-peds <- fam2ped(fam)
-peds
-```
-
-In the example above, the resulting tibble is nested by family ID.
-The `data` column contains the individual family information, while the `ped` column contains the pedigree object for that family.
-You can unnest any particular family:
-
-```{r}
-peds %>% 
-  dplyr::filter(fid=="testped1") %>% 
-  tidyr::unnest(cols=data)
-```
-
-You can also look at a single pedigree:
-
-```{r}
-peds$ped[[1]]
-```
-
-Or plot that pedigree:
-
-```{r plotped}
-plot(peds$ped[[1]], mar=c(1,4,1,4))
-```
-
-The `plot_pedigree` function in the skater package will walk over a list of pedigree objects, writing a multi-page PDF to file, with each page containing a pedigree from each of the families:
-
-```{r, eval=FALSE}
-plot_pedigree(peds$ped, file="3gens.ped.pdf")
-```
-
-The `ped2kinpair()` function takes a pedigree object and produces a pairwise list of relationships with their expected kinship coefficient.
-
-Run on a single family:
-
-```{r}
-ped2kinpair(peds$ped[[1]])
-```
-
-Map over all families:
-
-```{r}
-kinpairs <- 
-  peds %>% 
-  dplyr::mutate(pairs=purrr::map(ped, ped2kinpair)) %>% 
-  dplyr::select(fid, pairs) %>% 
-  tidyr::unnest(cols=pairs)
-kinpairs
-```
-
-Note that this maps `ped2kinpair` over all `ped` objects in the input tibble, and that relationships are not shown for between-family relationships (which should all be zero).
-
-### Degree inference
-
-The `dibble()` function creates a **d**egree **i**nference tibble, with degrees up to the specified `max_degree` (default=3), expected kinship coefficient, and lower (`l`) and upper (`u`) inference ranges (see [Manichaikul 2010](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3025716/)).
-Degree 0 corresponds to self / identity / MZ twins, with an expected kinship coefficient of 0.5, with inference range >=0.354.
-Anything beyond the maximum degree resolution is considered unrelated (degree `NA`), with expected kinship coefficient of 0.
-
-```{r}
-dibble()
-```
-
-The degree inference `max_degree` default is 3. Change this argument to allow more granular degree inference ranges:
-
-```{r}
-dibble(max_degree = 5)
-```
-
-Note that the distance between relationship degrees becomes smaller as the relationship degree becomes more distant.
-The `dibble()` function will throw a warning with `max_degree` >=10, and will stop with an error at >=12.
-
-The `kin2degree()` function infers the relationship degree given a kinship coefficient and a `max_degree` up to which anything more distant is treated as unrelated. Example first degree relative:
-
-```{r}
-kin2degree(.25, max_degree=3)
-```
-
-Example 4th degree relative, but using the default max_degree resolution of 3:
-
-```{r}
-kin2degree(.0312, max_degree=3)
-```
-
-Example 4th degree relative, but increasing the degree resolution:
-
-```{r}
-kin2degree(.0312, max_degree=5)
-```
-
-The `kin2degree()` function is vectorized over values of `k`, so can be used inside of a `mutate` on a tibble of kinship coefficients:
-
-```{r}
-# Get two pairs from each type of relationship we have in kinpairs:
-kinpairs_subset <- 
-  kinpairs %>% 
-  dplyr::group_by(k) %>% 
-  dplyr::slice(1:2)
-kinpairs_subset
-
-# Infer degree out to third degree relatives:
-kinpairs_subset %>% 
-  dplyr::mutate(degree=kin2degree(k, max_degree=3))
-```
-
-### Confusion matrix
-
-The skater package adapts functionality from the [m-clark/confusion_matrix](https://github.com/m-clark/confusionMatrix) package to create a confusion matrix and statistics with minimal dependencies.
-
-First, take the `kinpairs` data from above and randomly flip ~20% of the true relationship degrees.
-
-```{r}
-# Function to randomly flip levels of a factor (at 20%, by default)
-randomflip <- function(x, p=.2) ifelse(runif(length(x))<p, sample(unique(x)), x)
-
-# Infer degree (truth/target) using kin2degree, then randomly flip 20% of them
-set.seed(42)
-kinpairs_inferred <- kinpairs %>% 
-  dplyr::mutate(degree_truth=kin2degree(k, max_degree=3)) %>% 
-  dplyr::mutate(degree_truth=tidyr::replace_na(degree_truth, "unrelated")) %>% 
-  dplyr::mutate(degree_inferred=randomflip(degree_truth))
-kinpairs_inferred
-```
-
-The `confusion_matrix` function on its own outputs a list with three objects:
-
-1. Calculates accuracy, lower and upper bounds, the guessing rate and p-value of the accuracy vs. the guessing rate. 
-2. Calculates by class the following statistics:
-    - Sensitivity = A/(A+C)
-    - Specificity = D/(B+D)
-    - Prevalence = (A+C)/(A+B+C+D)
-    - PPV = (sensitivity * prevalence)/((sensitivity * prevalence) + ((1-specificity) * (1-prevalence)))
-    - NPV = (specificity * (1-prevalence))/(((1-sensitivity) * prevalence) + ((specificity) * (1-prevalence)))
-    - Detection Rate = A/(A+B+C+D)
-    - Detection Prevalence = (A+B)/(A+B+C+D)
-    - Balanced Accuracy = (sensitivity+specificity)/2
-    - Precision = A/(A+B)
-    - Recall = A/(A+C)
-    - F1 = harmonic mean of precision and recall
-    - False Discovery Rate = 1 - PPV
-    - False Omission Rate = 1 - NPV
-    - False Positive Rate = 1 - Specificity
-    - False Negative Rate = 1 - Sensitivity
-3. The contingency table object itself.
-
-```{r}
-confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
-                 target = kinpairs_inferred$degree_truth)
-```
-
-Pluck out just the contingency table:
-
-```{r}
-confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
-                 target = kinpairs_inferred$degree_truth) %>% 
-  purrr::pluck("Table")
-```
-
-Output in tidy (`longer=TRUE`) format, then spread stats by class:
-
-```{r}
-confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
-                 target = kinpairs_inferred$degree_truth, 
-                 longer = TRUE) %>% 
-  purrr::pluck("Other") %>% 
-  tidyr::spread(Class, Value) %>% 
-  dplyr::relocate(Average, .after=dplyr::last_col()) %>% 
-  dplyr::mutate_if(rlang::is_double, signif, 2) %>% 
-  knitr::kable()
-```
-
-## Package data
-
-### Package data objects
-
-Unexported objects `ped1kg` and `ped1kg_unrel` contain [pedigree data obtained from 1000 Genomes FTP](http://ftp.1000genomes.ebi.ac.uk/vol1/ftp/release/20130502/integrated_call_samples_v3.20200731.ALL.ped) for all 2504 individuals with sequencing data, and all individuals unrelated to anyone else in the data (momid, dadid, siblings, second_order, and third_order all =0).
-This code gets all the unique IDs for anyone listed as a mother, father, sibling, child, second order, or third order relative, _who also has sequencing data_, and removes any _individual ID_ who matches one of these relatives.
-See [data-raw/generate_sysdata.R](data-raw/generate_sysdata.R), specifically the code that begins with the `relatives <-` assignment.
-
-```{r}
-skater:::ped1kg
-skater:::ped1kg_unrel
-```
-
-There are **`r nrow(skater:::ped1kg)-nrow(skater:::ped1kg_unrel)`** samples who have known relatives in the 1000 Genomes ped file (included in `ped1kg` but not in `ped1kg_unrel`):
-
-```{r}
-skater:::ped1kg %>% 
-  dplyr::anti_join(skater:::ped1kg_unrel) %>% 
-  dplyr::select(fid, id, population, second_order:other_comments) %>% 
-  knitr::kable()
-```
-
-### Text files with unrelated founder sample IDs
-
-After creating the `ped1kg_unrel` object in [data-raw/generate_sysdata.R](data-raw/generate_sysdata.R) as described above, the code in [data-raw/write-sampleids-1000g-unrelated.R](data-raw/write-sampleids-1000g-unrelated.R) then writes one sample ID per line for each of those unrelated samples to `inst/extdata`.
-These can be used to pass to `bcftools view --samples-file <POP.txt>`.
-
-The directory containing these files can be found on any system where skater is installed.
-On MacOS, this is usually something like `/Library/Frameworks/R.framework/Versions/4.0/Resources/library/skater/extdata/sampleids-1000g-unrelated`.
-
-```{r, eval=FALSE}
-system.file("extdata", "sampleids-1000g-unrelated", package="skater", mustWork=TRUE)
-```
-
-The table below shows how many unrelated founders are available in any given population:
-
-```{r}
-system.file("extdata", "sampleids-1000g-unrelated", package="skater", mustWork=TRUE) %>% 
-  list.files(full.names=TRUE) %>% 
-  rlang::set_names(basename) %>% 
-  purrr::map_int(~length(readr::read_lines(.))) %>% 
-  tibble::enframe() %>% 
-  knitr::kable()
-```

### 2021-06-30 paper/paper.Rmd
introduction

@@
 ---
-title_full: "`skater`: An R package for SNP-based Kinship Analysis and Benchmarking"
+# title_full: "skater: An R package for SNP-based Kinship Analysis and Benchmarking"
+title_full: "skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
 title_short: "skater package for kinship analysis"
 
-# one of `Original Paper`, `Discovery Note`, `Application Note`, `Review`, `Letter to the Editor`, `Conference Paper`    
-subject_section: "Application Note"
+# One of the following below. See https://academic.oup.com/bioinformatics/pages/instructions_for_authors#Types%20of%20Manuscript
+# one of `Original Paper`, `Discovery Note`, `Applications Note`, `Review`, `Letter to the Editor`, `Conference Paper`    
+manuscript_type: Applications Note
+
+# One of the following below. See https://academic.oup.com/bioinformatics/pages/instructions_for_authors#Detailed%20Scope%20Guidelines
+# Genome analysis
+# Sequence analysis
+# Phylogenetics
+# Structural bioinformatics
+# Gene expression
+# Genetic and population analysis
+# Systems biology
+# Data and text mining
+# Databases and ontologies
+# Bioimage informatics
+subject_section:  Genetics and Population Analysis
 
@@
 
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+Inferring familial relationships between individuals using genetic data is a common problem in population genetics, medical genetics, and forensics.
+There are multiple approaches to estimating the relatedness between samples, including genome-wide relatedness measures such as those implemented in Plink [@purcell2007] or KING [@manichaikul2010], and methods that rely on identity by descent (IBD) segment detection such as GERMLINE [@gusev2009], hap-IBD [@zhou2020], and IBIS [@seidman2020].
+Recent efforts have focused on benchmarking these methods [@ramstetter2017; @vries2021] aided by tools for simulating pedigrees and genome-wide SNP data [@caballero2019].
+Analyzing results from genome-wide SNP-based kinship analysis or comparing analyses to simulated data for benchmarking have to this point required writing one-off analysis functions or utility scripts that are rarely shared, poorly documented, and at best, have an inconsistent design and API across multiple developers and programming languages.
+Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation.
 

### 2021-06-30 R/ped.R
removing experimental tag for fam2ped

@@ -1,7 +1,6 @@
 #' Fam to pedigree
 #'
 #' @description
-#' `r lifecycle::badge("experimental")`
 #'
 #' Converts a [PLINK-formatted fam file](https://www.cog-genomics.org/plink/1.9/formats#fam) to a pedigree object using [kinship2::pedigree].
 #'

### 2021-06-30 data-raw/write-sampleids-1000g-unrelated.R
removed unused data from pkg

@@ -1,10 +1,13 @@
 library(tidyverse)
-unrel_path <- here::here("inst/extdata/sampleids-1000g-unrelated")
-dir.create(unrel_path, showWarnings = FALSE)
-ped1kg_unrel <- skater:::ped1kg_unrel
-for(pop in unique(ped1kg_unrel$population)) {
-  unrelated_ids <- ped1kg_unrel %>% filter(population==pop) %>% pull(id)
-  unrelated_ids_path <- paste0(file.path(unrel_path, pop), ".txt")
-  message(glue::glue("Writing {length(unrelated_ids)} to path: {unrelated_ids_path}"))
-  write_lines(unrelated_ids, unrelated_ids_path)
-}
+
+## NOTE: commented out because we don't need to save this data for now ... uncomment if needed
+
+# unrel_path <- here::here("inst/extdata/sampleids-1000g-unrelated")
+# dir.create(unrel_path, showWarnings = FALSE)
+# ped1kg_unrel <- skater:::ped1kg_unrel
+# for(pop in unique(ped1kg_unrel$population)) {
+#   unrelated_ids <- ped1kg_unrel %>% filter(population==pop) %>% pull(id)
+#   unrelated_ids_path <- paste0(file.path(unrel_path, pop), ".txt")
+#   message(glue::glue("Writing {length(unrelated_ids)} to path: {unrelated_ids_path}"))
+#   write_lines(unrelated_ids, unrelated_ids_path)
+# }

### 2021-07-12 paper/paper.Rmd
finish first draft of paper

@@
 motivation: |
-  This section should specifically state the scientific question within the context of the field of study.
+  SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produce results that often require further downstream manipulation and analysis.
+A software package with a consistent design and implementation that offers this downstream analysis functionality is needed.
 
 results: |
-  This section should summarize the scientific advance or novel results of the study, and its impact on computational biology.
+  Here we present the skater package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation as an **R** package.
+The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
 
 availability: |
-  This section should state software availability if the paper focuses mainly on software development or on the implementation of an algorithm.
-Examples are: 'Freely available on the web at XXX.' Website implemented in Perl, MySQL and Apache, with all major browsers supported'; or 'Source code and binaries freely available for download at URL, implemented in C++ and supported on linux and MS Windows'.
-The complete address (URL) should be given.
-If the manuscript describes new software tools or the implementation of novel algorithms the software must be freely available to non-commercial users.
-Authors must also ensure that the software is available for a full TWO YEARS following publication.
-The editors of Bioinformatics encourage authors to make their source code available and, if possible, to provide access through an open source license.
+  The skater package is implemented as an R package and is released under the MIT license at https\:\/\/github.com/signaturescience/skater.
+Documentation is available at https\:\/\/signaturescience.github.io/skater.
 
@@
 Recent efforts have focused on benchmarking these methods [@ramstetter2017; @vries2021] aided by tools for simulating pedigrees and genome-wide SNP data [@caballero2019].
-Analyzing results from genome-wide SNP-based kinship analysis or comparing analyses to simulated data for benchmarking have to this point required writing one-off analysis functions or utility scripts that are rarely shared, poorly documented, and at best, have an inconsistent design and API across multiple developers and programming languages.
+Analyzing results from genome-wide SNP-based kinship analysis or comparing analyses to simulated data for benchmarking have to this point required writing one-off analysis functions or utility scripts that are rarely shared and/or poorly documented.
+There is a need in the field for a well-documented software package with a consistent design and API that contains functions to assist with downstream manipulation, benchmarking, and analysis of SNP-based kinship assessment methods.
 Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation.
@@
 
-Describe the approach.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+The skater package provides a tidyverse-friendly collection of analysis and utility functions for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation as an **R** package.
+Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
 
-Describe the approach.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+## Pedigree parsing, manipulation, and analysis
 
-Describe the approach.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+The skater package has several functions for importing, parsing, and analyzing pedigree data.
+Pedigrees define familial relationships in a hierarchical structure.
+Many genomics tools for working with pedigrees start with a .fam file, which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
+The skater package contains the functions `read_fam()` to read in a PLINK-formatted .fam file, and another function `fam2ped()` to convert these into an internal pedigree object as a nested tibble with one row per family.
+Further functions such as `plot_pedigree()` produce a multi-page PDF drawing a diagram of the pedigree for each family, while `ped2kinpair()` produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair.
 
-Describe the approach.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+## Relationship degree inference and benchmarking
 
-# Functions and usage
+The skater package includes functions to translate kinship coefficients to relationship degrees. The kinship coefficients could come from `ped2kinpair()` or other kinship estimation software.
 
-Detailed methods.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+The `dibble()` function creates a **d**egree **i**nference **t**ibble, with degrees up to the specified maximum degree resolution, expected kinship coefficient, and lower and upper inference ranges as defined in @manichaikul2010.
+The `kin2degree()` function infers the relationship degree given a kinship coefficient and a maximum degree resolution (e.g., 7th-degree relatives) up to which anything more distant is treated as unrelated.
 
-Detailed methods.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Detailed methods.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-## Pedigree processing
-
-Details for Method 1.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 1.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 1.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 1.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to truth.
+When aggregated over many relationships and inferences, this can help benchmark performance of a particular kinship analysis method.
+The skater package adapts a `confusion_matrix()` function from @clark2021 to provide standard contingency table metrics (e.g.
+sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
+The R-RMSE metric is defined more thoroughly in the skater package vignette (see Supplementary Material), and is a superior measure to classification accuracy when benchmarking relationship degree estimation.
+Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree).
 
@@
 
-Details for Method 2.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-## Additional features and built-in data
-
-Details for Method 2.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+Tools such as hap-IBD [@zhou2020], and IBIS [@seidman2020] detect shared IBD segments between individuals.
+The skater package includes functionality to take those IBD segments, compute shared genomic centimorgan (cM) length, and convert that shared cM to a kinship coefficient.
+In addition to inferred segments, these functions can estimate "truth" kinship from simulated IBD segments [@caballero2019].
+The `read_ibd()` function reads pairwise IBD segments from IBD inference tools and from simulated IBD segments.
+The `read_map()` function reads in genetic map in a standard format which is required to translate the total centimorgans shared IBD to a kinship coefficient using the `ibd2kin()` function.
 
@@
 
-Anything else?
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+The skater R package provides a thoroughly-documented software package for common data import, manipulation, and analysis tasks typically encountered when working with common SNP-based kinship analysis tools.
+All package functions are internally documented with examples, and the package contains a vignette demonstrating usage, inputs, outputs, and interpretation of all key functions (see Supplementary Material).
+The package contains internal tests which are automatically run with continuous integration via GitHub Actions whenever the package code is updated.
+The skater package is permissively licensed (MIT) and is easily extensible to accommodate outputs from new genome-wide relatedness and IBD segment methods as they become available.
 
-Anything else?
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+<!-- # Acknowledgements {.unnumbered} -->
 
-Anything else?
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
+<!-- These should be included at the end of the text and not in footnotes. Please ensure you acknowledge all sources of funding, see funding section below. -->
 
-Anything else?
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-Introduce your topic.
-Lorem ipsum ad nauseum.
-
-# Acknowledgements {.unnumbered}
-
-These should be included at the end of the text and not in footnotes. Please ensure you acknowledge all sources of funding, see funding section below.
-
-Details of all funding sources for the work in question should be given in a separate section entitled 'Funding'. This should appear before the 'Acknowledgements' section.
+<!-- Details of all funding sources for the work in question should be given in a separate section entitled 'Funding'. This should appear before the 'Acknowledgements' section. -->
 
@@
 
-The following rules should be followed:
+**UNT FOLKS: NEED YOUR INPUT HERE!** The following rules should be followed:
 

### 2021-07-12 paper/paper.Rmd
copy vignette to paper supplementary materials #50 #48

@@
 
+```{r, echo=FALSE, eval=FALSE}
+## Run this chunk interactively if you update the package vignette.
+## This renders the vignette as a PDF with a new title.
+
+# Where is the package vignette?
+vpath <- here::here("vignettes/basic_usage.Rmd")
+# Where do you want to save the new modified vignette rmarkdown?
+spath <- here::here("paper/supplement-vignette.Rmd")
+# Read in the vignette rmarkdown text
+vtext <- readLines(vpath)
+# Create the supplemental material rmarkdown by replacing the title YAML
+stext <- sub(pattern='title: "Basic Usage"', 
+             replacement = 'title: |\n    | Supplementary Material for:\n    | skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation\n    | _(skater v0.1.0 package vignette)_', 
+             x=vtext)
+# Write out that modified yaml to the supplementary material rmarkdown
+writeLines(text=stext, con=spath)
+# Render it to PDF
+rmarkdown::render(spath, output_format="pdf_document")
+```
+
 # Introduction

### 2021-07-12 paper/paper.tex
compile paper and vignette

@@
 \access{Advance Access Publication Date:   }
-\appnotes{Manuscript Category}
+\appnotes{Applications Note}
 
@@
 
-\subtitle{Application Note}
+\subtitle{Genetics and Population Analysis}
 
-\title[skater package for kinship analysis]{\texttt{skater}: An R
-package for SNP-based Kinship Analysis and Benchmarking}
+\title[skater package for kinship analysis]{skater: An R package for
+SNP-based Kinship Analysis, Testing, and Evaluation}
 
 \author[FirstAuthorLastName \textit{et~al}.]{
-Stephen D. Turner\,\textsuperscript{1},
-V.P. Nagraj\,\textsuperscript{1},
-Matthew Scholz\,\textsuperscript{1},
-Shakeel Jessa\,\textsuperscript{1},
-Carlos Acevedo\,\textsuperscript{1},
-Jianye Ge\,\textsuperscript{2},
-August Woener\,\textsuperscript{2},
-Bruce Budowle\,\textsuperscript{2},
+Stephen D. Turner\,\textsuperscript{1}, V.P.
+Nagraj\,\textsuperscript{1}, Matthew
+Scholz\,\textsuperscript{1}, Shakeel Jessa\,\textsuperscript{1}, Carlos
+Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2}, August
+Woener\,\textsuperscript{2}, Bruce Budowle\,\textsuperscript{2}
 }
@@
 \abstract{
-\textbf{Motivation:} This section should specifically state the
-scientific question within the context of the field of study.\\
-\textbf{Results:} This section should summarize the scientific advance
-or novel results of the study, and its impact on computational
-biology.\\
-\textbf{Availability:} This section should state software availability
-if the paper focuses mainly on software development or on the
-implementation of an algorithm. Examples are: `Freely available on the
-web at XXX.' Website implemented in Perl, MySQL and Apache, with all
-major browsers supported'; or `Source code and binaries freely available
-for download at URL, implemented in C++ and supported on linux and MS
-Windows'. The complete address (URL) should be given. If the manuscript
-describes new software tools or the implementation of novel algorithms
-the software must be freely available to non-commercial users. Authors
-must also ensure that the software is available for a full TWO YEARS
-following publication. The editors of Bioinformatics encourage authors
-to make their source code available and, if possible, to provide access
-through an open source license.\\
+\textbf{Motivation:} SNP-based kinship analysis with genome-wide
+relationship estimation and IBD segment analysis methods produce results
+that often require further downstream manipulation and analysis. A
+software package with a consistent design and implementation that offers
+this downstream analysis functionality is needed.\\
+\textbf{Results:} Here we present the skater package for
+\textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting,
+and \textbf{e}valuation as an \textbf{R} package. The skater package
+contains a suite of well-documented tools for importing, parsing, and
+analyzing pedigree data, performing relationship degree inference,
+benchmarking relationship degree classification, and summarizing IBD
+segment data.\\
+\textbf{Availability:} The skater package is implemented as an R package
+and is released under the MIT license at
+https://github.com/signaturescience/skater. Documentation is available
+at https://signaturescience.github.io/skater.\\
 \textbf{Contact:}sturner@signaturescience.com\\
@@
 
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
+Inferring familial relationships between individuals using genetic data
+is a common problem in population genetics, medical genetics, and
+forensics. There are multiple approaches to estimating the relatedness
+between samples, including genome-wide relatedness measures such as
+those implemented in Plink \citep{purcell2007} or KING
+\citep{manichaikul2010}, and methods that rely on identity by descent
+(IBD) segment detection such as GERMLINE \citep{gusev2009}, hap-IBD
+\citep{zhou2020}, and IBIS \citep{seidman2020}. Recent efforts have
+focused on benchmarking these methods \citep{ramstetter2017, vries2021}
+aided by tools for simulating pedigrees and genome-wide SNP data
+\citep{caballero2019}. Analyzing results from genome-wide SNP-based
+kinship analysis or comparing analyses to simulated data for
+benchmarking have to this point required writing one-off analysis
+functions or utility scripts that are rarely shared and/or poorly
+documented. There is a need in the field for a well-documented software
+package with a consistent design and API that contains functions to
+assist with downstream manipulation, benchmarking, and analysis of
+SNP-based kinship assessment methods. Here we present the skater R
+package for \textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis,
+\textbf{t}esting, and \textbf{e}valuation.
 
@@
 
-Describe the approach. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
+The skater package provides a tidyverse-friendly collection of analysis
+and utility functions for \textbf{S}NP-based \textbf{k}inship
+\textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation as an
+\textbf{R} package. Functions in the package include tools for
+importing, parsing, and analyzing pedigree data, performing relationship
+degree inference, benchmarking relationship degree classification, and
+summarizing IBD segment data.
 
-Describe the approach. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
+\subsection{Pedigree parsing, manipulation, and analysis}
 
-Describe the approach. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
+The skater package has several functions for importing, parsing, and
+analyzing pedigree data. Pedigrees define familial relationships in a
+hierarchical structure. Many genomics tools for working with pedigrees
+start with a .fam file, which is a tabular format with one row per
+individual and columns for unique IDs of the mother, father, and the
+family unit. The skater package contains the functions
+\texttt{read\_fam()} to read in a PLINK-formatted .fam file, and another
+function \texttt{fam2ped()} to convert these into an internal pedigree
+object as a nested tibble with one row per family. Further functions
+such as \texttt{plot\_pedigree()} produce a multi-page PDF drawing a
+diagram of the pedigree for each family, while \texttt{ped2kinpair()}
+produces a pairwise list of relationships between all individuals in the
+data with the expected kinship coefficients for each pair.
 
-Describe the approach. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
+\subsection{Relationship degree inference and benchmarking}
 
-\section{Functions and usage}
+The skater package includes functions to translate kinship coefficients
+to relationship degrees. The kinship coefficients could come from
+\texttt{ped2kinpair()} or other kinship estimation software.
 
-Detailed methods. Lorem ipsum ad nauseum. Introduce your topic. Lorem
-ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
+The \texttt{dibble()} function creates a \textbf{d}egree
+\textbf{i}nference \textbf{t}ibble, with degrees up to the specified
+maximum degree resolution, expected kinship coefficient, and lower and
+upper inference ranges as defined in \citet{manichaikul2010}. The
+\texttt{kin2degree()} function infers the relationship degree given a
+kinship coefficient and a maximum degree resolution (e.g., 7th-degree
+relatives) up to which anything more distant is treated as unrelated.
 
-Detailed methods. Lorem ipsum ad nauseum. Introduce your topic. Lorem
-ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Detailed methods. Lorem ipsum ad nauseum. Introduce your topic. Lorem
-ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-\subsection{Pedigree processing}
-
-Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 1. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
+Once estimated kinship is converted to degree, it may be of interest to
+compare the inferred degree to truth. When aggregated over many
+relationships and inferences, this can help benchmark performance of a
+particular kinship analysis method. The skater package adapts a
+\texttt{confusion\_matrix()} function from \citet{clark2021} to provide
+standard contingency table metrics (e.g.~sensitivity, specificity, PPV,
+precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
+The R-RMSE metric is defined more thoroughly in the skater package
+vignette (see Supplementary Material), and is a superior measure to
+classification accuracy when benchmarking relationship degree
+estimation. Taking the reciprocal of the target and predicted degree in
+a typical RMSE calculation results in larger penalties for more
+egregious misclassifications (e.g., classifying a first-degree relative
+pair as second degree) than misclassifications at more distant
+relationships (e.g., misclassifying a fourth-degree relative pair as
+fifth-degree).
 
@@
 
-Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-\subsection{Additional features and built-in data}
-
-Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
-
-Details for Method 2. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad nauseum.
-Introduce your topic. Lorem ipsum ad nauseum. Introduce your topic.
-Lorem ipsum ad nauseum.
+Tools such as hap-IBD \citep{zhou2020}, and IBIS \citep{seidman2020}
+detect shared IBD segments between individuals. The skater package
+includes functionality to take those IBD segments, compute shared
+genomic centimorgan (cM) length, and convert that shared cM to a kinship
+coefficient. In addition to inferred segments, these functions can
+estimate ``truth'' kinship from simulated IBD segments
+\citep{caballero2019}. The \texttt{read\_ibd()} function reads pairwise
+IBD segments from IBD inference tools and from simulated IBD segments.
+The \texttt{read\_map()} function reads in genetic map in a standard
+format which is required to translate the total centimorgans shared IBD
+to a kinship coefficient using the \texttt{ibd2kin()} function.
 
@@
 
-Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum
-ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your
-topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad
-nauseum.
-
-Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum
-ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your
-topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad
-nauseum.
-
-Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum
-ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your
-topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad
-nauseum.
-
-Anything else? Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum
-ad nauseum. Introduce your topic. Lorem ipsum ad nauseum. Introduce your
-topic. Lorem ipsum ad nauseum. Introduce your topic. Lorem ipsum ad
-nauseum.
-
-\section*{Acknowledgements}
-\addcontentsline{toc}{section}{Acknowledgements}
-
-These should be included at the end of the text and not in footnotes.
-Please ensure you acknowledge all sources of funding, see funding
-section below.
-
-Details of all funding sources for the work in question should be given
-in a separate section entitled `Funding'. This should appear before the
-`Acknowledgements' section.
+The skater R package provides a thoroughly-documented software package
+for common data import, manipulation, and analysis tasks typically
+encountered when working with common SNP-based kinship analysis tools.
+All package functions are internally documented with examples, and the
+package contains a vignette demonstrating usage, inputs, outputs, and
+interpretation of all key functions (see Supplementary Material). The
+package contains internal tests which are automatically run with
+continuous integration via GitHub Actions whenever the package code is
+updated. The skater package is permissively licensed (MIT) and is easily
+extensible to accommodate outputs from new genome-wide relatedness and
+IBD segment methods as they become available.
 
@@
 
-The following rules should be followed:
+\textbf{UNT FOLKS: NEED YOUR INPUT HERE!} The following rules should be
+followed:
 

### 2021-07-12 paper/supplement-vignette.Rmd
compile paper and vignette

---
title: |
    | Supplementary Material for:
    | skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation
    | _(skater v0.1.0 package vignette)_
output: 
  rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{Basic Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

## Overview

The `skater` package provides a collection of analysis and utility functions for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation as an **R** package. Functions in the package include tools for working with pedigree data, performing relationship degree inference, assessing classification accuracy, and summarizing IBD segment data.

```{r}
library(skater)
```

## Pedigree parsing and manipulation

Pedigrees define familial relationships in a hierarchical structure. Many genomics tools for working with pedigrees start with a "fam" file, which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.

One of the formats used by PLINK and other genetic analysis tools is the `.fam` file.[^plink-fam] The package includes `read_fam()` to read files in this format:

[^plink-fam]: https://www.cog-genomics.org/plink/1.9/formats#fam

```{r}
famfile <- system.file("extdata", "3gens.fam", package="skater", mustWork=TRUE)
fam <- read_fam(famfile)
fam
```

Family structures in the "fam" format can then be translated to the `pedigree` structure used by the `kinship2` package.[^kinship2-vignette] The "fam" format could include multiple families, and the `fam2ped()` function will collapse them all into a `tibble` with one row per family:

[^kinship2-vignette]: https://cran.r-project.org/web/packages/kinship2/vignettes/pedigree.html

```{r}
peds <- fam2ped(fam)
peds
```

In the example above, the resulting `tibble` is nested by family ID. The `data` column contains the individual family information, while the `ped` column contains the pedigree object for that family. You can unnest any particular family:

```{r}
peds %>% 
  dplyr::filter(fid=="testped1") %>% 
  tidyr::unnest(cols=data)
```

You can also look at a single pedigree:

```{r}
peds$ped[[1]]
```

Or plot that pedigree:

```{r plotped, fig.width=8, fig.height=8, fig.align="center"}
plot(peds$ped[[1]], mar=c(1,4,1,4))
```

The `plot_pedigree()` function from `skater` will walk over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from each of the families:

```{r, eval=FALSE}
plot_pedigree(peds$ped, file="3gens.ped.pdf")
```

The `ped2kinpair()` function takes a pedigree object and produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair.

The function can be run on a single family:

```{r}
ped2kinpair(peds$ped[[1]])
```

Or mapped over all families in the pedigree

```{r}
kinpairs <- 
  peds %>% 
  dplyr::mutate(pairs=purrr::map(ped, ped2kinpair)) %>% 
  dplyr::select(fid, pairs) %>% 
  tidyr::unnest(cols=pairs)
kinpairs
```

Note that this maps `ped2kinpair()` over all `ped` objects in the input `tibble`, and that relationships are not shown for between-family relationships (which should all be zero).

## Degree Inference

The `skater` package includes functions to translate kinship coefficients to relationship degrees. The kinship coefficients could come from `ped2kinpair()` or other kinship estimation software.

The `dibble()` function creates a **d**egree **i**nference `tibble`, with degrees up to the specified `max_degree` (default=3), expected kinship coefficient, and lower (`l`) and upper (`u`) inference ranges as defined in the KING paper.[^manichaikul] Degree 0 corresponds to self / identity / monozygotic twins, with an expected kinship coefficient of 0.5, with inference range >=0.354. Anything beyond the maximum degree resolution is considered unrelated (degree `NA`), with expected kinship coefficient of 0.

[^manichaikul]: Manichaikul, A., Mychaleckyj, J. C., Rich, S. S., Daly, K., Sale, M., & Chen, W. M. (2010). Robust relationship inference in genome-wide association studies. Bioinformatics (Oxford, England), 26(22), 2867–2873. https://doi.org/10.1093/bioinformatics/btq559

```{r}
dibble()
```

The degree inference `max_degree` default is 3. Change this argument to allow more granular degree inference ranges:

```{r}
dibble(max_degree = 5)
```

Note that the distance between relationship degrees becomes smaller as the relationship degree becomes more distant. The `dibble()` function will throw a warning with `max_degree` >=10, and will stop with an error at >=12.

The `kin2degree()` function infers the relationship degree given a kinship coefficient and a `max_degree` up to which anything more distant is treated as unrelated. Example first degree relative:

```{r}
kin2degree(.25, max_degree=3)
```

Example 4th degree relative, but using the default max_degree resolution of 3:

```{r}
kin2degree(.0312, max_degree=3)
```

Example 4th degree relative, but increasing the degree resolution:

```{r}
kin2degree(.0312, max_degree=5)
```

The `kin2degree()` function is vectorized over values of `k`, so it can be used inside of a `mutate` on a `tibble` of kinship coefficients:

```{r}
# Get two pairs from each type of relationship we have in kinpairs:
kinpairs_subset <- 
  kinpairs %>% 
  dplyr::group_by(k) %>% 
  dplyr::slice(1:2)
kinpairs_subset

# Infer degree out to third degree relatives:
kinpairs_subset %>% 
  dplyr::mutate(degree=kin2degree(k, max_degree=3))
```

## Benchmarking Degree Classification

Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to truth. When aggregated over many relationships and inferences, this method can help benchmark performance of a particular kinship analysis method.

The `skater` package adapts functionality from the `confusionMatrix` package[^confusionMatrix] in the `confusion_matrix()` function.

[^confusionMatrix]: https://github.com/m-clark/confusionMatrix

The `confusion_matrix()` function on its own outputs a list with three objects:

1. A `tibble` with calculated accuracy, lower and upper bounds, the guessing rate and p-value of the accuracy vs. the guessing rate. 
2. A `tibble` with the following statistics (for each class):
    - Sensitivity = A/(A+C)
    - Specificity = D/(B+D)
    - Prevalence = (A+C)/(A+B+C+D)
    - PPV = (sensitivity * prevalence)/((sensitivity * prevalence) + ((1-specificity) * (1-prevalence)))
    - NPV = (specificity * (1-prevalence))/(((1-sensitivity) * prevalence) + ((specificity) * (1-prevalence)))
    - Detection Rate = A/(A+B+C+D)
    - Detection Prevalence = (A+B)/(A+B+C+D)
    - Balanced Accuracy = (sensitivity+specificity)/2
    - Precision = A/(A+B)
    - Recall = A/(A+C)
    - F1 = harmonic mean of precision and recall
    - False Discovery Rate = 1 - PPV
    - False Omission Rate = 1 - NPV
    - False Positive Rate = 1 - Specificity
    - False Negative Rate = 1 - Sensitivity
3. A `matrix` with the contingency table object itself.
4. A `vector` with the reciprocal RMSE (R-RMSE). The R-RMSE is calculated as `sqrt(mean((1/(Target+.5)-1/(Predicted+.5))^2)))`, and is a superior measure to classification accuracy when benchmarking relationship degree estimation. Taking the reciprocal of the target and predicted degree results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree). The +0.5 adjustment prevents division-by-zero when a 0th-degree (identical) relative pair is introduced.

To illustrate the usage, first take the `kinpairs` data from above and randomly flip ~20% of the true relationship degrees.

```{r}
# Function to randomly flip levels of a factor (at 20%, by default)
randomflip <- function(x, p=.2) ifelse(runif(length(x))<p, sample(unique(x)), x)

# Infer degree (truth/target) using kin2degree, then randomly flip 20% of them
set.seed(42)
kinpairs_inferred <- kinpairs %>% 
  dplyr::mutate(degree_truth=kin2degree(k, max_degree=3)) %>% 
  dplyr::mutate(degree_truth=tidyr::replace_na(degree_truth, "unrelated")) %>% 
  dplyr::mutate(degree_inferred=randomflip(degree_truth))
kinpairs_inferred
```

```{r}
confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
                 target = kinpairs_inferred$degree_truth)
```

You can use `purrr::pluck()` to isolate just the contingency table:

```{r}
confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
                 target = kinpairs_inferred$degree_truth) %>% 
  purrr::pluck("Table")
```

Or optionally output in a tidy (`longer=TRUE`) format, then spread stats by class:

```{r}
confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
                 target = kinpairs_inferred$degree_truth, 
                 longer = TRUE) %>% 
  purrr::pluck("Other") %>% 
  tidyr::spread(Class, Value) %>% 
  dplyr::relocate(Average, .after=dplyr::last_col()) %>% 
  dplyr::mutate_if(rlang::is_double, signif, 2) %>% 
  knitr::kable()
```

## IBD Segment Analysis

Tools such as `hap-ibd`[^hap-ibd] are capable of inferring shared IBD segments between individuals. The `skater` package includes functionality to take those IBD segments, compute shared genomic centimorgan (cM) length, and convert that shared cM to a kinship coefficient. In addition to inferred segments, these functions can estimate "truth" kinship from data simulated by `ped-sim`.[^ped-sim]

[^hap-ibd]: https://github.com/browning-lab/hap-ibd#output-files
[^ped-sim]: https://github.com/williamslab/ped-sim#output-ibd-segments-file

The `read_ibd()` function reads in the pairwise IBD segment format. Input to this function can either be inferred IBD segments from hap-IBD (`source="hapibd"`) or simulated segments (`source="pedsim"`). The first example below uses data in the `hap-ibd` output format:

```{r}
hapibd_fp <- system.file("extdata", "GBR.sim.ibd.gz", package="skater", mustWork=TRUE)
hapibd_seg <- read_ibd(hapibd_fp, source = "hapibd")
hapibd_seg
```

In order to translate the shared genomic cM length to a kinship coefficient, you must load a genetic map with `read_map()`. Software for IBD segment inference and simulation requires a genetic map. The map loaded for kinship estimation should be the same one used for creating the shared IBD segment output. The example below uses a minimal genetic map created with `min_map`[^min_map] that ships with `skater`:

[^min_map]: https://github.com/williamslab/min_map

```{r}
gmapfile <- system.file("extdata", "sexspec-avg-min.plink.map", package="skater", mustWork=TRUE)
gmap <- read_map(gmapfile)
gmap
```

The `ibd2kin()` function takes the segments and map file and outputs a `tibble` with one row per pair of individuals and columns for individual 1 ID, individual 2 ID, and the kinship coefficient for the pair:

```{r}
ibd_dat <- ibd2kin(.ibd_data=hapibd_seg, .map=gmap)
ibd_dat
```

As noted above, the IBD segment kinship estimation can be performed on simulated segments. The package includes an example of IBD data in that format:

```{r}
pedsim_fp <- system.file("extdata", "GBR.sim.seg.gz", package="skater", mustWork=TRUE)
pedsim_seg <- read_ibd(pedsim_fp, source = "pedsim")
pedsim_seg
```

Notably, `ped-sim` differentiates IBD1 and IBD2 segments. Given that IBD1 and IBD2 segments are weighted differently in kinship calculation, this should be accounted for in processing. In the example below the shared IBD is calculated separately for IBD1 and IBD2 with `type="IBD1"` and `type="IBD2"` respectively. You can then combine those results and sum the IBD1 and IBD2 kinship coefficients to get the overall kinship coefficient:

```{r}
ibd1_dat <- ibd2kin(.ibd_data=pedsim_seg$IBD1, .map=gmap, type="IBD1")
ibd2_dat <- ibd2kin(.ibd_data=pedsim_seg$IBD2, .map=gmap, type="IBD2")
dplyr::bind_rows(ibd1_dat,ibd2_dat) %>%
  dplyr::group_by(id1,id2) %>%
  dplyr::summarise(kinship = sum(kinship), .groups = "drop")
```

### 2021-07-13 paper/paper.Rmd
vpn edits to skater paper

@@
 motivation: |
-  SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produce results that often require further downstream manipulation and analysis.
-A software package with a consistent design and implementation that offers this downstream analysis functionality is needed.
+  SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation.
+A dedicated software package that consistently and intuitively implements this analysis functionality is needed.
 
 results: |
-  Here we present the skater package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation as an **R** package.
+  Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**.
 The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
@@
 Inferring familial relationships between individuals using genetic data is a common problem in population genetics, medical genetics, and forensics.
-There are multiple approaches to estimating the relatedness between samples, including genome-wide relatedness measures such as those implemented in Plink [@purcell2007] or KING [@manichaikul2010], and methods that rely on identity by descent (IBD) segment detection such as GERMLINE [@gusev2009], hap-IBD [@zhou2020], and IBIS [@seidman2020].
-Recent efforts have focused on benchmarking these methods [@ramstetter2017; @vries2021] aided by tools for simulating pedigrees and genome-wide SNP data [@caballero2019].
-Analyzing results from genome-wide SNP-based kinship analysis or comparing analyses to simulated data for benchmarking have to this point required writing one-off analysis functions or utility scripts that are rarely shared and/or poorly documented.
+There are multiple approaches to estimating relatedness between samples, including genome-wide measures such as those implemented in Plink [@purcell2007] or KING [@manichaikul2010], and methods that rely on identity by descent (IBD) segment detection such as GERMLINE [@gusev2009], hap-IBD [@zhou2020], and IBIS [@seidman2020].
+Recent efforts focusing on benchmarking these methods [@ramstetter2017; @vries2021] have been aided by tools for simulating pedigrees and genome-wide SNP data [@caballero2019].
+Analyzing results from genome-wide SNP-based kinship analysis or comparing analyses to simulated data for benchmarking have to this point required writing one-off analysis functions or utility scripts that are seldom distributed with robust documentation, test suites, or narrative examples of usage.
 There is a need in the field for a well-documented software package with a consistent design and API that contains functions to assist with downstream manipulation, benchmarking, and analysis of SNP-based kinship assessment methods.
-Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation.
+Here we present the skater package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**.
 
@@
 
-The skater package provides a tidyverse-friendly collection of analysis and utility functions for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation as an **R** package.
+The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis.
 Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
+The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem [@Wickham2019].
 
@@
 Many genomics tools for working with pedigrees start with a .fam file, which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
-The skater package contains the functions `read_fam()` to read in a PLINK-formatted .fam file, and another function `fam2ped()` to convert these into an internal pedigree object as a nested tibble with one row per family.
+The skater package contains the function `read_fam()` to read in a PLINK-formatted .fam file, and another function `fam2ped()` to convert the content into a pedigree object as a nested tibble with one row per family.
+All pedigree processing from skater internally leverages a data structure from the kinship2 package [@kinship2].
 Further functions such as `plot_pedigree()` produce a multi-page PDF drawing a diagram of the pedigree for each family, while `ped2kinpair()` produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair.
@@
 sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
-The R-RMSE metric is defined more thoroughly in the skater package vignette (see Supplementary Material), and is a superior measure to classification accuracy when benchmarking relationship degree estimation.
-Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree).
+The R-RMSE metric is defined more thoroughly in the skater package vignette (see Supplementary Material), and may be a preferrable measure of classification accuracy when benchmarking relationship degree estimation.
+Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).
 
@@
 
-The skater R package provides a thoroughly-documented software package for common data import, manipulation, and analysis tasks typically encountered when working with common SNP-based kinship analysis tools.
+The skater R package provides a robust software package for data import, manipulation, and analysis tasks typically encountered when working with SNP-based kinship analysis tools.
 All package functions are internally documented with examples, and the package contains a vignette demonstrating usage, inputs, outputs, and interpretation of all key functions (see Supplementary Material).
-The package contains internal tests which are automatically run with continuous integration via GitHub Actions whenever the package code is updated.
+The package contains internal tests that are automatically run with continuous integration via GitHub Actions whenever the package code is updated.
 The skater package is permissively licensed (MIT) and is easily extensible to accommodate outputs from new genome-wide relatedness and IBD segment methods as they become available.

### 2021-07-13 paper/paper.tex
vpn edits to skater paper

@@
 \access{Advance Access Publication Date:   }
-\appnotes{Applications Note}
+\appnotes{Manuscript Category}
 
@@
 
-\title[skater package for kinship analysis]{skater: An R package for
-SNP-based Kinship Analysis, Testing, and Evaluation}
+\title[skater package for kinship analysis]{skater: An R package for SNP-based Kinship Analysis, Testing, and
+Evaluation}
 
 \author[FirstAuthorLastName \textit{et~al}.]{
-Stephen D. Turner\,\textsuperscript{1}, V.P.
-Nagraj\,\textsuperscript{1}, Matthew
-Scholz\,\textsuperscript{1}, Shakeel Jessa\,\textsuperscript{1}, Carlos
-Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2}, August
-Woener\,\textsuperscript{2}, Bruce Budowle\,\textsuperscript{2}
+Stephen D. Turner\,\textsuperscript{1},
+V.P. Nagraj\,\textsuperscript{1},
+Matthew Scholz\,\textsuperscript{1},
+Shakeel Jessa\,\textsuperscript{1},
+Carlos Acevedo\,\textsuperscript{1},
+Jianye Ge\,\textsuperscript{2},
+August Woener\,\textsuperscript{2},
+Bruce Budowle\,\textsuperscript{2},
 }
@@
 \address{
-\textsuperscript{1}Signature Science, LLC., Charlottesville, Virginia,
-22911, USA.\\
-\textsuperscript{2}University of North Texas Health Science Center, Fort
-Worth, Texas, 76107, USA.\\
+\textsuperscript{1}Signature Science, LLC., Charlottesville, Virginia, 22911, USA.\\
+\textsuperscript{2}University of North Texas Health Science Center, Fort Worth, Texas,
+76107, USA.\\
 }
@@
 \abstract{
-\textbf{Motivation:} SNP-based kinship analysis with genome-wide
-relationship estimation and IBD segment analysis methods produce results
-that often require further downstream manipulation and analysis. A
-software package with a consistent design and implementation that offers
-this downstream analysis functionality is needed.\\
-\textbf{Results:} Here we present the skater package for
-\textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting,
-and \textbf{e}valuation as an \textbf{R} package. The skater package
-contains a suite of well-documented tools for importing, parsing, and
-analyzing pedigree data, performing relationship degree inference,
-benchmarking relationship degree classification, and summarizing IBD
-segment data.\\
-\textbf{Availability:} The skater package is implemented as an R package
-and is released under the MIT license at
-https://github.com/signaturescience/skater. Documentation is available
-at https://signaturescience.github.io/skater.\\
+\textbf{Motivation:} SNP-based kinship analysis with genome-wide relationship estimation and
+IBD segment analysis methods produces results that often require further
+downstream processing and manipulation. A dedicated software package
+that consistently and intuitively implements this analysis functionality
+is needed.\\
+\textbf{Results:} Here we present the skater R package for \textbf{S}NP-based
+\textbf{k}inship \textbf{a}nalysis, \textbf{t}esting, and
+\textbf{e}valuation with \textbf{R}. The skater package contains a suite
+of well-documented tools for importing, parsing, and analyzing pedigree
+data, performing relationship degree inference, benchmarking
+relationship degree classification, and summarizing IBD segment data.\\
+\textbf{Availability:} The skater package is implemented as an R package and is released under
+the MIT license at https://github.com/signaturescience/skater.
+Documentation is available at https://signaturescience.github.io/skater.\\
 \textbf{Contact:}sturner@signaturescience.com\\
-\textbf{Supplementary information:} Supplementary data are available at
-Bioinformatics Online.}
+\textbf{Supplementary information:} Supplementary data are available at Bioinformatics Online.}
 
@@
 is a common problem in population genetics, medical genetics, and
-forensics. There are multiple approaches to estimating the relatedness
-between samples, including genome-wide relatedness measures such as
-those implemented in Plink \citep{purcell2007} or KING
+forensics. There are multiple approaches to estimating relatedness
+between samples, including genome-wide measures such as those
+implemented in Plink \citep{purcell2007} or KING
 \citep{manichaikul2010}, and methods that rely on identity by descent
 (IBD) segment detection such as GERMLINE \citep{gusev2009}, hap-IBD
-\citep{zhou2020}, and IBIS \citep{seidman2020}. Recent efforts have
-focused on benchmarking these methods \citep{ramstetter2017, vries2021}
-aided by tools for simulating pedigrees and genome-wide SNP data
+\citep{zhou2020}, and IBIS \citep{seidman2020}. Recent efforts focusing
+on benchmarking these methods \citep{ramstetter2017, vries2021} have
+been aided by tools for simulating pedigrees and genome-wide SNP data
 \citep{caballero2019}. Analyzing results from genome-wide SNP-based
@@
 benchmarking have to this point required writing one-off analysis
-functions or utility scripts that are rarely shared and/or poorly
-documented. There is a need in the field for a well-documented software
-package with a consistent design and API that contains functions to
-assist with downstream manipulation, benchmarking, and analysis of
-SNP-based kinship assessment methods. Here we present the skater R
-package for \textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis,
-\textbf{t}esting, and \textbf{e}valuation.
+functions or utility scripts that are seldom distributed with robust
+documentation, test suites, or narrative examples of usage. There is a
+need in the field for a well-documented software package with a
+consistent design and API that contains functions to assist with
+downstream manipulation, benchmarking, and analysis of SNP-based kinship
+assessment methods. Here we present the skater package for
+\textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting,
+and \textbf{e}valuation with \textbf{R}.
 
@@
 
-The skater package provides a tidyverse-friendly collection of analysis
-and utility functions for \textbf{S}NP-based \textbf{k}inship
-\textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation as an
-\textbf{R} package. Functions in the package include tools for
-importing, parsing, and analyzing pedigree data, performing relationship
-degree inference, benchmarking relationship degree classification, and
-summarizing IBD segment data.
+The skater package provides an intuitive collection of analysis and
+utility functions for SNP-based kinship analysis. Functions in the
+package include tools for importing, parsing, and analyzing pedigree
+data, performing relationship degree inference, benchmarking
+relationship degree classification, and summarizing IBD segment data.
+The package is designed to adhere to ``tidy'' data analysis principles,
+and builds upon the tools released under the tidyverse R ecosystem
+\citep{Wickham2019}.
 
@@
 individual and columns for unique IDs of the mother, father, and the
-family unit. The skater package contains the functions
+family unit. The skater package contains the function
 \texttt{read\_fam()} to read in a PLINK-formatted .fam file, and another
-function \texttt{fam2ped()} to convert these into an internal pedigree
-object as a nested tibble with one row per family. Further functions
-such as \texttt{plot\_pedigree()} produce a multi-page PDF drawing a
-diagram of the pedigree for each family, while \texttt{ped2kinpair()}
-produces a pairwise list of relationships between all individuals in the
-data with the expected kinship coefficients for each pair.
+function \texttt{fam2ped()} to convert the content into a pedigree
+object as a nested tibble with one row per family. All pedigree
+processing from skater internally leverages a data structure from the
+kinship2 package \citep{kinship2}. Further functions such as
+\texttt{plot\_pedigree()} produce a multi-page PDF drawing a diagram of
+the pedigree for each family, while \texttt{ped2kinpair()} produces a
+pairwise list of relationships between all individuals in the data with
+the expected kinship coefficients for each pair.
 
@@
 The R-RMSE metric is defined more thoroughly in the skater package
-vignette (see Supplementary Material), and is a superior measure to
-classification accuracy when benchmarking relationship degree
+vignette (see Supplementary Material), and may be a preferrable measure
+of classification accuracy when benchmarking relationship degree
 estimation. Taking the reciprocal of the target and predicted degree in
@@
 pair as second degree) than misclassifications at more distant
-relationships (e.g., misclassifying a fourth-degree relative pair as
+relationships (e.g., classifying a fourth-degree relative pair as
 fifth-degree).
@@
 
-The skater R package provides a thoroughly-documented software package
-for common data import, manipulation, and analysis tasks typically
-encountered when working with common SNP-based kinship analysis tools.
-All package functions are internally documented with examples, and the
-package contains a vignette demonstrating usage, inputs, outputs, and
-interpretation of all key functions (see Supplementary Material). The
-package contains internal tests which are automatically run with
-continuous integration via GitHub Actions whenever the package code is
-updated. The skater package is permissively licensed (MIT) and is easily
-extensible to accommodate outputs from new genome-wide relatedness and
-IBD segment methods as they become available.
+The skater R package provides a robust software package for data import,
+manipulation, and analysis tasks typically encountered when working with
+SNP-based kinship analysis tools. All package functions are internally
+documented with examples, and the package contains a vignette
+demonstrating usage, inputs, outputs, and interpretation of all key
+functions (see Supplementary Material). The package contains internal
+tests that are automatically run with continuous integration via GitHub
+Actions whenever the package code is updated. The skater package is
+permissively licensed (MIT) and is easily extensible to accommodate
+outputs from new genome-wide relatedness and IBD segment methods as they
+become available.
 

### 2021-07-13 paper/paper.tex
recompiled

@@
 
-\title[skater package for kinship analysis]{skater: An R package for SNP-based Kinship Analysis, Testing, and
-Evaluation}
+\title[skater package for kinship analysis]{skater: An R package for
+SNP-based Kinship Analysis, Testing, and Evaluation}
 
 \author[FirstAuthorLastName \textit{et~al}.]{
-Stephen D. Turner\,\textsuperscript{1},
-V.P. Nagraj\,\textsuperscript{1},
-Matthew Scholz\,\textsuperscript{1},
-Shakeel Jessa\,\textsuperscript{1},
-Carlos Acevedo\,\textsuperscript{1},
-Jianye Ge\,\textsuperscript{2},
-August Woener\,\textsuperscript{2},
-Bruce Budowle\,\textsuperscript{2},
+Stephen D. Turner\,\textsuperscript{1}, V.P.
+Nagraj\,\textsuperscript{1}, Matthew
+Scholz\,\textsuperscript{1}, Shakeel Jessa\,\textsuperscript{1}, Carlos
+Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2}, August
+Woener\,\textsuperscript{2}, Bruce Budowle\,\textsuperscript{2}
 }
@@
 \address{
-\textsuperscript{1}Signature Science, LLC., Charlottesville, Virginia, 22911, USA.\\
-\textsuperscript{2}University of North Texas Health Science Center, Fort Worth, Texas,
-76107, USA.\\
+\textsuperscript{1}Signature Science, LLC., Charlottesville, Virginia,
+22911, USA.\\
+\textsuperscript{2}University of North Texas Health Science Center, Fort
+Worth, Texas, 76107, USA.\\
 }
@@
 \abstract{
-\textbf{Motivation:} SNP-based kinship analysis with genome-wide relationship estimation and
-IBD segment analysis methods produces results that often require further
-downstream processing and manipulation. A dedicated software package
-that consistently and intuitively implements this analysis functionality
-is needed.\\
-\textbf{Results:} Here we present the skater R package for \textbf{S}NP-based
-\textbf{k}inship \textbf{a}nalysis, \textbf{t}esting, and
-\textbf{e}valuation with \textbf{R}. The skater package contains a suite
-of well-documented tools for importing, parsing, and analyzing pedigree
-data, performing relationship degree inference, benchmarking
+\textbf{Motivation:} SNP-based kinship analysis with genome-wide
+relationship estimation and IBD segment analysis methods produces
+results that often require further downstream processing and
+manipulation. A dedicated software package that consistently and
+intuitively implements this analysis functionality is needed.\\
+\textbf{Results:} Here we present the skater R package for
+\textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting,
+and \textbf{e}valuation with \textbf{R}. The skater package contains a
+suite of well-documented tools for importing, parsing, and analyzing
+pedigree data, performing relationship degree inference, benchmarking
 relationship degree classification, and summarizing IBD segment data.\\
-\textbf{Availability:} The skater package is implemented as an R package and is released under
-the MIT license at https://github.com/signaturescience/skater.
-Documentation is available at https://signaturescience.github.io/skater.\\
+\textbf{Availability:} The skater package is implemented as an R package
+and is released under the MIT license at
+https://github.com/signaturescience/skater. Documentation is available
+at https://signaturescience.github.io/skater.\\
 \textbf{Contact:}sturner@signaturescience.com\\
-\textbf{Supplementary information:} Supplementary data are available at Bioinformatics Online.}
+\textbf{Supplementary information:} Supplementary data are available at
+Bioinformatics Online.}
 

### 2021-07-13 paper/paper.Rmd
Update paper.Rmd
Minor edits.

@@
 
-Inferring familial relationships between individuals using genetic data is a common problem in population genetics, medical genetics, and forensics.
+Inferring familial relationships between individuals using genetic data is a common practice in population genetics, medical genetics, and forensics.
 There are multiple approaches to estimating relatedness between samples, including genome-wide measures such as those implemented in Plink [@purcell2007] or KING [@manichaikul2010], and methods that rely on identity by descent (IBD) segment detection such as GERMLINE [@gusev2009], hap-IBD [@zhou2020], and IBIS [@seidman2020].
@@
 
-The `dibble()` function creates a **d**egree **i**nference **t**ibble, with degrees up to the specified maximum degree resolution, expected kinship coefficient, and lower and upper inference ranges as defined in @manichaikul2010.
-The `kin2degree()` function infers the relationship degree given a kinship coefficient and a maximum degree resolution (e.g., 7th-degree relatives) up to which anything more distant is treated as unrelated.
+The `dibble()` function creates a **d**egree **i**nference t**ibble**, with degrees up to the specified maximum degree resolution, expected kinship coefficient, and lower and upper inference ranges as defined in @manichaikul2010.
+The `kin2degree()` function infers the relationship degree given a kinship coefficient and a maximum degree resolution (e.g., 7th-degree relatives) up to which anything more distant is classified as unrelated.
 
-Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to truth.
+Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to known degrees of relatedness.
 When aggregated over many relationships and inferences, this can help benchmark performance of a particular kinship analysis method.

### 2021-07-13 paper/supplement-vignette.Rmd
Update supplement-vignette.Rmd
Minor edits for flow.

@@
 Pedigrees define familial relationships in a hierarchical structure.
-Many genomics tools for working with pedigrees start with a "fam" file, which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
 
-One of the formats used by PLINK and other genetic analysis tools is the `.fam` file.[^plink-fam] The package includes `read_fam()` to read files in this format:
+One of the formats used by PLINK and other genetic analysis tools is the `.fam` file,[^plink-fam] which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
+The package includes `read_fam()` to read files in this format:
 
@@
 
-Family structures in the "fam" format can then be translated to the `pedigree` structure used by the `kinship2` package.[^kinship2-vignette] The "fam" format could include multiple families, and the `fam2ped()` function will collapse them all into a `tibble` with one row per family:
+Family structures imported from "fam" formated files can then be translated to the `pedigree` structure used by the `kinship2` package.[^kinship2-vignette] The "fam" format may include multiple families, and the `fam2ped()` function will collapse them all into a `tibble` with one row per family:
 
@@
 
-The `plot_pedigree()` function from `skater` will walk over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from each of the families:
+The `plot_pedigree()` function from `skater` will iterate over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from each family:
 

### 2021-07-14 paper/paper.tex
update supplement and recompile

@@
 Inferring familial relationships between individuals using genetic data
-is a common problem in population genetics, medical genetics, and
+is a common practice in population genetics, medical genetics, and
 forensics. There are multiple approaches to estimating relatedness
@@
 The \texttt{dibble()} function creates a \textbf{d}egree
-\textbf{i}nference \textbf{t}ibble, with degrees up to the specified
+\textbf{i}nference t\textbf{ibble}, with degrees up to the specified
 maximum degree resolution, expected kinship coefficient, and lower and
@@
 kinship coefficient and a maximum degree resolution (e.g., 7th-degree
-relatives) up to which anything more distant is treated as unrelated.
+relatives) up to which anything more distant is classified as unrelated.
 
 Once estimated kinship is converted to degree, it may be of interest to
-compare the inferred degree to truth. When aggregated over many
-relationships and inferences, this can help benchmark performance of a
-particular kinship analysis method. The skater package adapts a
-\texttt{confusion\_matrix()} function from \citet{clark2021} to provide
-standard contingency table metrics (e.g.~sensitivity, specificity, PPV,
-precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
-The R-RMSE metric is defined more thoroughly in the skater package
-vignette (see Supplementary Material), and may be a preferrable measure
-of classification accuracy when benchmarking relationship degree
-estimation. Taking the reciprocal of the target and predicted degree in
-a typical RMSE calculation results in larger penalties for more
-egregious misclassifications (e.g., classifying a first-degree relative
-pair as second degree) than misclassifications at more distant
-relationships (e.g., classifying a fourth-degree relative pair as
-fifth-degree).
+compare the inferred degree to known degrees of relatedness. When
+aggregated over many relationships and inferences, this can help
+benchmark performance of a particular kinship analysis method. The
+skater package adapts a \texttt{confusion\_matrix()} function from
+\citet{clark2021} to provide standard contingency table metrics
+(e.g.~sensitivity, specificity, PPV, precision, recall, F1, etc.) with a
+new reciprocal RMSE (R-RMSE) metric. The R-RMSE metric is defined more
+thoroughly in the skater package vignette (see Supplementary Material),
+and may be a preferrable measure of classification accuracy when
+benchmarking relationship degree estimation. Taking the reciprocal of
+the target and predicted degree in a typical RMSE calculation results in
+larger penalties for more egregious misclassifications (e.g.,
+classifying a first-degree relative pair as second degree) than
+misclassifications at more distant relationships (e.g., classifying a
+fourth-degree relative pair as fifth-degree).
 

### 2021-07-15 paper/paper.Rmd
affiliations

@@
   - name: Jianye Ge
-    affiliation: "2"
+    affiliation: "2,3"
   - name: August Woener
-    affiliation: "2"
+    affiliation: "2,3"
   - name: Bruce Budowle
-    affiliation: "2"
+    affiliation: "2,3"
 
@@
   - code: 2
-    address: University of North Texas Health Science Center, Fort Worth, Texas, 76107, USA.
+    address: Center for Human Identification, University of North Texas Health Science Center, 3500 Camp Bowie Blvd., Fort Worth, TX, 76107, USA
+  - code: 3
+    address: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, 3500 Camp Bowie Blvd., Fort Worth, TX, 76107, USA
 

### 2021-07-15 paper/paper.Rmd
typos, funding statement

@@
 Inferring familial relationships between individuals using genetic data is a common practice in population genetics, medical genetics, and forensics.
-There are multiple approaches to estimating relatedness between samples, including genome-wide measures such as those implemented in Plink [@purcell2007] or KING [@manichaikul2010], and methods that rely on identity by descent (IBD) segment detection such as GERMLINE [@gusev2009], hap-IBD [@zhou2020], and IBIS [@seidman2020].
+There are multiple approaches to estimating relatedness between samples, including genome-wide measures, such as those implemented in Plink [@purcell2007] or KING [@manichaikul2010], and methods that rely on identity by descent (IBD) segment detection, such as GERMLINE [@gusev2009], hap-IBD [@zhou2020], and IBIS [@seidman2020].
 Recent efforts focusing on benchmarking these methods [@ramstetter2017; @vries2021] have been aided by tools for simulating pedigrees and genome-wide SNP data [@caballero2019].
@@
 Many genomics tools for working with pedigrees start with a .fam file, which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
-The skater package contains the function `read_fam()` to read in a PLINK-formatted .fam file, and another function `fam2ped()` to convert the content into a pedigree object as a nested tibble with one row per family.
+The skater package contains the function `read_fam()` to read in a PLINK-formatted .fam file and another function `fam2ped()` to convert the content into a pedigree object as a nested tibble with one row per family.
 All pedigree processing from skater internally leverages a data structure from the kinship2 package [@kinship2].
@@
 sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
-The R-RMSE metric is defined more thoroughly in the skater package vignette (see Supplementary Material), and may be a preferrable measure of classification accuracy when benchmarking relationship degree estimation.
-Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).
+The R-RMSE metric is defined more thoroughly in the skater package vignette (see Supplementary Material) and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation.
+Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second-degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).
 
@@
 Tools such as hap-IBD [@zhou2020], and IBIS [@seidman2020] detect shared IBD segments between individuals.
-The skater package includes functionality to take those IBD segments, compute shared genomic centimorgan (cM) length, and convert that shared cM to a kinship coefficient.
+The skater package includes functionality to take those IBD segments, compute shared genomic centimorgan (cM) length, and converts that shared cM to a kinship coefficient.
 In addition to inferred segments, these functions can estimate "truth" kinship from simulated IBD segments [@caballero2019].
@@
 
-**UNT FOLKS: NEED YOUR INPUT HERE!** The following rules should be followed:
+This work was supported in part by award 2019-DU-BX-0046 (Dense DNA Data for Enhanced Missing Persons Identification) to B.B., awarded by the National Institute of Justice, Office of Justice Programs, U.S.
+Department of Justice and by internal funds from the Center for Human Identification.
+The opinions, findings, and conclusions
+or recommendations expressed are those of the authors and do not necessarily reflect those of the U.S. Department of Justice.
 
--   The sentence should begin: 'This work was supported by ...' -
--   The full official funding agency name should be given, i.e. 'National Institutes of Health', not 'NIH' (full RIN-approved list of UK funding agencies)
--   Grant numbers should be given in brackets as follows: '[grant number xxxx]'
--   Multiple grant numbers should be separated by a comma as follows: '[grant numbers xxxx, yyyy]'
--   Agencies should be separated by a semi-colon (plus 'and' before the last funding agency)
--   Where individuals need to be specified for certain sources of funding the following text should be added after the relevant agency or grant number 'to [author initials]'.
+<!-- -   The sentence should begin: 'This work was supported by ...' - -->
+<!-- -   The full official funding agency name should be given, i.e. 'National Institutes of Health', not 'NIH' (full RIN-approved list of UK funding agencies) -->
+<!-- -   Grant numbers should be given in brackets as follows: '[grant number xxxx]' -->
+<!-- -   Multiple grant numbers should be separated by a comma as follows: '[grant numbers xxxx, yyyy]' -->
+<!-- -   Agencies should be separated by a semi-colon (plus 'and' before the last funding agency) -->
+<!-- -   Where individuals need to be specified for certain sources of funding the following text should be added after the relevant agency or grant number 'to [author initials]'. -->
 
-An example is given here: 'This work was supported by the National Institutes of Health [AA123456 to C.S., BB765432 to M.H.]; and the Alcohol & Education Research Council [hfygr667789].'
+<!-- An example is given here: 'This work was supported by the National Institutes of Health [AA123456 to C.S., BB765432 to M.H.]; and the Alcohol & Education Research Council [hfygr667789].' -->
 
-Oxford Journals will deposit all NIH-funded articles in PubMed Central.
+<!-- Oxford Journals will deposit all NIH-funded articles in PubMed Central.
 See Depositing articles in repositories -- information for authors for details.
 Authors must ensure that manuscripts are clearly indicated as NIH-funded using the guidelines above.
+-->

### 2021-07-15 paper/paper.tex
recompile

@@
 Scholz\,\textsuperscript{1}, Shakeel Jessa\,\textsuperscript{1}, Carlos
-Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2}, August
-Woener\,\textsuperscript{2}, Bruce Budowle\,\textsuperscript{2}
+Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2,3}, August
+Woener\,\textsuperscript{2,3}, Bruce Budowle\,\textsuperscript{2,3}
 }
@@
 22911, USA.\\
-\textsuperscript{2}University of North Texas Health Science Center, Fort
-Worth, Texas, 76107, USA.\\
+\textsuperscript{2}Center for Human Identification, University of North
+Texas Health Science Center, 3500 Camp Bowie Blvd., Fort Worth, TX,
+76107, USA\\
+\textsuperscript{3}Department of Microbiology, Immunology, and Genetics,
+University of North Texas Health Science Center, 3500 Camp Bowie Blvd.,
+Fort Worth, TX, 76107, USA\\
 }
@@
 forensics. There are multiple approaches to estimating relatedness
-between samples, including genome-wide measures such as those
+between samples, including genome-wide measures, such as those
 implemented in Plink \citep{purcell2007} or KING
 \citep{manichaikul2010}, and methods that rely on identity by descent
-(IBD) segment detection such as GERMLINE \citep{gusev2009}, hap-IBD
+(IBD) segment detection, such as GERMLINE \citep{gusev2009}, hap-IBD
 \citep{zhou2020}, and IBIS \citep{seidman2020}. Recent efforts focusing
@@
 family unit. The skater package contains the function
-\texttt{read\_fam()} to read in a PLINK-formatted .fam file, and another
+\texttt{read\_fam()} to read in a PLINK-formatted .fam file and another
 function \texttt{fam2ped()} to convert the content into a pedigree
@@
 new reciprocal RMSE (R-RMSE) metric. The R-RMSE metric is defined more
-thoroughly in the skater package vignette (see Supplementary Material),
-and may be a preferrable measure of classification accuracy when
+thoroughly in the skater package vignette (see Supplementary Material)
+and may be a preferable measure of classification accuracy when
 benchmarking relationship degree estimation. Taking the reciprocal of
@@
 larger penalties for more egregious misclassifications (e.g.,
-classifying a first-degree relative pair as second degree) than
+classifying a first-degree relative pair as second-degree) than
 misclassifications at more distant relationships (e.g., classifying a
@@
 includes functionality to take those IBD segments, compute shared
-genomic centimorgan (cM) length, and convert that shared cM to a kinship
-coefficient. In addition to inferred segments, these functions can
-estimate ``truth'' kinship from simulated IBD segments
+genomic centimorgan (cM) length, and converts that shared cM to a
+kinship coefficient. In addition to inferred segments, these functions
+can estimate ``truth'' kinship from simulated IBD segments
 \citep{caballero2019}. The \texttt{read\_ibd()} function reads pairwise
@@
 
-\textbf{UNT FOLKS: NEED YOUR INPUT HERE!} The following rules should be
-followed:
-
-\begin{itemize}
-\tightlist
-\item
-  The sentence should begin: `This work was supported by \ldots{}' -
-\item
-  The full official funding agency name should be given, i.e.~`National
-  Institutes of Health', not `NIH' (full RIN-approved list of UK funding
-  agencies)
-\item
-  Grant numbers should be given in brackets as follows: `{[}grant number
-  xxxx{]}'
-\item
-  Multiple grant numbers should be separated by a comma as follows:
-  `{[}grant numbers xxxx, yyyy{]}'
-\item
-  Agencies should be separated by a semi-colon (plus `and' before the
-  last funding agency)
-\item
-  Where individuals need to be specified for certain sources of funding
-  the following text should be added after the relevant agency or grant
-  number `to {[}author initials{]}'.
-\end{itemize}
-
-An example is given here: `This work was supported by the National
-Institutes of Health {[}AA123456 to C.S., BB765432 to M.H.{]}; and the
-Alcohol \& Education Research Council {[}hfygr667789{]}.'
-
-Oxford Journals will deposit all NIH-funded articles in PubMed Central.
-See Depositing articles in repositories -- information for authors for
-details. Authors must ensure that manuscripts are clearly indicated as
-NIH-funded using the guidelines above.
+This work was supported in part by award 2019-DU-BX-0046 (Dense DNA Data
+for Enhanced Missing Persons Identification) to B.B., awarded by the
+National Institute of Justice, Office of Justice Programs, U.S.
+Department of Justice and by internal funds from the Center for Human
+Identification. The opinions, findings, and conclusions or
+recommendations expressed are those of the authors and do not
+necessarily reflect those of the U.S. Department of Justice.
 

### 2021-07-15 paper/paper.Rmd
typo

@@
     affiliation: "2,3"
-  - name: August Woener
+  - name: August Woerner
     affiliation: "2,3"

### 2021-07-15 paper/paper.tex
typo

@@
 Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2,3}, August
-Woener\,\textsuperscript{2,3}, Bruce Budowle\,\textsuperscript{2,3}
+Woerner\,\textsuperscript{2,3}, Bruce Budowle\,\textsuperscript{2,3}
 }

### 2021-07-19 paper/paper.Rmd
fix @ahhgust name

--- 2021-07-15 paper/paper.Rmd
@@
     affiliation: "2,3"
-  - name: August Woener
+  - name: August E. Woerner
     affiliation: "2,3"

### 2021-07-19 paper/paper.Rmd
refer to suppl materials in ped2kinpair and plot_pedigree h/t @ahhgust

@@
 All pedigree processing from skater internally leverages a data structure from the kinship2 package [@kinship2].
-Further functions such as `plot_pedigree()` produce a multi-page PDF drawing a diagram of the pedigree for each family, while `ped2kinpair()` produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair.
+Further functions such as `plot_pedigree()` produce a multi-page PDF drawing a diagram of the pedigree for each family, while `ped2kinpair()` produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair (see Supplementary Material).
 

### 2021-07-19 paper/paper.tex

--- 2021-07-15 paper/paper.tex
@@
 Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2,3}, August
-Woener\,\textsuperscript{2,3}, Bruce Budowle\,\textsuperscript{2,3}
+E. Woerner\,\textsuperscript{2,3}, Bruce Budowle\,\textsuperscript{2,3}
 }
@@
 pairwise list of relationships between all individuals in the data with
-the expected kinship coefficients for each pair.
+the expected kinship coefficients for each pair (see Supplementary
+Material).
 
@@
 and may be a preferable measure of classification accuracy when
-benchmarking relationship degree estimation. Taking the reciprocal of
-the target and predicted degree in a typical RMSE calculation results in
-larger penalties for more egregious misclassifications (e.g.,
-classifying a first-degree relative pair as second-degree) than
-misclassifications at more distant relationships (e.g., classifying a
-fourth-degree relative pair as fifth-degree).
+benchmarking relationship degree estimation. In many kinship
+benchmarking analyses, classification error is treated in a categorical
+manner (exact match plus or minus one degree), neglecting the true
+amount of sharing as a real number. Taking the reciprocal of the target
+and predicted degree in a typical RMSE calculation results in larger
+penalties for more egregious misclassifications (e.g., classifying a
+first-degree relative pair as second-degree) than misclassifications at
+more distant relationships (e.g., classifying a fourth-degree relative
+pair as fifth-degree).
 

### 2021-07-19 paper/supplement-vignette.Rmd

@@
 
-One of the formats used by PLINK and other genetic analysis tools is the `.fam` file,[^plink-fam] which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
+One of the formats used by PLINK and other genetic analysis tools is the `.fam` file.[^plink-fam] A `.fam` file is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
 The package includes `read_fam()` to read files in this format:
@@
 
-Family structures imported from "fam" formated files can then be translated to the `pedigree` structure used by the `kinship2` package.[^kinship2-vignette] The "fam" format may include multiple families, and the `fam2ped()` function will collapse them all into a `tibble` with one row per family:
+Family structures imported from ".fam" formated files can then be translated to the `pedigree` structure used by the `kinship2` package.[^kinship2-vignette] The "fam" format may include multiple families, and the `fam2ped()` function will collapse them all into a `tibble` with one row per family:
 
@@
 
-The `plot_pedigree()` function from `skater` will iterate over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from each family:
+The `plot_pedigree()` function from `skater` will iterate over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from family:
 

### 2021-07-19 README.Rmd
update readme (remove note about PAT)

@@
 
-A GitHub [personal access token](https://github.com/settings/tokens) must be in the environment variable `GITHUB_PAT` or supplied as a string to `auth_token`. See `?remotes::install_github`.
+<!-- A GitHub [personal access token](https://github.com/settings/tokens) must be in the environment variable `GITHUB_PAT` or supplied as a string to `auth_token`. See `?remotes::install_github`.  -->
 

### 2021-07-20 README.Rmd
link to documentation in readme

@@
 
-The "Basic Usage" vignette steps through the primary functionality of the `skater` package:
+The ["Basic Usage" vignette](https://signaturescience.github.io/skater/articles/basic_usage.html) steps through the primary functionality of the `skater` package:
 
@@
 ```
+
+Full documentation: <https://signaturescience.github.io/skater/>.

### 2021-07-21 paper/paper.Rmd
affiliations

@@
   - code: 1
-    address: Signature Science, LLC., Charlottesville, Virginia, 22911, USA.
+    address: Signature Science, LLC., Austin, TX 78759, USA.
   - code: 2
-    address: Center for Human Identification, University of North Texas Health Science Center, 3500 Camp Bowie Blvd., Fort Worth, TX, 76107, USA
+    address: Center for Human Identification, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
   - code: 3
-    address: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, 3500 Camp Bowie Blvd., Fort Worth, TX, 76107, USA
+    address: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
 
@@
 The R-RMSE metric is defined more thoroughly in the skater package vignette (see Supplementary Material) and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation.
+In many kinship benchmarking analyses, classification error is treated in a categorical manner (exact match plus or minus one degree), neglecting the true amount of sharing as a real number.
 Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second-degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).

### 2021-07-21 paper/paper.tex
affiliations

@@
 \address{
-\textsuperscript{1}Signature Science, LLC., Charlottesville, Virginia,
-22911, USA.\\
+\textsuperscript{1}Signature Science, LLC., Austin, TX 78759, USA.\\
 \textsuperscript{2}Center for Human Identification, University of North
-Texas Health Science Center, 3500 Camp Bowie Blvd., Fort Worth, TX,
-76107, USA\\
+Texas Health Science Center, Fort Worth, TX 76107, USA\\
 \textsuperscript{3}Department of Microbiology, Immunology, and Genetics,
-University of North Texas Health Science Center, 3500 Camp Bowie Blvd.,
-Fort Worth, TX, 76107, USA\\
+University of North Texas Health Science Center, Fort Worth, TX 76107,
+USA\\
 }

### 2021-07-22 paper/paper.Rmd
- remove unpublished citations (Bioinformatics rules) - remove tidyverse citation (bumps us over 2 pages) - recompile

@@
 There are multiple approaches to estimating relatedness between samples, including genome-wide measures, such as those implemented in Plink [@purcell2007] or KING [@manichaikul2010], and methods that rely on identity by descent (IBD) segment detection, such as GERMLINE [@gusev2009], hap-IBD [@zhou2020], and IBIS [@seidman2020].
-Recent efforts focusing on benchmarking these methods [@ramstetter2017; @vries2021] have been aided by tools for simulating pedigrees and genome-wide SNP data [@caballero2019].
+Recent efforts focusing on benchmarking these methods [@ramstetter2017] have been aided by tools for simulating pedigrees and genome-wide SNP data [@caballero2019].
 Analyzing results from genome-wide SNP-based kinship analysis or comparing analyses to simulated data for benchmarking have to this point required writing one-off analysis functions or utility scripts that are seldom distributed with robust documentation, test suites, or narrative examples of usage.
@@
 
+<!-- The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis.
+Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
+The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem [@Wickham2019].
+-->
+
 The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis.
 Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
-The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem [@Wickham2019].
+The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem.
 
@@
 The skater package contains the function `read_fam()` to read in a PLINK-formatted .fam file and another function `fam2ped()` to convert the content into a pedigree object as a nested tibble with one row per family.
-All pedigree processing from skater internally leverages a data structure from the kinship2 package [@kinship2].
+All pedigree processing from skater internally leverages a data structure from the kinship2 package [@sinnwell2014].
 Further functions such as `plot_pedigree()` produce a multi-page PDF drawing a diagram of the pedigree for each family, while `ped2kinpair()` produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair (see Supplementary Material).
@@
 Department of Justice and by internal funds from the Center for Human Identification.
-The opinions, findings, and conclusions
-or recommendations expressed are those of the authors and do not necessarily reflect those of the U.S. Department of Justice.
+The opinions, findings, and conclusions or recommendations expressed are those of the authors and do not necessarily reflect those of the U.S.
+Department of Justice.
 
 <!-- -   The sentence should begin: 'This work was supported by ...' - -->
+
 <!-- -   The full official funding agency name should be given, i.e. 'National Institutes of Health', not 'NIH' (full RIN-approved list of UK funding agencies) -->
+
 <!-- -   Grant numbers should be given in brackets as follows: '[grant number xxxx]' -->
+
 <!-- -   Multiple grant numbers should be separated by a comma as follows: '[grant numbers xxxx, yyyy]' -->
+
 <!-- -   Agencies should be separated by a semi-colon (plus 'and' before the last funding agency) -->
+
 <!-- -   Where individuals need to be specified for certain sources of funding the following text should be added after the relevant agency or grant number 'to [author initials]'. -->

### 2021-07-22 paper/paper.tex
- remove unpublished citations (Bioinformatics rules) - remove tidyverse citation (bumps us over 2 pages) - recompile

@@
 \citep{zhou2020}, and IBIS \citep{seidman2020}. Recent efforts focusing
-on benchmarking these methods \citep{ramstetter2017, vries2021} have
-been aided by tools for simulating pedigrees and genome-wide SNP data
+on benchmarking these methods \citep{ramstetter2017} have been aided by
+tools for simulating pedigrees and genome-wide SNP data
 \citep{caballero2019}. Analyzing results from genome-wide SNP-based
@@
 The package is designed to adhere to ``tidy'' data analysis principles,
-and builds upon the tools released under the tidyverse R ecosystem
-\citep{Wickham2019}.
+and builds upon the tools released under the tidyverse R ecosystem.
 
@@
 processing from skater internally leverages a data structure from the
-kinship2 package \citep{kinship2}. Further functions such as
+kinship2 package \citep{sinnwell2014}. Further functions such as
 \texttt{plot\_pedigree()} produce a multi-page PDF drawing a diagram of

### 2021-09-07 paper/paper.Rmd
change to JOSS format

@@
 ---
-# title_full: "skater: An R package for SNP-based Kinship Analysis and Benchmarking"
-title_full: "skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
-title_short: "skater package for kinship analysis"
+title: "`skater`: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
 
-# One of the following below. See https://academic.oup.com/bioinformatics/pages/instructions_for_authors#Types%20of%20Manuscript
-# one of `Original Paper`, `Discovery Note`, `Applications Note`, `Review`, `Letter to the Editor`, `Conference Paper`    
-manuscript_type: Applications Note
+tags:
+  - R
+  - genetics
+  - kinship analysis
+  - SNPs
+  - sequencing
 
-# One of the following below. See https://academic.oup.com/bioinformatics/pages/instructions_for_authors#Detailed%20Scope%20Guidelines
-# Genome analysis
-# Sequence analysis
-# Phylogenetics
-# Structural bioinformatics
-# Gene expression
-# Genetic and population analysis
-# Systems biology
-# Data and text mining
-# Databases and ontologies
-# Bioimage informatics
-subject_section:  Genetics and Population Analysis
+citation_author: Turner, S.D. et al.
 
-access_date:
-  - day: DD
-    month: MM
-    year: YY
-
-author:
+authors:
   - name: Stephen D. Turner
+    orcid: 0000-0001-9140-9028
     affiliation: "1"
   - name: V.P. Nagraj
+    orcid: 0000-0003-0060-566X
     affiliation: "1"
   - name: Matthew Scholz
+    orcid: 0000-0003-3686-1227
     affiliation: "1"
@@
   - name: Jianye Ge
+    orcid: 0000-0001-8724-075X
     affiliation: "2,3"
   - name: August E. Woerner
+    orcid: 0000-0002-9372-1127
     affiliation: "2,3"
   - name: Bruce Budowle
+    orcid: 0000-0003-4116-2930
     affiliation: "2,3"
 
-address:
-  - code: 1
-    address: Signature Science, LLC., Austin, TX 78759, USA.
-  - code: 2
-    address: Center for Human Identification, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
-  - code: 3
-    address: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
+affiliations:
+  - name: Signature Science, LLC., Austin, TX 78759, USA.
+    index: 1
+  - name: Center for Human Identification, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
+    index: 2
+  - name: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
+    index: 3
 
-# Full email address to be given, preferably an institution email address.  
-corresponding_author_text: "To whom correspondence should be addressed. E-mail: sturner\\@signaturescience.com"
-corresponding_author_email: "sturner\\@signaturescience.com"
+date: 07 September 2021
+year: 2021
+bibliography: bibliography.bib
+output: rticles::joss_article
+csl: apa.csl
+journal: JOSS
 
-motivation: |
-  SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation.
-A dedicated software package that consistently and intuitively implements this analysis functionality is needed.
-
-results: |
-  Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**.
-The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
-
-availability: |
-  The skater package is implemented as an R package and is released under the MIT license at https\:\/\/github.com/signaturescience/skater.
-Documentation is available at https\:\/\/signaturescience.github.io/skater.
-
-supplementary_info: |
-  Supplementary data are available at Bioinformatics Online.
-
-output:
-  rticles::bioinformatics_article:
-    citation_package: natbib
-
-bibliography: bibliography.bib
 ---
@@
 
+# Summary
+
+**Motivation:** SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation.
+A dedicated software package that consistently and intuitively implements this analysis functionality is needed.
+**Results:**  Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**.
+The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
+**Availability:** The skater package is implemented as an R package and is released under the MIT license at https\:\/\/github.com/signaturescience/skater.
+Documentation is available at https\:\/\/signaturescience.github.io/skater.
+
 # Introduction
@@
 
-<!-- The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis.
+The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis.
 Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
 The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem [@Wickham2019].
--->
-
-The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis.
-Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
-The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem.
 
@@
 
-# Funding {.unnumbered}
+# Acknowledgements
 
@@
 -->
+
+# References

### 2021-09-07 paper/paper.Rmd
remove references to supplementary material

@@
 ---
-title: "`skater`: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
+title: "skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
 
@@
 ```{r, echo=FALSE, eval=FALSE}
-## Run this chunk interactively if you update the package vignette.
+## Run this chunk interactively if you update the package vignette as supplementary material.
 ## This renders the vignette as a PDF with a new title.
@@
 All pedigree processing from skater internally leverages a data structure from the kinship2 package [@sinnwell2014].
-Further functions such as `plot_pedigree()` produce a multi-page PDF drawing a diagram of the pedigree for each family, while `ped2kinpair()` produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair (see Supplementary Material).
+Further functions such as `plot_pedigree()` produce a multi-page PDF drawing a diagram of the pedigree for each family, while `ped2kinpair()` produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair (see skater package vignette).
 
@@
 sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
-The R-RMSE metric is defined more thoroughly in the skater package vignette (see Supplementary Material) and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation.
+The R-RMSE metric is defined more thoroughly in the skater package vignette and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation.
 In many kinship benchmarking analyses, classification error is treated in a categorical manner (exact match plus or minus one degree), neglecting the true amount of sharing as a real number.
@@
 The skater R package provides a robust software package for data import, manipulation, and analysis tasks typically encountered when working with SNP-based kinship analysis tools.
-All package functions are internally documented with examples, and the package contains a vignette demonstrating usage, inputs, outputs, and interpretation of all key functions (see Supplementary Material).
+All package functions are internally documented with examples, and the package contains a vignette demonstrating usage, inputs, outputs, and interpretation of all key functions.
 The package contains internal tests that are automatically run with continuous integration via GitHub Actions whenever the package code is updated.

### 2021-09-07 paper/paper.Rmd
orcid IDs, remove comments

@@
   - name: Shakeel Jessa
+    orcid: 0000-0002-1953-6839
     affiliation: "1"
   - name: Carlos Acevedo
+    orcid: 0000-0002-1953-6839
     affiliation: "1"
@@
 
-<!-- # Acknowledgements {.unnumbered} -->
-
-<!-- These should be included at the end of the text and not in footnotes. Please ensure you acknowledge all sources of funding, see funding section below. -->
-
-<!-- Details of all funding sources for the work in question should be given in a separate section entitled 'Funding'. This should appear before the 'Acknowledgements' section. -->
-
 # Acknowledgements
@@
 
-<!-- -   The sentence should begin: 'This work was supported by ...' - -->
-
-<!-- -   The full official funding agency name should be given, i.e. 'National Institutes of Health', not 'NIH' (full RIN-approved list of UK funding agencies) -->
-
-<!-- -   Grant numbers should be given in brackets as follows: '[grant number xxxx]' -->
-
-<!-- -   Multiple grant numbers should be separated by a comma as follows: '[grant numbers xxxx, yyyy]' -->
-
-<!-- -   Agencies should be separated by a semi-colon (plus 'and' before the last funding agency) -->
-
-<!-- -   Where individuals need to be specified for certain sources of funding the following text should be added after the relevant agency or grant number 'to [author initials]'. -->
-
-<!-- An example is given here: 'This work was supported by the National Institutes of Health [AA123456 to C.S., BB765432 to M.H.]; and the Alcohol & Education Research Council [hfygr667789].' -->
-
-<!-- Oxford Journals will deposit all NIH-funded articles in PubMed Central.
-See Depositing articles in repositories -- information for authors for details.
-Authors must ensure that manuscripts are clearly indicated as NIH-funded using the guidelines above.
--->
-
 # References

### 2021-11-29 paper/paper.Rmd
first cut at moving to F1000 research

@@
 ---
-title: "skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
-
-tags:
-  - R
-  - genetics
-  - kinship analysis
-  - SNPs
-  - sequencing
-
-citation_author: Turner, S.D. et al.
-
-authors:
+title: "**skater**: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
+author: 
   - name: Stephen D. Turner
-    orcid: 0000-0001-9140-9028
-    affiliation: "1"
-  - name: V.P. Nagraj
-    orcid: 0000-0003-0060-566X
-    affiliation: "1"
+    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
+  - name: V. P. Nagraj
+    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
   - name: Matthew Scholz
-    orcid: 0000-0003-3686-1227
-    affiliation: "1"
+    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
   - name: Shakeel Jessa
-    orcid: 0000-0002-1953-6839
-    affiliation: "1"
+    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
   - name: Carlos Acevedo
-    orcid: 0000-0002-1953-6839
-    affiliation: "1"
+    affiliation: Signature Science, LLC., Austin, TX 78759, USA.
   - name: Jianye Ge
-    orcid: 0000-0001-8724-075X
-    affiliation: "2,3"
+    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
   - name: August E. Woerner
-    orcid: 0000-0002-9372-1127
-    affiliation: "2,3"
+    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
   - name: Bruce Budowle
-    orcid: 0000-0003-4116-2930
-    affiliation: "2,3"
-
-affiliations:
-  - name: Signature Science, LLC., Austin, TX 78759, USA.
-    index: 1
-  - name: Center for Human Identification, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
-    index: 2
-  - name: Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA
-    index: 3
-
-date: 07 September 2021
-year: 2021
+    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
+# Abstract up to 300 words. No citations, minimal abbreviations.
+abstract: | 
+    | 
+    | **Motivation:** SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation.
+A dedicated software package that consistently and intuitively implements this analysis functionality is needed.
+    | **Results:**  Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**.
+The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
+    | **Availability:** The skater package is implemented as an R package and is released under the MIT license at https\:\/\/github.com/signaturescience/skater.
+Documentation is available at https\:\/\/signaturescience.github.io/skater.
+keywords: bioinformatics, kinship, R, genealogy, SNPs, single nucleotide polymorphisms, relatedness
 bibliography: bibliography.bib
-output: rticles::joss_article
-csl: apa.csl
-journal: JOSS
-
+output: BiocWorkflowTools::f1000_article
 ---
 
-```{r, echo=FALSE, eval=FALSE}
-## Run this chunk interactively if you update the package vignette as supplementary material.
-## This renders the vignette as a PDF with a new title.
+**R version**: `r R.version.string`
 
-# Where is the package vignette?
-vpath <- here::here("vignettes/basic_usage.Rmd")
-# Where do you want to save the new modified vignette rmarkdown?
-spath <- here::here("paper/supplement-vignette.Rmd")
-# Read in the vignette rmarkdown text
-vtext <- readLines(vpath)
-# Create the supplemental material rmarkdown by replacing the title YAML
-stext <- sub(pattern='title: "Basic Usage"', 
-             replacement = 'title: |\n    | Supplementary Material for:\n    | skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation\n    | _(skater v0.1.0 package vignette)_', 
-             x=vtext)
-# Write out that modified yaml to the supplementary material rmarkdown
-writeLines(text=stext, con=spath)
-# Render it to PDF
-rmarkdown::render(spath, output_format="pdf_document")
-```
+<!-- **Bioconductor version**: `r BiocManager::version()` -->
 
-# Summary
+<!--  Update the name of the package below to be the name of the workflow package you are working on  -->
 
-**Motivation:** SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation.
-A dedicated software package that consistently and intuitively implements this analysis functionality is needed.
-**Results:**  Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**.
-The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
-**Availability:** The skater package is implemented as an R package and is released under the MIT license at https\:\/\/github.com/signaturescience/skater.
-Documentation is available at https\:\/\/signaturescience.github.io/skater.
+**skater package version**: `r packageVersion("skater")`
+
+----
 
@@
 
-# The skater package
+# Methods
+
+## Implementation
+
+<!-- For software tool papers, this section should address how the tool works and any relevant technical details required for implementation of the tool by other developers. -->
 
 The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis.
-Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
-The package is designed to adhere to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem [@Wickham2019].
+Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data, described in full in the _Use Cases_ section below.
+The package adheres to "tidy" data analysis principles, and builds upon the tools released under the tidyverse R ecosystem [@Wickham2019].
+
+The skater package is hosted in the Comprehensive R Archive Network (CRAN) which is the main repository for R packages: <http://CRAN.R-project.org/package=skater>.
+Users can install skater in R by executing the following code:
+
+```{r, eval=FALSE}
+install.packages("skater")
+```
+
+Alternatively, the development version of skater is available on GitHub at <https://github.com/signaturescience/skater>.
+The development version may contain new features which are not yet available in the version hosted on CRAN.
+This version can be installed using the `install_github()` function in the devtools package:
+
+```{r, eval=FALSE}
+install.packages("devtools")
+devtools::install_github("signaturescience/skater", build_vignettes=TRUE)
+```
+
+When installing skater, other packages which skater depends on are automatically installed, including magritr, tibble, dplyr, tidyr, readr, purrr, kinship2, corrr, rlang, and others.
+
+## Operation
+
+<!-- This part of the methods should include the minimal system requirements needed to run the software and an overview of the workflow for the tool for users of the tool. -->
+
+Minimal system requirements for installing and using skater include R (version 3.0.0 or higher) and several tidyverse packages [@Wickham2019] that many R users will already have installed.
+Use cases are demonstrated in detail below.
+In summary, the skater package has functions for:
+
+- Reading in various output files produced by commonly used tools in SNP-based kinship analysis
+- Pedigree parsing, manpulation, and analysis
+- Relationship degree inference
+- Benchmarking and assessing relationship classification accuracy
+- IBD segment analysis post-processing
+
+A comprehensive reference for all the functions in the skater package is available at <https://signaturescience.github.io/skater/>.
+
+<!-- # Results  -->
+
+<!-- Optional - only if novel data or analyses are included -->
+
+<!-- This section is only required if the paper includes novel data or analyses, and should be written as a traditional results section. -->
+
+# Use Cases
+
+<!-- Optional - only if NO new datasets are included -->
+
+<!-- This section is required if the paper does not include novel data or analyses.  -->
+<!-- Examples of input and output files should be provided with some explanatory context.
+Any novel or complex variable parameters should also be explained in sufficient detail to allow users to understand and use the tool's functionality.
+-->
 
@@
 
-# Conclusion
+<!-- # Discussion  -->
+
+<!-- Optional - only if novel data or analyses are included -->
+
+<!-- This section is only required if the paper includes novel data or analyses, and should be written in the same style as a traditional discussion section. -->
+
+<!-- Please include a brief discussion of allowances made (if any) for controlling bias or unwanted sources of variability, and the limitations of any novel datasets. -->
+
+<!-- # Conclusions  -->
+
+<!-- Optional - only if novel data or analyses are included -->
+
+<!-- This section is only required if the paper includes novel data or analyses, and should be written as a traditional conclusion. -->
+
+# Summary
+
+<!-- Optional - only if NO new datasets are included -->
 
@@
 
-# Acknowledgements
+<!-- # Data availability  -->
+
+<!-- Optional - only if novel data or analyses are included -->
+
+<!-- Please add details of where any datasets that are mentioned in the paper, and that have not have not previously been formally published, can be found.
+If previously published datasets are mentioned, these should be cited in the references, as per usual scholarly conventions.
+-->
+
+# Software availability
+
+<!-- This section will be generated by the Editorial Office before publication. Authors are asked to provide some initial information to assist the Editorial Office, as detailed below. -->
+
+<!-- 1. URL link to where the software can be downloaded from or used by a non-coder (AUTHOR TO PROVIDE; optional) -->
+<!-- 2. URL link to the author's version control system repository containing the source code (AUTHOR TO PROVIDE; required) -->
+<!-- 3. Link to source code as at time of publication (*F1000Research* TO GENERATE) -->
+<!-- 4. Link to archived source code as at time of publication (*F1000Research* TO GENERATE) -->
+<!-- 5. Software license (AUTHOR TO PROVIDE; required) -->
+
+1. Software available from: <http://CRAN.R-project.org/package=skater>.
+1. Source code available from: <https://github.com/signaturescience/skater>.
+1. Archived source code at time of publication: **FIXME Zenodo DOI to come here**.
+1. Software license: MIT License.
+
+# Author information
+
+SDT and VPN developed the R package.
+
+All authors contributed to method development.
+
+SDT wrote the first draft of the manuscript.
+
+All authors assisted with manuscript revision.
+
+All authors read and approved the final manuscript.
+
+# Competing interests
+
+<!-- All financial, personal, or professional competing interests for any of the authors that could be construed to unduly influence the content of the article must be disclosed and will be displayed alongside the article.
+If there are no relevant competing interests to declare, please add the following: 'No competing interests were disclosed'.
+-->
+
+No competing interests were disclosed.
+
+# Grant information
+
+<!-- Please state who funded the work discussed in this article, whether it is your employer, a grant funder etc.
+Please do not list funding that you have that is not relevant to this specific piece of research.
+For each funder, please state the funder's name, the grant number where applicable, and the individual to whom the grant was assigned.
+If your work was not funded by any grants, please include the line: 'The author(s) declared that no grants were involved in supporting this work.' -->
 
@@
 
-# References
+<!-- # Acknowledgments -->
+
+<!-- This section should acknowledge anyone who contributed to the research or the article but who does not qualify as an author based on the criteria provided earlier (e.g.
+someone or an organization that provided writing assistance).
+Please state how they contributed; authors should obtain permission to acknowledge from all those mentioned in the Acknowledgments section.
+-->
+
+<!-- Please do not list grant funding in this section. -->
+
+<!--
+
+# USING R MARKDOWN
+
+Some examples of commonly used markdown syntax are listed below, to help you get started.
+
+## Cross-references
+
+For portability between different output formats, use the syntax introduced by *bookdown*, such as `(\#label)` for labels and `\@ref(label)` for cross-references.
+The following sections provide examples of referencing tables, figures, and equations.
+
+## Citations
+
+You can include references in a standard Bibtex file.
+The name of this file is given in the header of the markdown document (in our case it is *sample.bib*).
+References to entries in the Bibtex file are made using square brackets and use an @ plus the key for the entry you are referencing [@Smith:2012qr].
+You can combine multiple entries by separating them with a semi-colon [@Smith:2012qr; @Smith:2013jd].
+The default bibliography style uses numerical citations.  For superscript or author-year citations set the header metadata field `natbiboptions` to either `super` or `round`, respectively.
+
+If you specify a figure caption to a code chunk using the chunk option `fig.cap`, the plot will be automatically labeled and numbered.
+The figure label is generated from the label of the code chunk by prefixing it with `fig:`, e.g., see Figure \@ref(fig:plot).
+
+## Tables
+
+Markdown syntax tends to lack some of the more sophisticated formatting features available in LaTeX, so you may need to edit the tables later to get the desired format.
+
+| First name  | Last Name | Grade |
+| ----------- | --------- | ----- |
+| John        | Doe       |   7.5 |
+| Richard     | Miles     |     2 |
+
+Table: Caption to table.
+
+Just like figures, tables with captions will also be numbered and can be referenced.
+Captions are entered as a paragraph beginning with the string "Table:" (or just ":"), which may appear either before or after the table.
+A label for the table should appear in the beginning of the caption in the form of `(\#tab:label)`, e.g., see Table \@ref(tab:table).
+
+: (\#tab:table) A table with text justification.
+
+| First name  | Last Name | Grade |
+| ----------- | :-------: | ----: |
+| John        | Doe       |   7.5 |
+| Richard     | Miles     |     2 |
+
+## Figures
+
+You can include static figures (i.e. no generated by code) using the `include_graphics()` function from the **knitr** package, in a standard code chunk.
+
+You can again use the `fig.cap` option to provide the figure caption, and reference the image based on the code chunk label.
+You can also use options such as `fig.align` and `fig.width` to adjust the position and size of the image within the final document, e.g.
+Figure \@ref(fig:frog-picture) is a frog.
+
+Alternatively, you can use the standard markdown syntax like so:
+
+![This is a smaller version of the same picture, inserted using the standard markdown syntax](frog.jpg){width=2cm height=2cm}
+
+Please give figures appropriate filenames, e.g.: figure1.pdf, figure2.png.
+
+Figure legends should briefly describe the key messages of the figure such that the figure can stand alone from the main text.
+However, all figures should also be discussed in the article text.
+Each legend should have a concise title of no more than 15 words.
+The legend itself should be succinct, while still explaining all symbols and abbreviations.
+Avoid lengthy descriptions of methods.
+
+For any figures reproduced from another publication (as long as appropriate permission has been obtained from the copyright holder -see under the heading 'Submission'), please include a line in the legend to state that: 'This figure has been reproduced with kind permission from [include original publication citation]'.
+
+-->

### 2021-11-29 paper/paper.tex
first cut at moving to F1000 research

@@
-\documentclass{bioinfo}
-\copyrightyear{2016} \pubyear{2016}
+\documentclass[9pt,a4paper,]{extarticle}
 
-%% Some pieces required from the pandoc template
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\usepackage{f1000_styles}
 
-% Pandoc citation processing
+\usepackage[pdfborder={0 0 0}]{hyperref}
 
-% hyperref makes the margins screwy.
-% https://groups.google.com/forum/#!topic/latexusersgroup/4W_SwGk6zx4
-% http://ansuz.sooke.bc.ca/software/latex-tricks.php
-% \usepackage[colorlinks=true, allcolors=blue]{hyperref}
+\usepackage[numbers]{natbib}
+\bibliographystyle{unsrtnat}
 
-\access{Advance Access Publication Date:   }
-\appnotes{Applications Note}
+%% maxwidth is the original width if it is less than linewidth
+%% otherwise use linewidth (to make sure the graphics do not exceed the margin)
+\makeatletter
+\def\maxwidth{ %
+  \ifdim\Gin@nat@width>\linewidth
+    \linewidth
+  \else
+    \Gin@nat@width
+  \fi
+}
+\makeatother
+
+\usepackage{color}
+\usepackage{fancyvrb}
+\newcommand{\VerbBar}{|}
+\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
+\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
+% Add ',fontsize=\small' for more characters per line
+\usepackage{framed}
+\definecolor{shadecolor}{RGB}{248,248,248}
+\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
+\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
+\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
+\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\BuiltInTok}[1]{#1}
+\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
+\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
+\newcommand{\ExtensionTok}[1]{#1}
+\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
+\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\ImportTok}[1]{#1}
+\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
+\newcommand{\NormalTok}[1]{#1}
+\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
+\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
+\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
+\newcommand{\RegionMarkerTok}[1]{#1}
+\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
+\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
+\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
+
+% disable code chunks background
+%\renewenvironment{Shaded}{}{}
+
+% disable section numbers
+\setcounter{secnumdepth}{0}
+
+%% added by MLS, this is not in the F1000 style by default %%
+
+\hypersetup{unicode=true,
+            pdftitle={skater: An R package for SNP-based Kinship Analysis, Testing, and Evaluation},
+            pdfkeywords={bioinformatics, kinship, R, genealogy, SNPs, single nucleotide polymorphisms, relatedness},
+            pdfborder={0 0 0},
+            breaklinks=true}
+
+%% End added by MLS %%
+
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{6pt plus 2pt minus 1pt}
 
 \begin{document}
-\firstpage{1}
+\pagestyle{front}
 
-\subtitle{Genetics and Population Analysis}
+\title{\textbf{skater}: An R package for SNP-based Kinship Analysis, Testing, and Evaluation}
 
-\title[skater package for kinship analysis]{skater: An R package for
-SNP-based Kinship Analysis, Testing, and Evaluation}
-
-\author[FirstAuthorLastName \textit{et~al}.]{
-Stephen D. Turner\,\textsuperscript{1}, V.P.
-Nagraj\,\textsuperscript{1}, Matthew
-Scholz\,\textsuperscript{1}, Shakeel Jessa\,\textsuperscript{1}, Carlos
-Acevedo\,\textsuperscript{1}, Jianye Ge\,\textsuperscript{2,3}, August
-E. Woerner\,\textsuperscript{2,3}, Bruce Budowle\,\textsuperscript{2,3}
-}
-
-\address{
-\textsuperscript{1}Signature Science, LLC., Austin, TX 78759, USA.\\
-\textsuperscript{2}Center for Human Identification, University of North
-Texas Health Science Center, Fort Worth, TX 76107, USA\\
-\textsuperscript{3}Department of Microbiology, Immunology, and Genetics,
-University of North Texas Health Science Center, Fort Worth, TX 76107,
-USA\\
-}
-
-\corresp{To whom correspondence should be addressed. E-mail:
-sturner@signaturescience.com}
-
-\history{Received on XXX; revised on XXX; accepted on XXX}
-
-\editor{Associate Editor: XXX}
-
-\abstract{
-\textbf{Motivation:} SNP-based kinship analysis with genome-wide
-relationship estimation and IBD segment analysis methods produces
-results that often require further downstream processing and
-manipulation. A dedicated software package that consistently and
-intuitively implements this analysis functionality is needed.\\
-\textbf{Results:} Here we present the skater R package for
-\textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting,
-and \textbf{e}valuation with \textbf{R}. The skater package contains a
-suite of well-documented tools for importing, parsing, and analyzing
-pedigree data, performing relationship degree inference, benchmarking
-relationship degree classification, and summarizing IBD segment data.\\
-\textbf{Availability:} The skater package is implemented as an R package
-and is released under the MIT license at
-https://github.com/signaturescience/skater. Documentation is available
-at https://signaturescience.github.io/skater.\\
-\textbf{Contact:}sturner@signaturescience.com\\
-\textbf{Supplementary information:} Supplementary data are available at
-Bioinformatics Online.}
+\author[1]{Stephen D. Turner}
+\author[1]{V. P. Nagraj}
+\author[1]{Matthew Scholz}
+\author[1]{Shakeel Jessa}
+\author[1]{Carlos Acevedo}
+\author[2]{Jianye Ge}
+\author[2]{August E. Woerner}
+\author[2]{Bruce Budowle}
+\affil[1]{Signature Science, LLC., Austin, TX 78759, USA.}
+\affil[2]{Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA.}
 
 \maketitle
+\thispagestyle{front}
 
-\section{Introduction}
+\begin{abstract}
+\hfill\break
+\textbf{Motivation:} SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation.
+A dedicated software package that consistently and intuitively implements this analysis functionality is needed.\\
+\textbf{Results:} Here we present the skater R package for \textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation with \textbf{R}.
+The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.\\
+\textbf{Availability:} The skater package is implemented as an R package and is released under the MIT license at https://github.com/signaturescience/skater.
+Documentation is available at https://signaturescience.github.io/skater.
+\end{abstract}
 
-Inferring familial relationships between individuals using genetic data
-is a common practice in population genetics, medical genetics, and
-forensics. There are multiple approaches to estimating relatedness
-between samples, including genome-wide measures, such as those
-implemented in Plink \citep{purcell2007} or KING
-\citep{manichaikul2010}, and methods that rely on identity by descent
-(IBD) segment detection, such as GERMLINE \citep{gusev2009}, hap-IBD
-\citep{zhou2020}, and IBIS \citep{seidman2020}. Recent efforts focusing
-on benchmarking these methods \citep{ramstetter2017} have been aided by
-tools for simulating pedigrees and genome-wide SNP data
-\citep{caballero2019}. Analyzing results from genome-wide SNP-based
-kinship analysis or comparing analyses to simulated data for
-benchmarking have to this point required writing one-off analysis
-functions or utility scripts that are seldom distributed with robust
-documentation, test suites, or narrative examples of usage. There is a
-need in the field for a well-documented software package with a
-consistent design and API that contains functions to assist with
-downstream manipulation, benchmarking, and analysis of SNP-based kinship
-assessment methods. Here we present the skater package for
-\textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting,
-and \textbf{e}valuation with \textbf{R}.
+\section*{Keywords}
+bioinformatics, kinship, R, genealogy, SNPs, single nucleotide polymorphisms, relatedness
 
-\section{The skater package}
+\clearpage
+\pagestyle{main}
 
-The skater package provides an intuitive collection of analysis and
-utility functions for SNP-based kinship analysis. Functions in the
-package include tools for importing, parsing, and analyzing pedigree
-data, performing relationship degree inference, benchmarking
-relationship degree classification, and summarizing IBD segment data.
-The package is designed to adhere to ``tidy'' data analysis principles,
-and builds upon the tools released under the tidyverse R ecosystem.
+\textbf{R version}: R version 4.0.4 (2021-02-15)
 
-\subsection{Pedigree parsing, manipulation, and analysis}
+\textbf{skater package version}: 0.1.0
 
-The skater package has several functions for importing, parsing, and
-analyzing pedigree data. Pedigrees define familial relationships in a
-hierarchical structure. Many genomics tools for working with pedigrees
-start with a .fam file, which is a tabular format with one row per
-individual and columns for unique IDs of the mother, father, and the
-family unit. The skater package contains the function
-\texttt{read\_fam()} to read in a PLINK-formatted .fam file and another
-function \texttt{fam2ped()} to convert the content into a pedigree
-object as a nested tibble with one row per family. All pedigree
-processing from skater internally leverages a data structure from the
-kinship2 package \citep{sinnwell2014}. Further functions such as
-\texttt{plot\_pedigree()} produce a multi-page PDF drawing a diagram of
-the pedigree for each family, while \texttt{ped2kinpair()} produces a
-pairwise list of relationships between all individuals in the data with
-the expected kinship coefficients for each pair (see Supplementary
-Material).
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
 
-\subsection{Relationship degree inference and benchmarking}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
-The skater package includes functions to translate kinship coefficients
-to relationship degrees. The kinship coefficients could come from
-\texttt{ped2kinpair()} or other kinship estimation software.
+Inferring familial relationships between individuals using genetic data is a common practice in population genetics, medical genetics, and forensics.
+There are multiple approaches to estimating relatedness between samples, including genome-wide measures, such as those implemented in Plink \citep{purcell2007} or KING \citep{manichaikul2010}, and methods that rely on identity by descent (IBD) segment detection, such as GERMLINE \citep{gusev2009}, hap-IBD \citep{zhou2020}, and IBIS \citep{seidman2020}.
+Recent efforts focusing on benchmarking these methods \citep{ramstetter2017} have been aided by tools for simulating pedigrees and genome-wide SNP data \citep{caballero2019}.
+Analyzing results from genome-wide SNP-based kinship analysis or comparing analyses to simulated data for benchmarking have to this point required writing one-off analysis functions or utility scripts that are seldom distributed with robust documentation, test suites, or narrative examples of usage.
+There is a need in the field for a well-documented software package with a consistent design and API that contains functions to assist with downstream manipulation, benchmarking, and analysis of SNP-based kinship assessment methods.
+Here we present the skater package for \textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation with \textbf{R}.
 
-The \texttt{dibble()} function creates a \textbf{d}egree
-\textbf{i}nference t\textbf{ibble}, with degrees up to the specified
-maximum degree resolution, expected kinship coefficient, and lower and
-upper inference ranges as defined in \citet{manichaikul2010}. The
-\texttt{kin2degree()} function infers the relationship degree given a
-kinship coefficient and a maximum degree resolution (e.g., 7th-degree
-relatives) up to which anything more distant is classified as unrelated.
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
 
-Once estimated kinship is converted to degree, it may be of interest to
-compare the inferred degree to known degrees of relatedness. When
-aggregated over many relationships and inferences, this can help
-benchmark performance of a particular kinship analysis method. The
-skater package adapts a \texttt{confusion\_matrix()} function from
-\citet{clark2021} to provide standard contingency table metrics
-(e.g.~sensitivity, specificity, PPV, precision, recall, F1, etc.) with a
-new reciprocal RMSE (R-RMSE) metric. The R-RMSE metric is defined more
-thoroughly in the skater package vignette (see Supplementary Material)
-and may be a preferable measure of classification accuracy when
-benchmarking relationship degree estimation. In many kinship
-benchmarking analyses, classification error is treated in a categorical
-manner (exact match plus or minus one degree), neglecting the true
-amount of sharing as a real number. Taking the reciprocal of the target
-and predicted degree in a typical RMSE calculation results in larger
-penalties for more egregious misclassifications (e.g., classifying a
-first-degree relative pair as second-degree) than misclassifications at
-more distant relationships (e.g., classifying a fourth-degree relative
-pair as fifth-degree).
+\hypertarget{implementation}{%
+\subsection{Implementation}\label{implementation}}
 
-\subsection{IBD segment analysis}
+The skater package provides an intuitive collection of analysis and utility functions for SNP-based kinship analysis.
+Functions in the package include tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data, described in full in the \emph{Use Cases} section below.
+The package adheres to ``tidy'' data analysis principles, and builds upon the tools released under the tidyverse R ecosystem \citep{Wickham2019}.
 
-Tools such as hap-IBD \citep{zhou2020}, and IBIS \citep{seidman2020}
-detect shared IBD segments between individuals. The skater package
-includes functionality to take those IBD segments, compute shared
-genomic centimorgan (cM) length, and converts that shared cM to a
-kinship coefficient. In addition to inferred segments, these functions
-can estimate ``truth'' kinship from simulated IBD segments
-\citep{caballero2019}. The \texttt{read\_ibd()} function reads pairwise
-IBD segments from IBD inference tools and from simulated IBD segments.
-The \texttt{read\_map()} function reads in genetic map in a standard
-format which is required to translate the total centimorgans shared IBD
-to a kinship coefficient using the \texttt{ibd2kin()} function.
+The skater package is hosted in the Comprehensive R Archive Network (CRAN) which is the main repository for R packages: \url{http://CRAN.R-project.org/package=skater}.
+Users can install skater in R by executing the following code:
 
-\section{Conclusion}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{install.packages}\NormalTok{(}\StringTok{"skater"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-The skater R package provides a robust software package for data import,
-manipulation, and analysis tasks typically encountered when working with
-SNP-based kinship analysis tools. All package functions are internally
-documented with examples, and the package contains a vignette
-demonstrating usage, inputs, outputs, and interpretation of all key
-functions (see Supplementary Material). The package contains internal
-tests that are automatically run with continuous integration via GitHub
-Actions whenever the package code is updated. The skater package is
-permissively licensed (MIT) and is easily extensible to accommodate
-outputs from new genome-wide relatedness and IBD segment methods as they
-become available.
+Alternatively, the development version of skater is available on GitHub at \url{https://github.com/signaturescience/skater}.
+The development version may contain new features which are not yet available in the version hosted on CRAN.
+This version can be installed using the \texttt{install\_github()} function in the devtools package:
 
-\section*{Funding}
-\addcontentsline{toc}{section}{Funding}
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}
+\NormalTok{devtools}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"signaturescience/skater"}\NormalTok{, }\AttributeTok{build\_vignettes=}\ConstantTok{TRUE}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
 
-This work was supported in part by award 2019-DU-BX-0046 (Dense DNA Data
-for Enhanced Missing Persons Identification) to B.B., awarded by the
-National Institute of Justice, Office of Justice Programs, U.S.
-Department of Justice and by internal funds from the Center for Human
-Identification. The opinions, findings, and conclusions or
-recommendations expressed are those of the authors and do not
-necessarily reflect those of the U.S. Department of Justice.
+When installing skater, other packages which skater depends on are automatically installed, including magritr, tibble, dplyr, tidyr, readr, purrr, kinship2, corrr, rlang, and others.
 
-% Bibliography
-\bibliographystyle{natbib}
-\bibliography{bibliography.bib}
+\hypertarget{operation}{%
+\subsection{Operation}\label{operation}}
+
+Minimal system requirements for installing and using skater include R (version 3.0.0 or higher) and several tidyverse packages \citep{Wickham2019} that many R users will already have installed.
+Use cases are demonstrated in detail below.
+In summary, the skater package has functions for:
+
+\begin{itemize}
+\item
+  Reading in various output files produced by commonly used tools in SNP-based kinship analysis
+\item
+  Pedigree parsing, manpulation, and analysis
+\item
+  Relationship degree inference
+\item
+  Benchmarking and assessing relationship classification accuracy
+\item
+  IBD segment analysis post-processing
+\end{itemize}
+
+A comprehensive reference for all the functions in the skater package is available at \url{https://signaturescience.github.io/skater/}.
+
+\hypertarget{use-cases}{%
+\section{Use Cases}\label{use-cases}}
+
+\hypertarget{pedigree-parsing-manipulation-and-analysis}{%
+\subsection{Pedigree parsing, manipulation, and analysis}\label{pedigree-parsing-manipulation-and-analysis}}
+
+The skater package has several functions for importing, parsing, and analyzing pedigree data.
+Pedigrees define familial relationships in a hierarchical structure.
+Many genomics tools for working with pedigrees start with a .fam file, which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
+The skater package contains the function \texttt{read\_fam()} to read in a PLINK-formatted .fam file and another function \texttt{fam2ped()} to convert the content into a pedigree object as a nested tibble with one row per family.
+All pedigree processing from skater internally leverages a data structure from the kinship2 package \citep{sinnwell2014}.
+Further functions such as \texttt{plot\_pedigree()} produce a multi-page PDF drawing a diagram of the pedigree for each family, while \texttt{ped2kinpair()} produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair (see skater package vignette).
+
+\hypertarget{relationship-degree-inference-and-benchmarking}{%
+\subsection{Relationship degree inference and benchmarking}\label{relationship-degree-inference-and-benchmarking}}
+
+The skater package includes functions to translate kinship coefficients to relationship degrees. The kinship coefficients could come from \texttt{ped2kinpair()} or other kinship estimation software.
+
+The \texttt{dibble()} function creates a \textbf{d}egree \textbf{i}nference t\textbf{ibble}, with degrees up to the specified maximum degree resolution, expected kinship coefficient, and lower and upper inference ranges as defined in \citet{manichaikul2010}.
+The \texttt{kin2degree()} function infers the relationship degree given a kinship coefficient and a maximum degree resolution (e.g., 7th-degree relatives) up to which anything more distant is classified as unrelated.
+
+Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to known degrees of relatedness.
+When aggregated over many relationships and inferences, this can help benchmark performance of a particular kinship analysis method.
+The skater package adapts a \texttt{confusion\_matrix()} function from \citet{clark2021} to provide standard contingency table metrics (e.g.~sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
+The R-RMSE metric is defined more thoroughly in the skater package vignette and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation.
+In many kinship benchmarking analyses, classification error is treated in a categorical manner (exact match plus or minus one degree), neglecting the true amount of sharing as a real number.
+Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second-degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).
+
+\hypertarget{ibd-segment-analysis}{%
+\subsection{IBD segment analysis}\label{ibd-segment-analysis}}
+
+Tools such as hap-IBD \citep{zhou2020}, and IBIS \citep{seidman2020} detect shared IBD segments between individuals.
+The skater package includes functionality to take those IBD segments, compute shared genomic centimorgan (cM) length, and converts that shared cM to a kinship coefficient.
+In addition to inferred segments, these functions can estimate ``truth'' kinship from simulated IBD segments \citep{caballero2019}.
+The \texttt{read\_ibd()} function reads pairwise IBD segments from IBD inference tools and from simulated IBD segments.
+The \texttt{read\_map()} function reads in genetic map in a standard format which is required to translate the total centimorgans shared IBD to a kinship coefficient using the \texttt{ibd2kin()} function.
+
+\hypertarget{summary}{%
+\section{Summary}\label{summary}}
+
+The skater R package provides a robust software package for data import, manipulation, and analysis tasks typically encountered when working with SNP-based kinship analysis tools.
+All package functions are internally documented with examples, and the package contains a vignette demonstrating usage, inputs, outputs, and interpretation of all key functions.
+The package contains internal tests that are automatically run with continuous integration via GitHub Actions whenever the package code is updated.
+The skater package is permissively licensed (MIT) and is easily extensible to accommodate outputs from new genome-wide relatedness and IBD segment methods as they become available.
+
+\hypertarget{software-availability}{%
+\section{Software availability}\label{software-availability}}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Software available from: \url{http://CRAN.R-project.org/package=skater}.
+\item
+  Source code available from: \url{https://github.com/signaturescience/skater}.
+\item
+  Archived source code at time of publication: \textbf{FIXME Zenodo DOI to come here}.
+\item
+  Software license: MIT License.
+\end{enumerate}
+
+\hypertarget{author-information}{%
+\section{Author information}\label{author-information}}
+
+SDT and VPN developed the R package.
+
+All authors contributed to method development.
+
+SDT wrote the first draft of the manuscript.
+
+All authors assisted with manuscript revision.
+
+All authors read and approved the final manuscript.
+
+\hypertarget{competing-interests}{%
+\section{Competing interests}\label{competing-interests}}
+
+No competing interests were disclosed.
+
+\hypertarget{grant-information}{%
+\section{Grant information}\label{grant-information}}
+
+This work was supported in part by award 2019-DU-BX-0046 (Dense DNA Data for Enhanced Missing Persons Identification) to B.B., awarded by the National Institute of Justice, Office of Justice Programs, U.S.
+Department of Justice and by internal funds from the Center for Human Identification.
+The opinions, findings, and conclusions or recommendations expressed are those of the authors and do not necessarily reflect those of the U.S.
+Department of Justice.
+
+{\small\bibliography{bibliography.bib}}
 

### 2021-11-29 R/confusion_matrix.R
remove unicode characters causing latex errors when creating pdf version

@@ -97,7 +97,7 @@ calc_accuracy <- function(tabble) {
 #' \deqn{False Positive Rate = 1 - Specificity}
 #' \deqn{False Negative Rate = 1 - Sensitivity}
 #' \deqn{D' = qnorm(Sensitivity) - qnorm(1 - Specificity)}
-#' \deqn{AUC ≈ pnorm(D'/sqrt(2))}
+#' \deqn{AUC ~= pnorm(D'/sqrt(2))}
 #'
 #' See the references for discussions of the first five formulas.
 #' Abbreviations:
@@ -115,8 +115,8 @@ calc_accuracy <- function(tabble) {
 #'   \item{Sensitivity: True Positive Rate, Recall, Hit Rate, Power}{}
 #'   \item{Specificity: True Negative Rate}{}
 #'   \item{Positive Predictive Value: Precision}{}
-#'   \item{False Negative Rate: Miss Rate, Type II error rate, β}{}
-#'   \item{False Positive Rate: Fallout, Type I error rate, α}{}
+#'   \item{False Negative Rate: Miss Rate, Type II error rate, beta}{}
+#'   \item{False Positive Rate: Fallout, Type I error rate, alpha}{}
 #' }
 #'
 #' This function is called by \code{confusion_matrix}, but if this is all you

### 2021-11-29 R/read.R
remove warning about doi formatting

@@ -330,7 +330,7 @@ read_ibd <- function(file, source) {
 #' @references <http://zzz.bwh.harvard.edu/plink/data.shtml#map>
 #' @references <http://bochet.gcc.biostat.washington.edu/beagle/genetic_maps/>
 #' @references <https://github.com/williamslab/ped-sim#map-file>
-#' @references <http://dx.doi.org/10.1038/ncomms14994>
+#' @references <https://www.nature.com/articles/ncomms14994>
 #' @references <https://www.nature.com/articles/ncomms14994>
 #' @references <https://github.com/cbherer/Bherer_etal_SexualDimorphismRecombination>
 #'

### 2021-11-29 R/confusion_matrix.R
fix redirecting URL

@@ -127,7 +127,7 @@ calc_accuracy <- function(tabble) {
 #'
 #' @references Kuhn, M. (2008), "Building predictive models in R using the
 #' caret package, " \emph{Journal of Statistical Software},
-#' (\url{https://www.jstatsoft.org/article/view/v028i05/v28i05.pdf}).
+#' (\url{https://www.jstatsoft.org/article/view/v028i05}).
 #'
 #' Altman, D.G., Bland, J.M. (1994) "Diagnostic tests 1: sensitivity and
 #' specificity", \emph{British Medical Journal}, vol 308, 1552.

### 2021-11-29 R/skater-package.R
remove lifecycle dependency

@@ -7,9 +7,6 @@
 ## usethis namespace: end
 NULL
 
-# Suppress R CMD check note
-#' @importFrom lifecycle badge
-
 # Quiet CRAN check notes for unbound global variables
 if(getRversion() >= "2.15.1")  utils::globalVariables(c(".data",
                                                         "data",

### 2021-11-29 R/utils.R

@@ -80,7 +80,6 @@ kin2degree <- function(k, max_degree=3L) {
 #' Kinship coefficient to cM
 #'
 #' @description
-#' `r lifecycle::badge("experimental")`
 #'
 #' "Converts" a kinship coefficient to put on the same scale as shared cM using the formula
 #' \eqn{cm <- pmin(3560, 4*pmax(0, k)*3560)}.

### 2021-11-29 paper/paper.Rmd
updated first draft of f1000research paper

@@
 
+The `skater` package provides a collection of analysis and utility functions for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation as an **R** package.
+Functions in the package include tools for working with pedigree data, performing relationship degree inference, assessing classification accuracy, and summarizing IBD segment data.
+
+```{r}
+library(skater)
+```
+
 <!-- Optional - only if NO new datasets are included -->
@@
 
-The skater package has several functions for importing, parsing, and analyzing pedigree data.
+<!-- The commented line below was used in the shorter JOSS/Bioinformatics submission, expanded here -->
+<!-- The skater package has several functions for importing, parsing, and analyzing pedigree data.
 Pedigrees define familial relationships in a hierarchical structure.
@@
 Further functions such as `plot_pedigree()` produce a multi-page PDF drawing a diagram of the pedigree for each family, while `ped2kinpair()` produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair (see skater package vignette).
+-->
+
+Pedigrees define familial relationships in a hierarchical structure.
+One of the common formats used by PLINK [@purcell2007] and other genetic analysis tools is the `.fam` file.
+A `.fam` file is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
+The package includes `read_fam()` to read files in this format:
+
+```{r}
+famfile <- system.file("extdata", "3gens.fam", package="skater", mustWork=TRUE)
+fam <- read_fam(famfile)
+fam
+```
+
+Family structures imported from `.fam` formated files can then be translated to the `pedigree` structure used by the `kinship2` package [@sinnwell2014].
+The "fam" format may include multiple families, and the `fam2ped()` function will collapse them all into a `tibble` with one row per family:
+
+```{r}
+peds <- fam2ped(fam)
+peds
+```
+
+In the example above, the resulting `tibble` is nested by family ID.
+The `data` column contains the individual family information, while the `ped` column contains the pedigree object for that family.
+Using standard tidyverse operations, the resulting tibble can be unnested for any particular family:
+
+```{r}
+peds %>% 
+  dplyr::filter(fid=="testped1") %>% 
+  tidyr::unnest(cols=data)
+```
+
+A single pedigree can also be inspected or visualized (standard base R plot arguments such as `mar` or `cex` can be used to adjust aesthetics):
+
+```{r}
+peds$ped[[1]]
+```
+
+```{r plotped, fig.width=6, fig.height=3}
+plot(peds$ped[[1]], mar=c(1,4,1,4), cex=.7)
+```
+
+The `plot_pedigree()` function from `skater` will iterate over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from family:
+
+```{r, eval=FALSE}
+plot_pedigree(peds$ped, file="3gens.ped.pdf")
+```
+
+The `ped2kinpair()` function takes a pedigree object and produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair.
+
+The function can be run on a single family:
+
+```{r}
+ped2kinpair(peds$ped[[1]])
+```
+
+This function can also be mapped over all families in the pedigree:
+
+```{r}
+kinpairs <- 
+  peds %>% 
+  dplyr::mutate(pairs=purrr::map(ped, ped2kinpair)) %>% 
+  dplyr::select(fid, pairs) %>% 
+  tidyr::unnest(cols=pairs)
+kinpairs
+```
+
+Note that this maps `ped2kinpair()` over all `ped` objects in the input `tibble`, and that relationships are not shown for between-family relationships.
 
@@
 
-The `dibble()` function creates a **d**egree **i**nference t**ibble**, with degrees up to the specified maximum degree resolution, expected kinship coefficient, and lower and upper inference ranges as defined in @manichaikul2010.
+<!-- The commented line below was used in the shorter JOSS/Bioinformatics submission, expanded here -->
+<!-- The `dibble()` function creates a **d**egree **i**nference t**ibble**, with degrees up to the specified maximum degree resolution, expected kinship coefficient, and lower and upper inference ranges as defined in @manichaikul2010.
 The `kin2degree()` function infers the relationship degree given a kinship coefficient and a maximum degree resolution (e.g., 7th-degree relatives) up to which anything more distant is classified as unrelated.
+-->
 
-Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to known degrees of relatedness.
+The `dibble()` function creates a **d**egree **i**nference `tibble`, with degrees up to the specified `max_degree` (default=3), expected kinship coefficient, and lower (`l`) and upper (`u`) inference ranges as defined in @manichaikul2010.
+Degree 0 corresponds to self / identity / monozygotic twins, with an expected kinship coefficient of 0.5, with inference range >=0.354.
+Anything beyond the maximum degree resolution is considered unrelated (degree `NA`).
+
+```{r}
+dibble()
+```
+
+The degree inference `max_degree` default is 3. Change this argument to allow more granular degree inference ranges:
+
+```{r}
+dibble(max_degree = 5)
+```
+
+Note that the distance between relationship degrees becomes smaller as the relationship degree becomes more distant.
+The `dibble()` function will emit a warning with `max_degree` >=10, and will stop with an error at >=12.
+
+The `kin2degree()` function infers the relationship degree given a kinship coefficient and a `max_degree` up to which anything more distant is treated as unrelated. Example first degree relative:
+
+```{r}
+kin2degree(.25, max_degree=3)
+```
+
+Example 4th degree relative, but using the default max_degree resolution of 3:
+
+```{r}
+kin2degree(.0312, max_degree=3)
+```
+
+Example 4th degree relative, but increasing the degree resolution:
+
+```{r}
+kin2degree(.0312, max_degree=5)
+```
+
+The `kin2degree()` function is vectorized over values of `k`, so it can be used inside of a `mutate` on a `tibble` of kinship coefficients:
+
+```{r}
+# Get two pairs from each type of relationship we have in kinpairs:
+kinpairs_subset <- 
+  kinpairs %>% 
+  dplyr::group_by(k) %>% 
+  dplyr::slice(1:2)
+kinpairs_subset
+
+# Infer degree out to third degree relatives:
+kinpairs_subset %>% 
+  dplyr::mutate(degree=kin2degree(k, max_degree=3))
+```
+
+## Benchmarking Degree Classification
+
+<!-- The commented line below was used in the shorter JOSS/Bioinformatics submission, expanded here -->
+<!-- Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to known degrees of relatedness.
 When aggregated over many relationships and inferences, this can help benchmark performance of a particular kinship analysis method.
@@
 Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second-degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).
+-->
+
+Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to truth.
+When aggregated over many relationships and inferences, this method can help benchmark performance of a particular kinship analysis method.
+
+The skater package adapts a `confusion_matrix()` function from @clark2021 to provide standard contingency table metrics (e.g.
+sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
+The `confusion_matrix()` function on its own outputs a list with four objects:
+
+1. A `tibble` with calculated accuracy, lower and upper bounds, the guessing rate and p-value of the accuracy vs. the guessing rate. 
+2. A `tibble` with contingency table statistics calculated for each class. Details on the statistics calculated for each class can be reviewed on the help page for `?confusion_matrix`.
+3. A `matrix` with the contingency table object itself.
+4.
+A `vector` with the reciprocal RMSE (R-RMSE).
+The R-RMSE is calculated as `sqrt(mean((1/(Target+.5)-1/(Predicted+.5))^2)))`, and represents an alternative to classification accuracy when benchmarking relationship degree estimation.
+Taking the reciprocal of the target and predicted degree results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree).
+The +0.5 adjustment prevents division-by-zero when a 0th-degree (identical) relative pair is introduced.
+
+To illustrate the usage, this example will start with the `kinpairs` data from above and randomly flip ~20% of the true relationship degrees.
+
+```{r}
+# Function to randomly flip levels of a factor (at 20%, by default)
+randomflip <- function(x, p=.2) ifelse(runif(length(x))<p, sample(unique(x)), x)
+
+# Infer degree (truth/target) using kin2degree, then randomly flip 20% of them
+set.seed(42)
+kinpairs_inferred <- kinpairs %>% 
+  dplyr::mutate(degree_truth=kin2degree(k, max_degree=3)) %>% 
+  dplyr::mutate(degree_truth=tidyr::replace_na(degree_truth, "unrelated")) %>% 
+  dplyr::mutate(degree_inferred=randomflip(degree_truth))
+kinpairs_inferred
+```
+
+Next, running the `confusion_matrix()` function will return all four objects noted above.
+
+```{r}
+confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
+                 target = kinpairs_inferred$degree_truth)
+```
+
+Standard tidyverse functions such as `purrr::pluck()` can be used to isolate just the contingency table:
+
+```{r}
+confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
+                 target = kinpairs_inferred$degree_truth) %>% 
+  purrr::pluck("Table")
+```
+
+The `confusion_matrix()` function includes an argument to output in a tidy (`longer=TRUE`) format, and the example below illustrates how to spread contingency table statistics by class:
+
+```{r}
+confusion_matrix(prediction = kinpairs_inferred$degree_inferred, 
+                 target = kinpairs_inferred$degree_truth, 
+                 longer = TRUE) %>% 
+  purrr::pluck("Other") %>% 
+  tidyr::spread(Class, Value) %>% 
+  dplyr::relocate(Average, .after=dplyr::last_col()) %>% 
+  dplyr::mutate_if(rlang::is_double, signif, 2)
+```
 
@@
 The `read_map()` function reads in genetic map in a standard format which is required to translate the total centimorgans shared IBD to a kinship coefficient using the `ibd2kin()` function.
+See `?read_ibd` and `?read_map` for additional details on expected format.
+
+The `read_ibd()` function reads in the pairwise IBD segment format.
+Input to this function can either be inferred IBD segments from hap-IBD (`source="hapibd"`) or simulated segments (`source="pedsim"`).
+The first example below uses data in the `hap-ibd` output format:
+
+```{r, message=FALSE}
+hapibd_filepath <- system.file("extdata", "GBR.sim.ibd.gz", 
+                               package="skater")
+hapibd_seg <- read_ibd(hapibd_filepath, source = "hapibd")
+hapibd_seg
+```
+
+In order to translate the shared genomic cM length to a kinship coefficient, a genetic map must first be read in with `read_map()`.
+Software for IBD segment inference and simulation requires a genetic map.
+The map loaded for kinship estimation should be the same one used for creating the shared IBD segment output.
+The example below uses a minimal genetic map that ships with `skater`:
+
+```{r}
+gmap_filepath <- system.file("extdata", "sexspec-avg-min.plink.map", 
+                             package="skater")
+gmap <- read_map(gmap_filepath)
+gmap
+```
+
+The `ibd2kin()` function takes the segments and map file and outputs a `tibble` with one row per pair of individuals and columns for individual 1 ID, individual 2 ID, and the kinship coefficient for the pair:
+
+```{r}
+ibd_dat <- ibd2kin(.ibd_data=hapibd_seg, .map=gmap)
+ibd_dat
+```
 

### 2021-11-29 paper/paper.tex
updated first draft of f1000research paper

@@
 
+The \texttt{skater} package provides a collection of analysis and utility functions for \textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation as an \textbf{R} package.
+Functions in the package include tools for working with pedigree data, performing relationship degree inference, assessing classification accuracy, and summarizing IBD segment data.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{library}\NormalTok{(skater)}
+\end{Highlighting}
+\end{Shaded}
+
 \hypertarget{pedigree-parsing-manipulation-and-analysis}{%
@@
 
-The skater package has several functions for importing, parsing, and analyzing pedigree data.
 Pedigrees define familial relationships in a hierarchical structure.
-Many genomics tools for working with pedigrees start with a .fam file, which is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
-The skater package contains the function \texttt{read\_fam()} to read in a PLINK-formatted .fam file and another function \texttt{fam2ped()} to convert the content into a pedigree object as a nested tibble with one row per family.
-All pedigree processing from skater internally leverages a data structure from the kinship2 package \citep{sinnwell2014}.
-Further functions such as \texttt{plot\_pedigree()} produce a multi-page PDF drawing a diagram of the pedigree for each family, while \texttt{ped2kinpair()} produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair (see skater package vignette).
+One of the common formats used by PLINK \citep{purcell2007} and other genetic analysis tools is the \texttt{.fam} file.
+A \texttt{.fam} file is a tabular format with one row per individual and columns for unique IDs of the mother, father, and the family unit.
+The package includes \texttt{read\_fam()} to read files in this format:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{famfile }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"3gens.fam"}\NormalTok{, }\AttributeTok{package=}\StringTok{"skater"}\NormalTok{, }\AttributeTok{mustWork=}\ConstantTok{TRUE}\NormalTok{)}
+\NormalTok{fam }\OtherTok{\textless{}{-}} \FunctionTok{read\_fam}\NormalTok{(famfile)}
+\NormalTok{fam}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 64 x 6
+##    fid      id                dadid             momid               sex affected
+##    <chr>    <chr>             <chr>             <chr>             <int>    <int>
+##  1 testped1 testped1_g1-b1-s1 0                 0                     1        1
+##  2 testped1 testped1_g1-b1-i1 0                 0                     2        1
+##  3 testped1 testped1_g2-b1-s1 0                 0                     1        1
+##  4 testped1 testped1_g2-b1-i1 testped1_g1-b1-s1 testped1_g1-b1-i1     2        1
+##  5 testped1 testped1_g2-b2-s1 0                 0                     1        1
+##  6 testped1 testped1_g2-b2-i1 testped1_g1-b1-s1 testped1_g1-b1-i1     2        1
+##  7 testped1 testped1_g3-b1-i1 testped1_g2-b1-s1 testped1_g2-b1-i1     2        1
+##  8 testped1 testped1_g3-b2-i1 testped1_g2-b2-s1 testped1_g2-b2-i1     1        1
+##  9 testped2 testped2_g1-b1-s1 0                 0                     2        1
+## 10 testped2 testped2_g1-b1-i1 0                 0                     1        1
+## # ... with 54 more rows
+\end{verbatim}
+
+Family structures imported from \texttt{.fam} formated files can then be translated to the \texttt{pedigree} structure used by the \texttt{kinship2} package \citep{sinnwell2014}.
+The ``fam'' format may include multiple families, and the \texttt{fam2ped()} function will collapse them all into a \texttt{tibble} with one row per family:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{peds }\OtherTok{\textless{}{-}} \FunctionTok{fam2ped}\NormalTok{(fam)}
+\NormalTok{peds}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 8 x 3
+##   fid      data             ped       
+##   <chr>    <list>           <list>    
+## 1 testped1 <tibble [8 x 5]> <pedigree>
+## 2 testped2 <tibble [8 x 5]> <pedigree>
+## 3 testped3 <tibble [8 x 5]> <pedigree>
+## 4 testped4 <tibble [8 x 5]> <pedigree>
+## 5 testped5 <tibble [8 x 5]> <pedigree>
+## 6 testped6 <tibble [8 x 5]> <pedigree>
+## 7 testped7 <tibble [8 x 5]> <pedigree>
+## 8 testped8 <tibble [8 x 5]> <pedigree>
+\end{verbatim}
+
+In the example above, the resulting \texttt{tibble} is nested by family ID.
+The \texttt{data} column contains the individual family information, while the \texttt{ped} column contains the pedigree object for that family.
+Using standard tidyverse operations, the resulting tibble can be unnested for any particular family:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{peds }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(fid}\SpecialCharTok{==}\StringTok{"testped1"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols=}\NormalTok{data)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 8 x 7
+##   fid      id                dadid             momid        sex affected ped    
+##   <chr>    <chr>             <chr>             <chr>      <int>    <dbl> <list> 
+## 1 testped1 testped1_g1-b1-s1 <NA>              <NA>           1        1 <pedig~
+## 2 testped1 testped1_g1-b1-i1 <NA>              <NA>           2        1 <pedig~
+## 3 testped1 testped1_g2-b1-s1 <NA>              <NA>           1        1 <pedig~
+## 4 testped1 testped1_g2-b1-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
+## 5 testped1 testped1_g2-b2-s1 <NA>              <NA>           1        1 <pedig~
+## 6 testped1 testped1_g2-b2-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
+## 7 testped1 testped1_g3-b1-i1 testped1_g2-b1-s1 testped1_~     2        1 <pedig~
+## 8 testped1 testped1_g3-b2-i1 testped1_g2-b2-s1 testped1_~     1        1 <pedig~
+\end{verbatim}
+
+A single pedigree can also be inspected or visualized (standard base R plot arguments such as \texttt{mar} or \texttt{cex} can be used to adjust aesthetics):
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]]}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## Pedigree object with 8 subjects
+## Bit size= 4
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plot}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]], }\AttributeTok{mar=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{), }\AttributeTok{cex=}\NormalTok{.}\DecValTok{7}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{center}\includegraphics{paper_files/figure-latex/plotped-1} \end{center}
+
+The \texttt{plot\_pedigree()} function from \texttt{skater} will iterate over a list of pedigree objects, writing a multi-page PDF, with each page containing a pedigree from family:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{plot\_pedigree}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped, }\AttributeTok{file=}\StringTok{"3gens.ped.pdf"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+The \texttt{ped2kinpair()} function takes a pedigree object and produces a pairwise list of relationships between all individuals in the data with the expected kinship coefficients for each pair.
+
+The function can be run on a single family:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{ped2kinpair}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]])}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 36 x 3
+##    id1               id2                   k
+##    <chr>             <chr>             <dbl>
+##  1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5  
+##  2 testped1_g1-b1-i1 testped1_g1-b1-s1 0    
+##  3 testped1_g1-b1-s1 testped1_g2-b1-s1 0    
+##  4 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25 
+##  5 testped1_g1-b1-s1 testped1_g2-b2-s1 0    
+##  6 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25 
+##  7 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125
+##  8 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125
+##  9 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5  
+## 10 testped1_g1-b1-i1 testped1_g2-b1-s1 0    
+## # ... with 26 more rows
+\end{verbatim}
+
+This function can also be mapped over all families in the pedigree:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{kinpairs }\OtherTok{\textless{}{-}} 
+\NormalTok{  peds }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{pairs=}\NormalTok{purrr}\SpecialCharTok{::}\FunctionTok{map}\NormalTok{(ped, ped2kinpair)) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(fid, pairs) }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols=}\NormalTok{pairs)}
+\NormalTok{kinpairs}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 288 x 4
+##    fid      id1               id2                   k
+##    <chr>    <chr>             <chr>             <dbl>
+##  1 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5  
+##  2 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0    
+##  3 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0    
+##  4 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25 
+##  5 testped1 testped1_g1-b1-s1 testped1_g2-b2-s1 0    
+##  6 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25 
+##  7 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125
+##  8 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125
+##  9 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5  
+## 10 testped1 testped1_g1-b1-i1 testped1_g2-b1-s1 0    
+## # ... with 278 more rows
+\end{verbatim}
+
+Note that this maps \texttt{ped2kinpair()} over all \texttt{ped} objects in the input \texttt{tibble}, and that relationships are not shown for between-family relationships.
 
@@
 
-The \texttt{dibble()} function creates a \textbf{d}egree \textbf{i}nference t\textbf{ibble}, with degrees up to the specified maximum degree resolution, expected kinship coefficient, and lower and upper inference ranges as defined in \citet{manichaikul2010}.
-The \texttt{kin2degree()} function infers the relationship degree given a kinship coefficient and a maximum degree resolution (e.g., 7th-degree relatives) up to which anything more distant is classified as unrelated.
+The \texttt{dibble()} function creates a \textbf{d}egree \textbf{i}nference \texttt{tibble}, with degrees up to the specified \texttt{max\_degree} (default=3), expected kinship coefficient, and lower (\texttt{l}) and upper (\texttt{u}) inference ranges as defined in \citet{manichaikul2010}.
+Degree 0 corresponds to self / identity / monozygotic twins, with an expected kinship coefficient of 0.5, with inference range \textgreater=0.354.
+Anything beyond the maximum degree resolution is considered unrelated (degree \texttt{NA}).
 
-Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to known degrees of relatedness.
-When aggregated over many relationships and inferences, this can help benchmark performance of a particular kinship analysis method.
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{dibble}\NormalTok{()}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 5 x 4
+##   degree      k       l      u
+##    <int>  <dbl>   <dbl>  <dbl>
+## 1      0 0.5     0.354  1     
+## 2      1 0.25    0.177  0.354 
+## 3      2 0.125   0.0884 0.177 
+## 4      3 0.0625  0.0442 0.0884
+## 5     NA 0      -1      0.0442
+\end{verbatim}
+
+The degree inference \texttt{max\_degree} default is 3. Change this argument to allow more granular degree inference ranges:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{dibble}\NormalTok{(}\AttributeTok{max\_degree =} \DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 7 x 4
+##   degree      k       l      u
+##    <int>  <dbl>   <dbl>  <dbl>
+## 1      0 0.5     0.354  1     
+## 2      1 0.25    0.177  0.354 
+## 3      2 0.125   0.0884 0.177 
+## 4      3 0.0625  0.0442 0.0884
+## 5      4 0.0312  0.0221 0.0442
+## 6      5 0.0156  0.0110 0.0221
+## 7     NA 0      -1      0.0110
+\end{verbatim}
+
+Note that the distance between relationship degrees becomes smaller as the relationship degree becomes more distant.
+The \texttt{dibble()} function will emit a warning with \texttt{max\_degree} \textgreater=10, and will stop with an error at \textgreater=12.
+
+The \texttt{kin2degree()} function infers the relationship degree given a kinship coefficient and a \texttt{max\_degree} up to which anything more distant is treated as unrelated.
+Example first degree relative:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{25}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 1
+\end{verbatim}
+
+Example 4th degree relative, but using the default max\_degree resolution of 3:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] NA
+\end{verbatim}
+
+Example 4th degree relative, but increasing the degree resolution:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{5}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## [1] 4
+\end{verbatim}
+
+The \texttt{kin2degree()} function is vectorized over values of \texttt{k}, so it can be used inside of a \texttt{mutate} on a \texttt{tibble} of kinship coefficients:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Get two pairs from each type of relationship we have in kinpairs:}
+\NormalTok{kinpairs\_subset }\OtherTok{\textless{}{-}} 
+\NormalTok{  kinpairs }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{group\_by}\NormalTok{(k) }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{slice}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{2}\NormalTok{)}
+\NormalTok{kinpairs\_subset}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 10 x 4
+## # Groups:   k [5]
+##    fid      id1               id2                    k
+##    <chr>    <chr>             <chr>              <dbl>
+##  1 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0     
+##  2 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0     
+##  3 testped1 testped1_g3-b1-i1 testped1_g3-b2-i1 0.0625
+##  4 testped2 testped2_g3-b1-i1 testped2_g3-b2-i1 0.0625
+##  5 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125 
+##  6 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125 
+##  7 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25  
+##  8 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25  
+##  9 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5   
+## 10 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Infer degree out to third degree relatives:}
+\NormalTok{kinpairs\_subset }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree=}\FunctionTok{kin2degree}\NormalTok{(k, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{))}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 10 x 5
+## # Groups:   k [5]
+##    fid      id1               id2                    k degree
+##    <chr>    <chr>             <chr>              <dbl>  <int>
+##  1 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0          NA
+##  2 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0          NA
+##  3 testped1 testped1_g3-b1-i1 testped1_g3-b2-i1 0.0625      3
+##  4 testped2 testped2_g3-b1-i1 testped2_g3-b2-i1 0.0625      3
+##  5 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125       2
+##  6 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125       2
+##  7 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25        1
+##  8 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25        1
+##  9 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5         0
+## 10 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5         0
+\end{verbatim}
+
+\hypertarget{benchmarking-degree-classification}{%
+\subsection{Benchmarking Degree Classification}\label{benchmarking-degree-classification}}
+
+Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to truth.
+When aggregated over many relationships and inferences, this method can help benchmark performance of a particular kinship analysis method.
+
 The skater package adapts a \texttt{confusion\_matrix()} function from \citet{clark2021} to provide standard contingency table metrics (e.g.~sensitivity, specificity, PPV, precision, recall, F1, etc.) with a new reciprocal RMSE (R-RMSE) metric.
-The R-RMSE metric is defined more thoroughly in the skater package vignette and may be a preferable measure of classification accuracy when benchmarking relationship degree estimation.
-In many kinship benchmarking analyses, classification error is treated in a categorical manner (exact match plus or minus one degree), neglecting the true amount of sharing as a real number.
-Taking the reciprocal of the target and predicted degree in a typical RMSE calculation results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second-degree) than misclassifications at more distant relationships (e.g., classifying a fourth-degree relative pair as fifth-degree).
+The \texttt{confusion\_matrix()} function on its own outputs a list with four objects:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  A \texttt{tibble} with calculated accuracy, lower and upper bounds, the guessing rate and p-value of the accuracy vs.~the guessing rate.
+\item
+  A \texttt{tibble} with contingency table statistics calculated for each class. Details on the statistics calculated for each class can be reviewed on the help page for \texttt{?confusion\_matrix}.
+\item
+  A \texttt{matrix} with the contingency table object itself.
+\item
+  A \texttt{vector} with the reciprocal RMSE (R-RMSE).
+The R-RMSE is calculated as \texttt{sqrt(mean((1/(Target+.5)-1/(Predicted+.5))\^{}2)))}, and represents an alternative to classification accuracy when benchmarking relationship degree estimation.
+Taking the reciprocal of the target and predicted degree results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree).
+The +0.5 adjustment prevents division-by-zero when a 0th-degree (identical) relative pair is introduced.
+\end{enumerate}
+
+To illustrate the usage, this example will start with the \texttt{kinpairs} data from above and randomly flip \textasciitilde20\% of the true relationship degrees.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\CommentTok{\# Function to randomly flip levels of a factor (at 20\%, by default)}
+\NormalTok{randomflip }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x, }\AttributeTok{p=}\NormalTok{.}\DecValTok{2}\NormalTok{) }\FunctionTok{ifelse}\NormalTok{(}\FunctionTok{runif}\NormalTok{(}\FunctionTok{length}\NormalTok{(x))}\SpecialCharTok{\textless{}}\NormalTok{p, }\FunctionTok{sample}\NormalTok{(}\FunctionTok{unique}\NormalTok{(x)), x)}
+
+\CommentTok{\# Infer degree (truth/target) using kin2degree, then randomly flip 20\% of them}
+\FunctionTok{set.seed}\NormalTok{(}\DecValTok{42}\NormalTok{)}
+\NormalTok{kinpairs\_inferred }\OtherTok{\textless{}{-}}\NormalTok{ kinpairs }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_truth=}\FunctionTok{kin2degree}\NormalTok{(k, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_truth=}\NormalTok{tidyr}\SpecialCharTok{::}\FunctionTok{replace\_na}\NormalTok{(degree\_truth, }\StringTok{"unrelated"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_inferred=}\FunctionTok{randomflip}\NormalTok{(degree\_truth))}
+\NormalTok{kinpairs\_inferred}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 288 x 6
+##    fid      id1               id2                   k degree_truth degree_inferred
+##    <chr>    <chr>             <chr>             <dbl> <chr>        <chr>          
+##  1 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5   0            0              
+##  2 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0     unrelated    unrelated      
+##  3 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0     unrelated    unrelated      
+##  4 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25  1            1              
+##  5 testped1 testped1_g1-b1-s1 testped1_g2-b2-s1 0     unrelated    unrelated      
+##  6 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25  1            1              
+##  7 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125 2            2              
+##  8 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125 2            1              
+##  9 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5   0            0              
+## 10 testped1 testped1_g1-b1-i1 testped1_g2-b1-s1 0     unrelated    unrelated      
+## # ... with 278 more rows
+\end{verbatim}
+
+Next, running the \texttt{confusion\_matrix()} function will return all four objects noted above.
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
+                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## $Accuracy
+## # A tibble: 1 x 5
+##   Accuracy `Accuracy LL` `Accuracy UL` `Accuracy Guessing` `Accuracy P-value`
+##      <dbl>         <dbl>         <dbl>               <dbl>              <dbl>
+## 1    0.812         0.763         0.856               0.333           1.09e-62
+## 
+## $Other
+## # A tibble: 6 x 15
+##   Class         N `Sensitivity/Re~ `Specificity/TN~ `PPV/Precision`   NPV `F1/Dice`
+##   <chr>     <dbl>            <dbl>            <dbl>           <dbl> <dbl>     <dbl>
+## 1 0          64              0.75             0.964           0.857 0.931     0.8  
+## 2 1          72              0.806            0.944           0.829 0.936     0.817
+## 3 2          48              0.833            0.967           0.833 0.967     0.833
+## 4 3           8              0.75             0.936           0.25  0.992     0.375
+## 5 unrelated  96              0.854            0.958           0.911 0.929     0.882
+## 6 Average    57.6            0.799            0.954           0.736 0.951     0.741
+## # ... with 8 more variables: Prevalence <dbl>, Detection Rate <dbl>,
+## #   Detection Prevalence <dbl>, Balanced Accuracy <dbl>, FDR <dbl>, FOR <dbl>,
+## #   FPR/Fallout <dbl>, FNR <dbl>
+## 
+## $Table
+##            Target
+## Predicted    0  1  2  3 unrelated
+##   0         48  4  2  1         1
+##   1          5 58  4  0         3
+##   2          0  3 40  1         4
+##   3          8  4  0  6         6
+##   unrelated  3  3  2  0        82
+## 
+## $recip_rmse
+## [1] 0.4665971
+\end{verbatim}
+
+Standard tidyverse functions such as \texttt{purrr::pluck()} can be used to isolate just the contingency table:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
+                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth) }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  purrr}\SpecialCharTok{::}\FunctionTok{pluck}\NormalTok{(}\StringTok{"Table"}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+##            Target
+## Predicted    0  1  2  3 unrelated
+##   0         48  4  2  1         1
+##   1          5 58  4  0         3
+##   2          0  3 40  1         4
+##   3          8  4  0  6         6
+##   unrelated  3  3  2  0        82
+\end{verbatim}
+
+The \texttt{confusion\_matrix()} function includes an argument to output in a tidy (\texttt{longer=TRUE}) format, and the example below illustrates how to spread contingency table statistics by class:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
+                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth, }
+                 \AttributeTok{longer =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  purrr}\SpecialCharTok{::}\FunctionTok{pluck}\NormalTok{(}\StringTok{"Other"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{spread}\NormalTok{(Class, Value) }\SpecialCharTok{\%\textgreater{}\%} 
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{relocate}\NormalTok{(Average, }\AttributeTok{.after=}\NormalTok{dplyr}\SpecialCharTok{::}\FunctionTok{last\_col}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
+\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate\_if}\NormalTok{(rlang}\SpecialCharTok{::}\NormalTok{is\_double, signif, }\DecValTok{2}\NormalTok{)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 14 x 7
+##    Statistic                 `0`    `1`    `2`    `3` unrelated Average
+##    <chr>                   <dbl>  <dbl>  <dbl>  <dbl>     <dbl>   <dbl>
+##  1 Balanced Accuracy       0.86   0.88   0.9   0.84       0.91    0.88 
+##  2 Detection Prevalence    0.19   0.24   0.17  0.083      0.31    0.2  
+##  3 Detection Rate          0.17   0.2    0.14  0.021      0.28    0.16 
+##  4 F1/Dice                 0.8    0.82   0.83  0.38       0.88    0.74 
+##  5 FDR                     0.14   0.17   0.17  0.75       0.089   0.26 
+##  6 FNR                     0.25   0.19   0.17  0.25       0.15    0.2  
+##  7 FOR                     0.069  0.064  0.033 0.0076     0.071   0.049
+##  8 FPR/Fallout             0.036  0.056  0.033 0.064      0.042   0.046
+##  9 N                      64     72     48     8         96      58    
+## 10 NPV                     0.93   0.94   0.97  0.99       0.93    0.95 
+## 11 PPV/Precision           0.86   0.83   0.83  0.25       0.91    0.74 
+## 12 Prevalence              0.22   0.25   0.17  0.028      0.33    0.2  
+## 13 Sensitivity/Recall/TPR  0.75   0.81   0.83  0.75       0.85    0.8  
+## 14 Specificity/TNR         0.96   0.94   0.97  0.94       0.96    0.95
+\end{verbatim}
 
@@
 The \texttt{read\_map()} function reads in genetic map in a standard format which is required to translate the total centimorgans shared IBD to a kinship coefficient using the \texttt{ibd2kin()} function.
+See \texttt{?read\_ibd} and \texttt{?read\_map} for additional details on expected format.
+
+The \texttt{read\_ibd()} function reads in the pairwise IBD segment format.
+Input to this function can either be inferred IBD segments from hap-IBD (\texttt{source="hapibd"}) or simulated segments (\texttt{source="pedsim"}).
+The first example below uses data in the \texttt{hap-ibd} output format:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{hapibd\_filepath }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"GBR.sim.ibd.gz"}\NormalTok{, }
+                               \AttributeTok{package=}\StringTok{"skater"}\NormalTok{)}
+\NormalTok{hapibd\_seg }\OtherTok{\textless{}{-}} \FunctionTok{read\_ibd}\NormalTok{(hapibd\_filepath, }\AttributeTok{source =} \StringTok{"hapibd"}\NormalTok{)}
+\NormalTok{hapibd\_seg}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 3,954 x 6
+##    id1               id2                 chr     start       end length
+##    <chr>             <chr>             <dbl>     <dbl>     <dbl>  <dbl>
+##  1 testped1_g1-b1-s1 testped1_g3-b1-i1     1 197661576 234863602   47.1
+##  2 testped1_g2-b2-i1 testped1_g3-b1-i1     1 197661576 231017545   39.8
+##  3 testped1_g3-b1-i1 testped1_g3-b2-i1     1 197661576 212799139   20.3
+##  4 testped3_g1-b1-s1 testped3_g3-b2-i1     1   2352146  10862397   17.7
+##  5 testped3_g2-b2-i1 testped3_g3-b2-i1     1   2352146  10862397   17.7
+##  6 testped1_g1-b1-s1 testped1_g2-b1-i1     1   3328659  64123868   86.4
+##  7 testped1_g1-b1-s1 testped1_g3-b1-i1     1   3328659  33476811   51.2
+##  8 testped1_g2-b2-s1 testped1_g3-b2-i1     1   5003504  32315147   45.9
+##  9 testped2_g1-b1-i1 testped2_g3-b1-i1     1 240810528 248578622   15.9
+## 10 testped2_g1-b1-i1 testped2_g2-b2-i1     1 241186056 249170711   15.5
+## # ... with 3,944 more rows
+\end{verbatim}
+
+In order to translate the shared genomic cM length to a kinship coefficient, a genetic map must first be read in with \texttt{read\_map()}.
+Software for IBD segment inference and simulation requires a genetic map.
+The map loaded for kinship estimation should be the same one used for creating the shared IBD segment output.
+The example below uses a minimal genetic map that ships with \texttt{skater}:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{gmap\_filepath }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"sexspec{-}avg{-}min.plink.map"}\NormalTok{, }
+                             \AttributeTok{package=}\StringTok{"skater"}\NormalTok{)}
+\NormalTok{gmap }\OtherTok{\textless{}{-}} \FunctionTok{read\_map}\NormalTok{(gmap\_filepath)}
+\NormalTok{gmap}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 28,726 x 3
+##      chr  value      bp
+##    <dbl>  <dbl>   <dbl>
+##  1     1 0       752721
+##  2     1 0.0292 1066029
+##  3     1 0.0829 1099342
+##  4     1 0.157  1106473
+##  5     1 0.246  1152631
+##  6     1 0.294  1314015
+##  7     1 0.469  1510801
+##  8     1 0.991  1612540
+##  9     1 1.12   1892325
+## 10     1 1.41   1916587
+## # ... with 28,716 more rows
+\end{verbatim}
+
+The \texttt{ibd2kin()} function takes the segments and map file and outputs a \texttt{tibble} with one row per pair of individuals and columns for individual 1 ID, individual 2 ID, and the kinship coefficient for the pair:
+
+\begin{Shaded}
+\begin{Highlighting}[]
+\NormalTok{ibd\_dat }\OtherTok{\textless{}{-}} \FunctionTok{ibd2kin}\NormalTok{(}\AttributeTok{.ibd\_data=}\NormalTok{hapibd\_seg, }\AttributeTok{.map=}\NormalTok{gmap)}
+\NormalTok{ibd\_dat}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## # A tibble: 196 x 3
+##    id1               id2                kinship
+##    <chr>             <chr>                <dbl>
+##  1 testped1_g1-b1-i1 testped1_g1-b1-s1 0.000316
+##  2 testped1_g1-b1-i1 testped1_g2-b1-i1 0.261   
+##  3 testped1_g1-b1-i1 testped1_g2-b2-i1 0.263   
+##  4 testped1_g1-b1-i1 testped1_g2-b2-s1 0.000150
+##  5 testped1_g1-b1-i1 testped1_g3-b1-i1 0.145   
+##  6 testped1_g1-b1-i1 testped1_g3-b2-i1 0.133   
+##  7 testped1_g1-b1-i1 testped2_g1-b1-i1 0.000165
+##  8 testped1_g1-b1-i1 testped2_g1-b1-s1 0.000323
+##  9 testped1_g1-b1-i1 testped2_g2-b1-i1 0.000499
+## 10 testped1_g1-b1-i1 testped2_g2-b1-s1 0.000318
+## # ... with 186 more rows
+\end{verbatim}
 

### 2021-11-29 README.Rmd

@@
 <!-- badges: start -->
+
 [![R-CMD-check-stable](https://github.com/signaturescience/skater/workflows/R-CMD-check-stable/badge.svg)](https://github.com/signaturescience/skater/actions)
@@
 [![R-CMD-check-dev](https://github.com/signaturescience/skater/workflows/R-CMD-check-dev/badge.svg)](https://github.com/signaturescience/skater/actions)
+
 <!-- badges: end -->

### 2021-11-30 paper/paper.Rmd

@@
 ---
-title: "**skater**: An R package for SNP-based Kinship Analysis, Testing, and Evaluation"
+title: '**skater**: An R package for SNP-based Kinship Analysis, Testing, and Evaluation'
 author:
@@
 - name: Jianye Ge
-    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
+  affiliation: Center for Human Identification, Department of Microbiology, Immunology,
+    and Genetics, University of North Texas Health Science Center, Fort Worth, TX
+    76107, USA.
 - name: August E. Woerner
-    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
+  affiliation: Center for Human Identification, Department of Microbiology, Immunology,
+    and Genetics, University of North Texas Health Science Center, Fort Worth, TX
+    76107, USA.
 - name: Bruce Budowle
-    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
-# Abstract up to 300 words. No citations, minimal abbreviations.
-abstract: | 
-    | 
-    | **Motivation:** SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation.
-A dedicated software package that consistently and intuitively implements this analysis functionality is needed.
-    | **Results:**  Here we present the skater R package for **S**NP-based **k**inship **a**nalysis, **t**esting, and **e**valuation with **R**.
-The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.
-    | **Availability:** The skater package is implemented as an R package and is released under the MIT license at https\:\/\/github.com/signaturescience/skater.
-Documentation is available at https\:\/\/signaturescience.github.io/skater.
-keywords: bioinformatics, kinship, R, genealogy, SNPs, single nucleotide polymorphisms, relatedness
+  affiliation: Center for Human Identification, Department of Microbiology, Immunology,
+    and Genetics, University of North Texas Health Science Center, Fort Worth, TX
+    76107, USA.
+abstract: "| \n| **Motivation:** SNP-based kinship analysis with genome-wide relationship
+  estimation and IBD segment analysis methods produces results that often require
+  further downstream processing and manipulation. A dedicated software package that
+  consistently and intuitively implements this analysis functionality is needed.  \n|
+  **a**nalysis, **t**esting, and **e**valuation with **R**. The skater package contains
+  a suite of well-documented tools for importing, parsing, and analyzing pedigree
+  data, performing relationship degree inference, benchmarking relationship degree
+  package is implemented as an R package and is released under the MIT license at
+  **Results:**  Here we present the skater R package for **S**NP-based **k**inship
+  classification, and summarizing IBD segment data.  \n| **Availability:** The skater
+  https\\:\\/\\/github.com/signaturescience/skater. Documentation is available at
+  https\\:\\/\\/signaturescience.github.io/skater.\n"
+output:
+  BiocWorkflowTools::f1000_article: default
 bibliography: bibliography.bib
-output: BiocWorkflowTools::f1000_article
+keywords: bioinformatics, kinship, R, genealogy, SNPs, single nucleotide polymorphisms,
+  relatedness
 ---
@@
 Once estimated kinship is converted to degree, it may be of interest to compare the inferred degree to truth.
-When aggregated over many relationships and inferences, this method can help benchmark performance of a particular kinship analysis method.
+When aggregated over many relationships and inferences, this approach can help benchmark performance of a particular kinship analysis method.
 
@@
 A `vector` with the reciprocal RMSE (R-RMSE).
-The R-RMSE is calculated as `sqrt(mean((1/(Target+.5)-1/(Predicted+.5))^2)))`, and represents an alternative to classification accuracy when benchmarking relationship degree estimation.
+The R-RMSE represents an alternative to classification accuracy when benchmarking relationship degree estimation and is calculated using the formula in (1).
 Taking the reciprocal of the target and predicted degree results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree).
@@
 
-To illustrate the usage, this example will start with the `kinpairs` data from above and randomly flip ~20% of the true relationship degrees.
+\begin{equation}
+\sqrt{\frac{\sum_{i=1}^{k}(\frac{1}{\text{Target}+0.5}-\frac{1}{\text{Predicted}+0.5})^2}{k}}
+\end{equation}
+
+To illustrate the usage, this example will start with the `kinpairs` data from above and randomly flip ~20% of the true relationship degrees:
 
@@
 
-Next, running the `confusion_matrix()` function will return all four objects noted above.
+Next, running the `confusion_matrix()` function will return all four objects noted above:
 
@@
 
-SDT and VPN developed the R package.
+SDT, VPN, and MBS developed the R package.
 

### 2021-11-30 paper/paper.tex

@@
 \begin{abstract}
-\hfill\break
 \textbf{Motivation:} SNP-based kinship analysis with genome-wide relationship estimation and IBD segment analysis methods produces results that often require further downstream processing and manipulation.
 A dedicated software package that consistently and intuitively implements this analysis functionality is needed.\\
-\textbf{Results:} Here we present the skater R package for \textbf{S}NP-based \textbf{k}inship \textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation with \textbf{R}.
-The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree classification, and summarizing IBD segment data.\\
-\textbf{Availability:} The skater package is implemented as an R package and is released under the MIT license at https://github.com/signaturescience/skater.
-Documentation is available at https://signaturescience.github.io/skater.
+\textbf{a}nalysis, \textbf{t}esting, and \textbf{e}valuation with \textbf{R}.
+The skater package contains a suite of well-documented tools for importing, parsing, and analyzing pedigree data, performing relationship degree inference, benchmarking relationship degree package is implemented as an R package and is released under the MIT license at \textbf{Results:} Here we present the skater R package for \textbf{S}NP-based \textbf{k}inship classification, and summarizing IBD segment data.\\
+\textbf{Availability:} The skater https://github.com/signaturescience/skater. Documentation is available at https://signaturescience.github.io/skater.
 \end{abstract}
@@
 
-\textbf{R version}: R version 4.0.4 (2021-02-15)
+\textbf{R version}: R version 4.0.2 (2020-06-22)
 
@@
 \begin{Highlighting}[]
-\FunctionTok{install.packages}\NormalTok{(}\StringTok{"skater"}\NormalTok{)}
+\KeywordTok{install.packages}\NormalTok{(}\StringTok{"skater"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}
-\NormalTok{devtools}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{"signaturescience/skater"}\NormalTok{, }\AttributeTok{build\_vignettes=}\ConstantTok{TRUE}\NormalTok{)}
+\KeywordTok{install.packages}\NormalTok{(}\StringTok{"devtools"}\NormalTok{)}
+\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"signaturescience/skater"}\NormalTok{, }\DataTypeTok{build_vignettes=}\OtherTok{TRUE}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{library}\NormalTok{(skater)}
+\KeywordTok{library}\NormalTok{(skater)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{famfile }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"3gens.fam"}\NormalTok{, }\AttributeTok{package=}\StringTok{"skater"}\NormalTok{, }\AttributeTok{mustWork=}\ConstantTok{TRUE}\NormalTok{)}
-\NormalTok{fam }\OtherTok{\textless{}{-}} \FunctionTok{read\_fam}\NormalTok{(famfile)}
+\NormalTok{famfile <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"3gens.fam"}\NormalTok{, }\DataTypeTok{package=}\StringTok{"skater"}\NormalTok{, }\DataTypeTok{mustWork=}\OtherTok{TRUE}\NormalTok{)}
+\NormalTok{fam <-}\StringTok{ }\KeywordTok{read_fam}\NormalTok{(famfile)}
 \NormalTok{fam}
@@
 \begin{Highlighting}[]
-\NormalTok{peds }\OtherTok{\textless{}{-}} \FunctionTok{fam2ped}\NormalTok{(fam)}
+\NormalTok{peds <-}\StringTok{ }\KeywordTok{fam2ped}\NormalTok{(fam)}
+\end{Highlighting}
+\end{Shaded}
+
+\begin{verbatim}
+## Warning: replacing previous import 'vctrs::data_frame' by 'tibble::data_frame'
+## when loading 'dplyr'
+\end{verbatim}
+
+\begin{Shaded}
+\begin{Highlighting}[]
 \NormalTok{peds}
@@
 \begin{Highlighting}[]
-\NormalTok{peds }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{filter}\NormalTok{(fid}\SpecialCharTok{==}\StringTok{"testped1"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols=}\NormalTok{data)}
+\NormalTok{peds }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{filter}\NormalTok{(fid}\OperatorTok{==}\StringTok{"testped1"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols=}\NormalTok{data)}
 \end{Highlighting}
@@
 ##   <chr>    <chr>             <chr>             <chr>      <int>    <dbl> <list> 
-## 1 testped1 testped1_g1-b1-s1 <NA>              <NA>           1        1 <pedig~
-## 2 testped1 testped1_g1-b1-i1 <NA>              <NA>           2        1 <pedig~
-## 3 testped1 testped1_g2-b1-s1 <NA>              <NA>           1        1 <pedig~
+## 1 testped1 testped1_g1-b1-s1 NA                NA             1        1 <pedig~
+## 2 testped1 testped1_g1-b1-i1 NA                NA             2        1 <pedig~
+## 3 testped1 testped1_g2-b1-s1 NA                NA             1        1 <pedig~
 ## 4 testped1 testped1_g2-b1-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
-## 5 testped1 testped1_g2-b2-s1 <NA>              <NA>           1        1 <pedig~
+## 5 testped1 testped1_g2-b2-s1 NA                NA             1        1 <pedig~
 ## 6 testped1 testped1_g2-b2-i1 testped1_g1-b1-s1 testped1_~     2        1 <pedig~
@@
 \begin{Highlighting}[]
-\NormalTok{peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]]}
+\NormalTok{peds}\OperatorTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]]}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{plot}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]], }\AttributeTok{mar=}\FunctionTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{), }\AttributeTok{cex=}\NormalTok{.}\DecValTok{7}\NormalTok{)}
+\KeywordTok{plot}\NormalTok{(peds}\OperatorTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]], }\DataTypeTok{mar=}\KeywordTok{c}\NormalTok{(}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{,}\DecValTok{1}\NormalTok{,}\DecValTok{4}\NormalTok{), }\DataTypeTok{cex=}\NormalTok{.}\DecValTok{7}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{plot\_pedigree}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped, }\AttributeTok{file=}\StringTok{"3gens.ped.pdf"}\NormalTok{)}
+\KeywordTok{plot_pedigree}\NormalTok{(peds}\OperatorTok{$}\NormalTok{ped, }\DataTypeTok{file=}\StringTok{"3gens.ped.pdf"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{ped2kinpair}\NormalTok{(peds}\SpecialCharTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]])}
+\KeywordTok{ped2kinpair}\NormalTok{(peds}\OperatorTok{$}\NormalTok{ped[[}\DecValTok{1}\NormalTok{]])}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{kinpairs }\OtherTok{\textless{}{-}} 
-\NormalTok{  peds }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{pairs=}\NormalTok{purrr}\SpecialCharTok{::}\FunctionTok{map}\NormalTok{(ped, ped2kinpair)) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{select}\NormalTok{(fid, pairs) }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{unnest}\NormalTok{(}\AttributeTok{cols=}\NormalTok{pairs)}
+\NormalTok{kinpairs <-}\StringTok{ }
+\StringTok{  }\NormalTok{peds }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{pairs=}\NormalTok{purrr}\OperatorTok{::}\KeywordTok{map}\NormalTok{(ped, ped2kinpair)) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{select}\NormalTok{(fid, pairs) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{unnest}\NormalTok{(}\DataTypeTok{cols=}\NormalTok{pairs)}
 \NormalTok{kinpairs}
@@
 \begin{Highlighting}[]
-\FunctionTok{dibble}\NormalTok{()}
+\KeywordTok{dibble}\NormalTok{()}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{dibble}\NormalTok{(}\AttributeTok{max\_degree =} \DecValTok{5}\NormalTok{)}
+\KeywordTok{dibble}\NormalTok{(}\DataTypeTok{max_degree =} \DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{25}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)}
+\KeywordTok{kin2degree}\NormalTok{(.}\DecValTok{25}\NormalTok{, }\DataTypeTok{max_degree=}\DecValTok{3}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)}
+\KeywordTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\DataTypeTok{max_degree=}\DecValTok{3}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\AttributeTok{max\_degree=}\DecValTok{5}\NormalTok{)}
+\KeywordTok{kin2degree}\NormalTok{(.}\DecValTok{0312}\NormalTok{, }\DataTypeTok{max_degree=}\DecValTok{5}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{\# Get two pairs from each type of relationship we have in kinpairs:}
-\NormalTok{kinpairs\_subset }\OtherTok{\textless{}{-}} 
-\NormalTok{  kinpairs }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{group\_by}\NormalTok{(k) }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{slice}\NormalTok{(}\DecValTok{1}\SpecialCharTok{:}\DecValTok{2}\NormalTok{)}
-\NormalTok{kinpairs\_subset}
+\CommentTok{# Get two pairs from each type of relationship we have in kinpairs:}
+\NormalTok{kinpairs_subset <-}\StringTok{ }
+\StringTok{  }\NormalTok{kinpairs }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{group_by}\NormalTok{(k) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{slice}\NormalTok{(}\DecValTok{1}\OperatorTok{:}\DecValTok{2}\NormalTok{)}
+\NormalTok{kinpairs_subset}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\CommentTok{\# Infer degree out to third degree relatives:}
-\NormalTok{kinpairs\_subset }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree=}\FunctionTok{kin2degree}\NormalTok{(k, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{))}
+\CommentTok{# Infer degree out to third degree relatives:}
+\NormalTok{kinpairs_subset }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{degree=}\KeywordTok{kin2degree}\NormalTok{(k, }\DataTypeTok{max_degree=}\DecValTok{3}\NormalTok{))}
 \end{Highlighting}
@@
   A \texttt{vector} with the reciprocal RMSE (R-RMSE).
-The R-RMSE is calculated as \texttt{sqrt(mean((1/(Target+.5)-1/(Predicted+.5))\^{}2)))}, and represents an alternative to classification accuracy when benchmarking relationship degree estimation.
+The R-RMSE represents an alternative to classification accuracy when benchmarking relationship degree estimation and is calculated using the formula in (1).
 Taking the reciprocal of the target and predicted degree results in larger penalties for more egregious misclassifications (e.g., classifying a first-degree relative pair as second degree) than misclassifications at more distant relationships (e.g., misclassifying a fourth-degree relative pair as fifth-degree).
@@
 
-To illustrate the usage, this example will start with the \texttt{kinpairs} data from above and randomly flip \textasciitilde20\% of the true relationship degrees.
+\begin{equation}
+\sqrt{\frac{\sum_{i=1}^{k}(\frac{1}{\text{Target}+0.5}-\frac{1}{\text{Predicted}+0.5})^2}{k}}
+\end{equation}
+
+To illustrate the usage, this example will start with the \texttt{kinpairs} data from above and randomly flip \textasciitilde20\% of the true relationship degrees:
 
@@
 \begin{Highlighting}[]
-\CommentTok{\# Function to randomly flip levels of a factor (at 20\%, by default)}
-\NormalTok{randomflip }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(x, }\AttributeTok{p=}\NormalTok{.}\DecValTok{2}\NormalTok{) }\FunctionTok{ifelse}\NormalTok{(}\FunctionTok{runif}\NormalTok{(}\FunctionTok{length}\NormalTok{(x))}\SpecialCharTok{\textless{}}\NormalTok{p, }\FunctionTok{sample}\NormalTok{(}\FunctionTok{unique}\NormalTok{(x)), x)}
+\CommentTok{# Function to randomly flip levels of a factor (at 20%, by default)}
+\NormalTok{randomflip <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x, }\DataTypeTok{p=}\NormalTok{.}\DecValTok{2}\NormalTok{) }\KeywordTok{ifelse}\NormalTok{(}\KeywordTok{runif}\NormalTok{(}\KeywordTok{length}\NormalTok{(x))}\OperatorTok{<}\NormalTok{p, }\KeywordTok{sample}\NormalTok{(}\KeywordTok{unique}\NormalTok{(x)), x)}
 
-\CommentTok{\# Infer degree (truth/target) using kin2degree, then randomly flip 20\% of them}
-\FunctionTok{set.seed}\NormalTok{(}\DecValTok{42}\NormalTok{)}
-\NormalTok{kinpairs\_inferred }\OtherTok{\textless{}{-}}\NormalTok{ kinpairs }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_truth=}\FunctionTok{kin2degree}\NormalTok{(k, }\AttributeTok{max\_degree=}\DecValTok{3}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_truth=}\NormalTok{tidyr}\SpecialCharTok{::}\FunctionTok{replace\_na}\NormalTok{(degree\_truth, }\StringTok{"unrelated"}\NormalTok{)) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate}\NormalTok{(}\AttributeTok{degree\_inferred=}\FunctionTok{randomflip}\NormalTok{(degree\_truth))}
-\NormalTok{kinpairs\_inferred}
+\CommentTok{# Infer degree (truth/target) using kin2degree, then randomly flip 20% of them}
+\KeywordTok{set.seed}\NormalTok{(}\DecValTok{42}\NormalTok{)}
+\NormalTok{kinpairs_inferred <-}\StringTok{ }\NormalTok{kinpairs }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{degree_truth=}\KeywordTok{kin2degree}\NormalTok{(k, }\DataTypeTok{max_degree=}\DecValTok{3}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{degree_truth=}\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{replace_na}\NormalTok{(degree_truth, }\StringTok{"unrelated"}\NormalTok{)) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate}\NormalTok{(}\DataTypeTok{degree_inferred=}\KeywordTok{randomflip}\NormalTok{(degree_truth))}
+\NormalTok{kinpairs_inferred}
 \end{Highlighting}
@@
 ##    <chr>    <chr>             <chr>           <dbl> <chr>        <chr>          
-##  1 testped1 testped1_g1-b1-s1 testped1_g1-b1-s1 0.5   0            0              
-##  2 testped1 testped1_g1-b1-i1 testped1_g1-b1-s1 0     unrelated    unrelated      
-##  3 testped1 testped1_g1-b1-s1 testped1_g2-b1-s1 0     unrelated    unrelated      
-##  4 testped1 testped1_g1-b1-s1 testped1_g2-b1-i1 0.25  1            1              
-##  5 testped1 testped1_g1-b1-s1 testped1_g2-b2-s1 0     unrelated    unrelated      
-##  6 testped1 testped1_g1-b1-s1 testped1_g2-b2-i1 0.25  1            1              
-##  7 testped1 testped1_g1-b1-s1 testped1_g3-b1-i1 0.125 2            2              
-##  8 testped1 testped1_g1-b1-s1 testped1_g3-b2-i1 0.125 2            1              
-##  9 testped1 testped1_g1-b1-i1 testped1_g1-b1-i1 0.5   0            0              
-## 10 testped1 testped1_g1-b1-i1 testped1_g2-b1-s1 0     unrelated    unrelated      
+##  1 testped1 testped1_g1-b1-s1 testped1_g1-b1~ 0.5   0            0              
+##  2 testped1 testped1_g1-b1-i1 testped1_g1-b1~ 0     unrelated    unrelated      
+##  3 testped1 testped1_g1-b1-s1 testped1_g2-b1~ 0     unrelated    unrelated      
+##  4 testped1 testped1_g1-b1-s1 testped1_g2-b1~ 0.25  1            1              
+##  5 testped1 testped1_g1-b1-s1 testped1_g2-b2~ 0     unrelated    unrelated      
+##  6 testped1 testped1_g1-b1-s1 testped1_g2-b2~ 0.25  1            1              
+##  7 testped1 testped1_g1-b1-s1 testped1_g3-b1~ 0.125 2            2              
+##  8 testped1 testped1_g1-b1-s1 testped1_g3-b2~ 0.125 2            1              
+##  9 testped1 testped1_g1-b1-i1 testped1_g1-b1~ 0.5   0            0              
+## 10 testped1 testped1_g1-b1-i1 testped1_g2-b1~ 0     unrelated    unrelated      
 ## # ... with 278 more rows
@@
 
-Next, running the \texttt{confusion\_matrix()} function will return all four objects noted above.
+Next, running the \texttt{confusion\_matrix()} function will return all four objects noted above:
 
@@
 \begin{Highlighting}[]
-\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
-                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth)}
+\KeywordTok{confusion_matrix}\NormalTok{(}\DataTypeTok{prediction =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_inferred, }
+                 \DataTypeTok{target =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_truth)}
 \end{Highlighting}
@@
 ## 4 3        8              0.75             0.936           0.25  0.992     0.375
-## 5 unrelated  96              0.854            0.958           0.911 0.929     0.882
-## 6 Average    57.6            0.799            0.954           0.736 0.951     0.741
+## 5 unrel~  96              0.854            0.958           0.911 0.929     0.882
+## 6 Avera~  57.6            0.799            0.954           0.736 0.951     0.741
 ## # ... with 8 more variables: Prevalence <dbl>, Detection Rate <dbl>,
@@
 \begin{Highlighting}[]
-\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
-                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth) }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  purrr}\SpecialCharTok{::}\FunctionTok{pluck}\NormalTok{(}\StringTok{"Table"}\NormalTok{)}
+\KeywordTok{confusion_matrix}\NormalTok{(}\DataTypeTok{prediction =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_inferred, }
+                 \DataTypeTok{target =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_truth) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{purrr}\OperatorTok{::}\KeywordTok{pluck}\NormalTok{(}\StringTok{"Table"}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\FunctionTok{confusion\_matrix}\NormalTok{(}\AttributeTok{prediction =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_inferred, }
-                 \AttributeTok{target =}\NormalTok{ kinpairs\_inferred}\SpecialCharTok{$}\NormalTok{degree\_truth, }
-                 \AttributeTok{longer =} \ConstantTok{TRUE}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  purrr}\SpecialCharTok{::}\FunctionTok{pluck}\NormalTok{(}\StringTok{"Other"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  tidyr}\SpecialCharTok{::}\FunctionTok{spread}\NormalTok{(Class, Value) }\SpecialCharTok{\%\textgreater{}\%} 
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{relocate}\NormalTok{(Average, }\AttributeTok{.after=}\NormalTok{dplyr}\SpecialCharTok{::}\FunctionTok{last\_col}\NormalTok{()) }\SpecialCharTok{\%\textgreater{}\%}
-\NormalTok{  dplyr}\SpecialCharTok{::}\FunctionTok{mutate\_if}\NormalTok{(rlang}\SpecialCharTok{::}\NormalTok{is\_double, signif, }\DecValTok{2}\NormalTok{)}
+\KeywordTok{confusion_matrix}\NormalTok{(}\DataTypeTok{prediction =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_inferred, }
+                 \DataTypeTok{target =}\NormalTok{ kinpairs_inferred}\OperatorTok{$}\NormalTok{degree_truth, }
+                 \DataTypeTok{longer =} \OtherTok{TRUE}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{purrr}\OperatorTok{::}\KeywordTok{pluck}\NormalTok{(}\StringTok{"Other"}\NormalTok{) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{tidyr}\OperatorTok{::}\KeywordTok{spread}\NormalTok{(Class, Value) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{relocate}\NormalTok{(Average, }\DataTypeTok{.after=}\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{last_col}\NormalTok{()) }\OperatorTok{%>%}\StringTok{ }
+\StringTok{  }\NormalTok{dplyr}\OperatorTok{::}\KeywordTok{mutate_if}\NormalTok{(rlang}\OperatorTok{::}\NormalTok{is_double, signif, }\DecValTok{2}\NormalTok{)}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{hapibd\_filepath }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"GBR.sim.ibd.gz"}\NormalTok{, }
-                               \AttributeTok{package=}\StringTok{"skater"}\NormalTok{)}
-\NormalTok{hapibd\_seg }\OtherTok{\textless{}{-}} \FunctionTok{read\_ibd}\NormalTok{(hapibd\_filepath, }\AttributeTok{source =} \StringTok{"hapibd"}\NormalTok{)}
-\NormalTok{hapibd\_seg}
+\NormalTok{hapibd_filepath <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"GBR.sim.ibd.gz"}\NormalTok{, }
+                               \DataTypeTok{package=}\StringTok{"skater"}\NormalTok{)}
+\NormalTok{hapibd_seg <-}\StringTok{ }\KeywordTok{read_ibd}\NormalTok{(hapibd_filepath, }\DataTypeTok{source =} \StringTok{"hapibd"}\NormalTok{)}
+\NormalTok{hapibd_seg}
 \end{Highlighting}
@@
 \begin{Highlighting}[]
-\NormalTok{gmap\_filepath }\OtherTok{\textless{}{-}} \FunctionTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"sexspec{-}avg{-}min.plink.map"}\NormalTok{, }
-                             \AttributeTok{package=}\StringTok{"skater"}\NormalTok{)}
-\NormalTok{gmap }\OtherTok{\textless{}{-}} \FunctionTok{read\_map}\NormalTok{(gmap\_filepath)}
+\NormalTok{gmap_filepath <-}\StringTok{ }\KeywordTok{system.file}\NormalTok{(}\StringTok{"extdata"}\NormalTok{, }\StringTok{"sexspec-avg-min.plink.map"}\NormalTok{, }
+                             \DataTypeTok{package=}\StringTok{"skater"}\NormalTok{)}
+\NormalTok{gmap <-}\StringTok{ }\KeywordTok{read_map}\NormalTok{(gmap_filepath)}
 \NormalTok{gmap}
@@
 \begin{Highlighting}[]
-\NormalTok{ibd\_dat }\OtherTok{\textless{}{-}} \FunctionTok{ibd2kin}\NormalTok{(}\AttributeTok{.ibd\_data=}\NormalTok{hapibd\_seg, }\AttributeTok{.map=}\NormalTok{gmap)}
-\NormalTok{ibd\_dat}
+\NormalTok{ibd_dat <-}\StringTok{ }\KeywordTok{ibd2kin}\NormalTok{(}\DataTypeTok{.ibd_data=}\NormalTok{hapibd_seg, }\DataTypeTok{.map=}\NormalTok{gmap)}
+\NormalTok{ibd_dat}
 \end{Highlighting}

### 2021-11-30 paper/paper.Rmd
fix abstract

--- 2021-11-29 paper/paper.Rmd
@@
 - name: Jianye Ge
-    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
+  affiliation: Center for Human Identification, Department of Microbiology, Immunology,
+    and Genetics, University of North Texas Health Science Center, Fort Worth, TX
+    76107, USA.
 - name: August E. Woerner
-    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
+  affiliation: Center for Human Identification, Department of Microbiology, Immunology,
+    and Genetics, University of North Texas Health Science Center, Fort Worth, TX
+    76107, USA.
 - name: Bruce Budowle
-    affiliation: Center for Human Identification, Department of Microbiology, Immunology, and Genetics, University of North Texas Health Science Center, Fort Worth, TX 76107, USA. 
-# Abstract up to 300 words. No citations, minimal abbreviations.
+  affiliation: Center for Human Identification, Department of Microbiology, Immunology,
+    and Genetics, University of North Texas Health Science Center, Fort Worth, TX
+    76107, USA.
 abstract: | 
@@
 Documentation is available at https\:\/\/signaturescience.github.io/skater.
+output:
+  BiocWorkflowTools::f1000_article: default
+bibliography: bibliography.bib
 keywords: bioinformatics, kinship, R, genealogy, SNPs, single nucleotide polymorphisms, relatedness
-bibliography: bibliography.bib
-output: BiocWorkflowTools::f1000_article
 ---

### 2021-12-01 R/ibd.R
explanation of ibd2kin math

@@ -2,14 +2,14 @@
 #'
 #' @description
 #' This function is used to retrieve a relatedness measure from IBD segments.
-#' The relatedness value returned is the kinship coefficent.
+#' The relatedness value returned is the kinship coefficient.
 #'
 #' @details
 #' The input data should be pairwise IBD segments prepared via
 #' \link[skater]{read_ibd}. The function will internally loop over each
 #' chromosome, and use a specified genetic map to convert shared segments to
 #' genetic units. After doing so, the function converts the shared length to a
-#' kinship coefficient.
+#' kinship coefficient by summing \eqn{0.5*IBD2 + 0.25*IBD1}.
 #'
 #' Note that the data read in by [read_ibd] when `source="pedsim"` returns a
 #' list with separate tibbles for IBD1 and IBD2 segments. The current

### 2021-12-01 R/confusion_matrix.R
stop exporting internal functions

@@ -18,11 +18,12 @@ utils::globalVariables(c('Class', 'Positive', 'N Positive', 'N Negative', 'N'))
 #' @seealso \code{\link{binom.test}}
 #'
 #' @examples
+#' \dontrun{
 #' p = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
 #' o = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
 #' calc_accuracy(table(p, o))
+#' }
 #'
-#' @export
 calc_accuracy <- function(tabble) {
 
   acc <- sum(diag(tabble))/sum(tabble)
@@ -140,11 +141,12 @@ calc_accuracy <- function(tabble) {
 #' reduction.," \emph{Genetic Epidemiology}, vol 4, 306.
 #'
 #' @examples
+#' \dontrun{
 #' p = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
 #' o = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
 #' calc_stats(table(p, o), positive='a')
+#' }
 #'
-#' @export
 calc_stats <- function(tabble, prevalence = NULL, positive, ...) {
   # checks
   # using original all.equal checks will fail
@@ -302,7 +304,6 @@ calc_stats <- function(tabble, prevalence = NULL, positive, ...) {
 #'   \code{confusionMatrix} function from the \code{caret} package.
 #'
 #' @seealso \code{\link{calc_accuracy}} \code{\link{calc_stats}}
-#' \code{\link{confusion_matrix}}
 #'
 #' @return A list of tibble(s) with the associated statistics and possibly the
 #'   frequency table as list column of the first element. If classes contain >1

### 2021-12-01 R/utils.R
make class of return value explicit

@@ -46,7 +46,7 @@ dibble <- function(max_degree=3L) {
 #'   [dibble][skater::dibble]. Anything below the inference range of
 #'   `max_degree` will report `NA`. See [dibble][skater::dibble].
 #'
-#' @return Inferred degree, up to the maximum degree in `dibble` (anything more distant is `NA`, i.e., unrelated).
+#' @return A vector with inferred degree, up to the maximum degree in `dibble` (anything more distant is `NA`, i.e., unrelated).
 #'
 #' @examples
 #' kin2degree(0.5)
@@ -87,7 +87,7 @@ kin2degree <- function(k, max_degree=3L) {
 #'
 #' @param k Kinship coefficient (numeric, typically between 0 and .5, although KING can produce values <0).
 #'
-#' @return Estimated cM, ranging from 0-3560.
+#' @return A vector of numeric estimated cM, ranging from 0-3560.
 #'
 #' @examples
 #' kin2cm(.25)

### 2021-12-03 R/confusion_matrix.R
remove examples for unexported functions

@@ -17,12 +17,10 @@ utils::globalVariables(c('Class', 'Positive', 'N Positive', 'N Negative', 'N'))
 #'
 #' @seealso \code{\link{binom.test}}
 #'
-#' @examples
-#' \dontrun{
-#' p = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
-#' o = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
-#' calc_accuracy(table(p, o))
-#' }
+# #' @examples
+# #' p = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
+# #' o = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
+# #' calc_accuracy(table(p, o))
 #'
 calc_accuracy <- function(tabble) {
 
@@ -140,12 +138,10 @@ calc_accuracy <- function(tabble) {
 #' modeling in imbalanced datasets using multifactor dimensionality
 #' reduction.," \emph{Genetic Epidemiology}, vol 4, 306.
 #'
-#' @examples
-#' \dontrun{
-#' p = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
-#' o = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
-#' calc_stats(table(p, o), positive='a')
-#' }
+# #' @examples
+# #' p = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
+# #' o = sample(letters[1:4], 250, replace = TRUE, prob = 1:4)
+# #' calc_stats(table(p, o), positive='a')
 #'
 calc_stats <- function(tabble, prevalence = NULL, positive, ...) {
   # checks

### 2021-12-06 README.Rmd
add badges

@@
 <!-- badges: start -->
+
+[![CRAN status](https://www.r-pkg.org/badges/version/skater)](https://CRAN.R-project.org/package=skater)
+
+[![biorXiv](https://img.shields.io/badge/biorXiv-10.1101%2F2021.07.21.453083-red)](https://www.biorxiv.org/content/10.1101/2021.07.21.453083v1)
+
+[![DOI](https://zenodo.org/badge/339462170.svg)](https://zenodo.org/badge/latestdoi/339462170)
 

### 2021-12-06 paper/paper.Rmd
boundary explanation thanks @ahhgust
updated paper with zenodo doi

@@
 Anything beyond the maximum degree resolution is considered unrelated (degree `NA`).
+Note also that while the theoretical upper boundary for the kinship coefficient is 0.5, the inference range for 0-degree (same person or identical twins) extends to 1 to allow for floating point arithmetic and stochastic effects resulting in kinship coefficients above 0.5.
 
@@
 1. Source code available from: <https://github.com/signaturescience/skater>.
-1. Archived source code at time of publication: **FIXME Zenodo DOI to come here**.
+1. Archived source code at time of publication: <https://doi.org/10.5281/zenodo.5761996>.
 1. Software license: MIT License.

### 2021-12-06 paper/paper.tex
updated paper with zenodo doi

@@
 Anything beyond the maximum degree resolution is considered unrelated (degree \texttt{NA}).
+Note also that while the theoretical upper boundary for the kinship coefficient is 0.5, the inference range for 0-degree (same person or identical twins) extends to 1 to allow for floating point arithmetic and stochastic effects resulting in kinship coefficients above 0.5.
 
@@
 \item
-  Archived source code at time of publication: \textbf{FIXME Zenodo DOI to come here}.
+  Archived source code at time of publication: \url{https://doi.org/10.5281/zenodo.5761996}.
 \item

### 2022-01-12 paper/paper.tex
figure caption

@@
 
-\textbf{skater package version}: 0.1.0
+\textbf{skater package version}: 0.1.1
 
@@
 
-\begin{center}\includegraphics{paper_files/figure-latex/plotped-1} \end{center}
+\begin{figure}
+
+{\centering \includegraphics{paper_files/figure-latex/plotped-1}
+
+}
+
+\caption{Pedigree diagram for the first family in the pedigree shown in the code above.}\label{fig:plotped}
+\end{figure}
 
